---
title: "calculate_HF_performance_2024"
output: html_notebook
---

# Load libraries
```{r}
library(caret)
library(data.table)
library(dplyr)
library(sets)
library(scales)
library(tidyr)
library(stringr)
library(tibble)
library(stringi)
library(ggplot2)

library(psych)
library(tidyverse)
library(irr)
library(RColorBrewer)

library(readr) # For read_tsv

```


# Model input files: 
- input file:  
    - Circ_allstates_includemissing_modelinput_DF1.csv
- output file:



# Model Circ HF results: 
- input file: DF1_results_individualfolds_Circ_2024.csv
- 

- train file: private_CHIL_repo/scripts/bal_allfeats_MLHC_nosmote_sgs_evaluate_baselines_nestKfold.py
- output file: private_CHIL_repo/scripts/3-21-ACC/Circ_HF_results_individualfolds.csv

```{r}

fairness_full_hf_results <- read.csv("../scripts/individualfolds_Circ_HF_results_fairness_final.csv") # individualfolds_Circ_HF_results_fairness_final

full_hf_results <- read.csv("../scripts/final_results_individualfolds_Circ_2024.csv") # 

```

# Subset results: 
```{r}

logr_hf_results <- full_hf_results[full_hf_results$model == "logr", ]
xgb_hf_results <- full_hf_results[full_hf_results$model == "xgb", ]
rf_hf_results <- full_hf_results[full_hf_results$model == "rf", ]



fairness_full_hf_results <- fairness_full_hf_results[fairness_full_hf_results$sg_key == "race", ]

fair_logr_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "logr", ]
fair_xgb_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "xgb", ]
fair_rf_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "rf", ]



full_hf_results
fairness_full_hf_results
```

# Print md table of all feature set results: 
- AUC
- F1
# Subgroups: blackorwhite, black, white.
```{r}
# Required library for markdown tables

#curr_model_results <- xgb_hf_results # logr_hf_results, xgb_hf_results

library(dplyr)
library(tidyr)
library(knitr)

#     "M1",
#        "M2",
#        "M3_county_DF1_nm",
#        "M4_DF1_nm",
#        "M5",
#        "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
#    
#        "DF1_nm_county_AHRQ_domain1",
#        "DF1_nm_county_AHRQ_domain2",
#        "DF1_nm_county_AHRQ_domain3",
#        "DF1_nm_county_AHRQ_domain4",
#        "DF1_nm_county_AHRQ_domain5",
#
#        "DF1_nm_tract_AHRQ_domain1",
#        "DF1_nm_tract_AHRQ_domain2",
#        "DF1_nm_tract_AHRQ_domain3",
#        "DF1_nm_tract_AHRQ_domain4",
#        "DF1_nm_tract_AHRQ_domain5"

feat_sets <- c(
   "M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm",
        "M5",
        "M6_total_ahrq_cty_DF1_nm")


sdoh_feat_sets <- c(
    "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
    
        "DF1_nm_county_AHRQ_domain1",
        "DF1_nm_county_AHRQ_domain2",
        "DF1_nm_county_AHRQ_domain3",
        "DF1_nm_county_AHRQ_domain4",
        "DF1_nm_county_AHRQ_domain5",

        "DF1_nm_tract_AHRQ_domain1",
        "DF1_nm_tract_AHRQ_domain2",
        "DF1_nm_tract_AHRQ_domain3",
        "DF1_nm_tract_AHRQ_domain4",
        "DF1_nm_tract_AHRQ_domain5")


# Ensure all required subgroups are present in the data
grouped_results <- curr_model_results %>%
  filter(feat %in% sdoh_feat_sets, subgroup %in% c("blackorwhite", "black", "white")) %>%
  group_by(feat, subgroup) %>%
  summarize(
    auc = mean(auc, na.rm = TRUE),
    f1 = mean(f1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = subgroup,
    values_from = c(auc, f1),
    names_glue = "{subgroup}_{.value}"
  )

# Reorder columns in grouped_results
grouped_results <- grouped_results %>%
  select(
    feat, 
    blackorwhite_auc, black_auc, white_auc, 
    blackorwhite_f1, black_f1, white_f1
  )

# Format numeric columns to three decimal places
grouped_results <- grouped_results %>%
  mutate(
    across(
      starts_with(c("blackorwhite_", "black_", "white_")),
      ~ sprintf("%.3f", .) # Format as a string with 3 decimal places
    )
  )

# Create the markdown table
kable(
  grouped_results,
  col.names = c(
    "Feature Set", 
    "AUROC (All Patients)", "AUROC (Black)", "AUROC (White)", 
    "F1 (All Patients)", "F1 (Black)", "F1 (White)"
  ),
  caption = "Summary of AUROC and F1 Score by Feature Set for Each Subgroup"
)


colnames(logr_hf_results)
```




# Print with Wilcoxon test results:
```{r}
library(dplyr)
library(tidyr)
library(knitr)

# Prepare the results dataframe
curr_model_results <- xgb_hf_results # or logr_hf_results

feat_sets <- c(
   "M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm_demo",
        "M5_and_demo",
        "M6_total_ahrq_cty_DF1_nm")


county_only_feats <- c("M6_total_ahrq_cty_DF1_nm",
        "DF1_nm_county_AHRQ_domain1",
        "DF1_nm_county_AHRQ_domain2",
        "DF1_nm_county_AHRQ_domain3",
        "DF1_nm_county_AHRQ_domain4",
        "DF1_nm_county_AHRQ_domain5")

tract_only_feats <- c( "M6_total_ahrq_trct_DF1_nm",
        "DF1_nm_tract_AHRQ_domain1",
        "DF1_nm_tract_AHRQ_domain2",
        "DF1_nm_tract_AHRQ_domain3",
        "DF1_nm_tract_AHRQ_domain4",
        "DF1_nm_tract_AHRQ_domain5")

# Prepare a dataframe to store Wilcoxon test results
wilcoxon_results <- data.frame(
  feature_set = character(),
  subgroup = character(),
  metric = character(),
  p_value = numeric(),
  stringsAsFactors = FALSE
)

# Perform Wilcoxon signed-rank test for each feature set compared to baseline
for (current_feat in county_only_feats[-1]) {  # Skip the baseline feature set
  # Prepare baseline data (M1)
  baseline_data <- curr_model_results %>%
    filter(feat == "M1")
  
  # Prepare current feature set data
  current_data <- curr_model_results %>%
    filter(feat == current_feat)
  
  # Subgroups to analyze
  subgroups <- c("blackorwhite", "black", "white")
  
  # Metrics to test
  metrics <- c("auc", "f1")
  
  # Perform Wilcoxon signed-rank test for each subgroup and metric
  for (subgroup in subgroups) {
    for (metric in metrics) {
      # Ensure data is properly matched by fold
      baseline_fold_metrics <- baseline_data %>%
        filter(subgroup == !!subgroup) %>%
        arrange(fold) %>%
        pull(!!metric)
      
      current_fold_metrics <- current_data %>%
        filter(subgroup == !!subgroup) %>%
        arrange(fold) %>%
        pull(!!metric)
      
      # Perform Wilcoxon signed-rank test
      wilcox_test <- wilcox.test(
        baseline_fold_metrics, 
        current_fold_metrics, 
        paired = TRUE, 
        alternative = "two.sided"
      )
      
      # Store results
      wilcoxon_results <- rbind(wilcoxon_results, data.frame(
        feature_set = current_feat,
        subgroup = subgroup,
        metric = metric,
        p_value = wilcox_test$p.value
      ))
    }
  }
}

# Prepare the table of Wilcoxon test results
wilcoxon_table <- wilcoxon_results %>%
  pivot_wider(
    names_from = c(subgroup, metric),
    values_from = p_value,
    names_glue = "{subgroup}_{metric}_p"
  ) %>%
  mutate(across(where(is.numeric), ~ sprintf("%.3f", .)))

# Print the Wilcoxon signed-rank test results
kable(
  wilcoxon_table,
  col.names = c(
    "Feature Set", 
    "All Patients AUROC p", "All Patients F1 p",
    "Black AUROC p", "Black F1 p",
    "White AUROC p", "White F1 p"
  ),
  caption = "Wilcoxon Signed-Rank Test P-values Comparing Feature Sets to Baseline"
)
```






# [PERFORMANCE] Print md table: for all 3 models:
```{r}
# Combine the logr, rf, and xgb results
curr_model_results <- bind_rows(
  logr_hf_results %>% mutate(model = "logr"),
  rf_hf_results %>% mutate(model = "rf"),
  xgb_hf_results %>% mutate(model = "xgb")
)


feat_sets <- c(
   "M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm",
        "M5",
        "M6_total_ahrq_cty_DF1_nm")

sdoh_feat_sets <- c(
  "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
  "DF1_nm_county_AHRQ_domain1", "DF1_nm_county_AHRQ_domain2", "DF1_nm_county_AHRQ_domain3",
  "DF1_nm_county_AHRQ_domain4", "DF1_nm_county_AHRQ_domain5", "DF1_nm_tract_AHRQ_domain1",
  "DF1_nm_tract_AHRQ_domain2", "DF1_nm_tract_AHRQ_domain3", "DF1_nm_tract_AHRQ_domain4",
  "DF1_nm_tract_AHRQ_domain5"
)

# Ensure all required subgroups are present in the data
grouped_results <- curr_model_results %>%
  filter(feat %in% sdoh_feat_sets, subgroup %in% c("blackorwhite", "black", "white")) %>%
  group_by(feat, model) %>%
  summarize(
    auroc_all = mean(auc[subgroup == "blackorwhite"], na.rm = TRUE),
    auroc_black = mean(auc[subgroup == "black"], na.rm = TRUE), 
    auroc_white = mean(auc[subgroup == "white"], na.rm = TRUE),
    f1_all = mean(f1[subgroup == "blackorwhite"], na.rm = TRUE),
    f1_black = mean(f1[subgroup == "black"], na.rm = TRUE),
    f1_white = mean(f1[subgroup == "white"], na.rm = TRUE),
    .groups = "drop"
  )

# Format numeric columns to three decimal places
grouped_results <- grouped_results %>%
  mutate(
    across(
      starts_with(c("auroc_", "f1_")),
      ~ sprintf("%.3f", .)
    )
  )

# Create the markdown table
kable(
  grouped_results,
  col.names = c(
    "Feature Set", "Model",
    "AUROC (All Patients)", "AUROC (Black)", "AUROC (White)",
    "F1 (All Patients)", "F1 (Black)", "F1 (White)"
  ),
  caption = "Summary of AUROC and F1 Score by Feature Set for Each Subgroup", format = "markdown"
)

```

# Fairness results:
```{r}
# Required libraries
library(dplyr)
library(tidyr)
library(knitr)

# Define feature sets
feat_sets <- c(
    "M1",
    "M2",
    "M3_county_DF1_nm",
    "M4_DF1_nm",
    "M5",
    "M6_total_ahrq_cty_DF1_nm"
)

sdoh_feat_sets <- c(
    "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
    "DF1_nm_county_AHRQ_domain1",
    "DF1_nm_county_AHRQ_domain2",
    "DF1_nm_county_AHRQ_domain3",
    "DF1_nm_county_AHRQ_domain4",
    "DF1_nm_county_AHRQ_domain5",
    "DF1_nm_tract_AHRQ_domain1",
    "DF1_nm_tract_AHRQ_domain2",
    "DF1_nm_tract_AHRQ_domain3",
    "DF1_nm_tract_AHRQ_domain4",
    "DF1_nm_tract_AHRQ_domain5"
)

# Filter, group, and summarize results
grouped_results <- curr_model_results %>%
  filter(feat %in% sdoh_feat_sets) %>%
  group_by(feat, fold) %>%
  summarize(
    eo_ratio = mean(eo_ratio, na.rm = TRUE),
    fpr_parity = mean(fpr_parity, na.rm = TRUE),
    tpr_parity = mean(tpr_parity, na.rm = TRUE),
    fnr_parity = mean(fnr_parity, na.rm = TRUE),
    dp_ratio = mean(dp_ratio, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(feat) %>%
  summarize(
    eo_ratio = round(mean(eo_ratio, na.rm = TRUE), 3),
    fpr_parity = round(mean(fpr_parity, na.rm = TRUE), 3),
    tpr_parity = round(mean(tpr_parity, na.rm = TRUE), 3),
    fnr_parity = round(mean(fnr_parity, na.rm = TRUE), 3),
    dp_ratio = round(mean(dp_ratio, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  select(feat, eo_ratio, fpr_parity, tpr_parity, fnr_parity, dp_ratio) # Reorder columns

# Create the markdown table
kable(
  grouped_results,
  col.names = c(
    "Feature Set", 
    "EO Ratio", "FPR Parity", "TPR Parity", 
    "FNR Parity", "DP Ratio"
  ),
  caption = "Summary of fairness by Feature Set for all patients (Black vs. White)"
)


```

# [FAIRNESS] Print md table: for all 3 models:
```{r}


fairness_full_hf_results <- fairness_full_hf_results[fairness_full_hf_results$sg_key == "race", ]

fair_logr_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "logr", ]
fair_xgb_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "xgb", ]
fair_rf_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "rf", ]


# Combine the logr, rf, and xgb results
curr_model_results_fair <- bind_rows(
  fair_logr_hf_results %>% mutate(model = "logr"),
  fair_xgb_hf_results %>% mutate(model = "rf"),
  fair_rf_hf_results %>% mutate(model = "xgb")
)


feat_sets <- c(
   "M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm",
        "M5",
        "M6_total_ahrq_cty_DF1_nm")

sdoh_feat_sets <- c(
  "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
  "DF1_nm_county_AHRQ_domain1", "DF1_nm_county_AHRQ_domain2", "DF1_nm_county_AHRQ_domain3",
  "DF1_nm_county_AHRQ_domain4", "DF1_nm_county_AHRQ_domain5", "DF1_nm_tract_AHRQ_domain1",
  "DF1_nm_tract_AHRQ_domain2", "DF1_nm_tract_AHRQ_domain3", "DF1_nm_tract_AHRQ_domain4",
  "DF1_nm_tract_AHRQ_domain5"
)

fairness_full_hf_results$dp_ratio

# Ensure all required subgroups are present in the data
grouped_results <- curr_model_results_fair %>%
  filter(feat %in% sdoh_feat_sets) %>%
  group_by(feat, model) %>%
  summarize(
    eo_ratio = mean(eo_ratio[sg_key == "race"], na.rm = TRUE),
    fpr_parity = mean(fpr_parity[sg_key == "race"], na.rm = TRUE), 
    tpr_parity = mean(tpr_parity[sg_key == "race"], na.rm = TRUE), 
    fnr_ratio = mean(fnr_parity[sg_key == "race"], na.rm = TRUE), 
    dp_ratio = mean(dp_ratio[sg_key == "race"], na.rm = TRUE), 
    .groups = "drop"
  )

# Format numeric columns to three decimal places
grouped_results <- grouped_results %>%
  mutate(
    across(
      starts_with(c("eo_", "fpr_", "tpr_", "fnr_", "dp_")),
      ~ sprintf("%.3f", .)
    )
  )

# Create the markdown table
kable(
  grouped_results,
  col.names = c(
    "Feature Set", "Model",
    "EO Ratio", "FPR Ratio", "TPR Ratio", "FNR Ratio",
    "DP Ratio"
  ),
  caption = "Summary of fairness metrics by model and feature set.", format = "markdown"
)

```

























# Below, create plots of fairness vs. predictive performance.

```{r}

fair_hf_results <- read.csv("../scripts/individualfolds_Circ_HF_results_fairness_final.csv") # "../v3_fairness_circulation_results.csv"
fair_hf_results

```

# 1.) Fairness:

sk_key: 
- race ** Main focus (self-reported race of Black vs. White)
- sex
- combos
- age_quartile
- hf_types
- insur_groups
- CCI_groups
- CCI_quartile

- Hypertension_groups
- CAD_groups
- CKD_groups

```{r}

# Filter out sg_key == "all_patients".
fair_hf_results <- fair_hf_results[fair_hf_results$sg_key != "all_patients", ]
#### perf_hf_results: Our main analysis df.
perf_hf_results <- fair_hf_results[fair_hf_results$sg_key == "race", ]


# M1: M1
# total_ahrq_cty_nomissing: M6
county_feat_sets <- c("M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm_demo",
        "M5_and_demo",
        "M6_total_ahrq_cty_DF1_nm")
county_only_feats <- c( "DF1_nm_county_AHRQ_domain1",
        "DF1_nm_county_AHRQ_domain2",
        "DF1_nm_county_AHRQ_domain3",
        "DF1_nm_county_AHRQ_domain4",
        "DF1_nm_county_AHRQ_domain5")
tract_only_feats <- c( "DF1_nm_tract_AHRQ_domain1",
        "DF1_nm_tract_AHRQ_domain2",
        "DF1_nm_tract_AHRQ_domain3",
        "DF1_nm_tract_AHRQ_domain4",
        "DF1_nm_tract_AHRQ_domain5", "M6_total_ahrq_trct_DF1_nm")


model_order <- c("logr", "rf", "xgb")

# ACROSS ALL folds. paired t-test for XGBoost vs. Logr.

curr_feat <- tract_only_feats

# 1.) EO Ratio: eo_ratio
eo_ratio_avg <- fair_hf_results[fair_hf_results$feat %in% curr_feat, ] %>% group_by(feat, model) %>% summarise(mean_eo_ratio = mean(eo_ratio)) %>% mutate(mean_eo_ratio = round(mean_eo_ratio, 3)) %>% mutate(model = factor(model, levels = model_order)) %>%  mutate(
    feat = factor(feat, levels = curr_feat),
    model = factor(model, levels = model_order)
  ) %>% arrange(feat, model)
# Print total df:
eo_ratio_avg
# Print results column for Excel:
cat("EO Ratio: \n")
cat(paste(eo_ratio_avg %>% pull(mean_eo_ratio), collapse = "\n"))


# 2.) FPR Parity: fpr_parity
fpr_parity_avg <- fair_hf_results[fair_hf_results$feat %in% curr_feat, ] %>% group_by(feat, model) %>% summarise(mean_fpr_parity = mean(fpr_parity)) %>% mutate(mean_fpr_parity = round(mean_fpr_parity, 3)) %>% mutate(model = factor(model, levels = model_order)) %>%  mutate(
    feat = factor(feat, levels = curr_feat),
    model = factor(model, levels = model_order)
  ) %>% arrange(feat, model)
# Print total df:
fpr_parity_avg
# Print results column for Excel:
cat("FPR Parity: \n")
cat(paste(fpr_parity_avg %>% pull(mean_fpr_parity), collapse = "\n"))



# 3.) TPR Parity: tpr_parity
tpr_parity_avg <- fair_hf_results[fair_hf_results$feat %in% curr_feat, ] %>% group_by(feat, model) %>% summarise(mean_tpr_parity = mean(tpr_parity)) %>% mutate(mean_tpr_parity = round(mean_tpr_parity, 3)) %>% mutate(model = factor(model, levels = model_order)) %>%  mutate(
    feat = factor(feat, levels = curr_feat),
    model = factor(model, levels = model_order)
  ) %>% arrange(feat, model)
# Print total df:
tpr_parity_avg
# Print results column for Excel:
cat("TPR Parity: \n")
cat(paste(tpr_parity_avg %>% pull(mean_tpr_parity), collapse = "\n"))


# 4.) FNR Parity: fnr_parity
fnr_parity_avg <- fair_hf_results[fair_hf_results$feat %in% curr_feat, ] %>% group_by(feat, model) %>% summarise(mean_fnr_parity = mean(fnr_parity)) %>% mutate(mean_fnr_parity = round(mean_fnr_parity, 3)) %>% mutate(model = factor(model, levels = model_order)) %>%  mutate(
    feat = factor(feat, levels = curr_feat),
    model = factor(model, levels = model_order)
  ) %>% arrange(feat, model)
# Print total df:
fnr_parity_avg
# Print results column for Excel:
cat("FNR Parity: \n")
cat(paste(fnr_parity_avg %>% pull(mean_fnr_parity), collapse = "\n"))

# 5.) Demographic Parity ratio: dp_ratio
dp_ratio_avg <- fair_hf_results[fair_hf_results$feat %in% curr_feat, ] %>% group_by(feat, model) %>% summarise(mean_dp_ratio = mean(dp_ratio)) %>% mutate(mean_dp_ratio = round(mean_dp_ratio, 3)) %>% mutate(model = factor(model, levels = model_order)) %>%  mutate(
    feat = factor(feat, levels = curr_feat),
    model = factor(model, levels = model_order)
  ) %>% arrange(feat, model)
# Print total df:
dp_ratio_avg
# Print results column for Excel:
cat("DP Ratio: \n")
cat(paste(dp_ratio_avg %>% pull(mean_dp_ratio), collapse = "\n"))


```

# COMPARE: performance vs. fairness. 
```{r}
# PREDICTION PERFORMANCE:
pred_hf_results <- full_hf_results <- read.csv("../scripts/final_results_individualfolds_Circ_2024.csv")# "../Circ_HF_results_individualfolds.csv"
pred_hf_results <- pred_hf_results[pred_hf_results$subgroup == "blackorwhite", ] # blackorwhite, black, white

# FAIRNESS: 
# Filter out sg_key == "all_patients". 
total_fair_hf_results <- read.csv("../scripts/individualfolds_Circ_HF_results_fairness_final.csv") #"../v3_fairness_circulation_results.csv"

fair_hf_results <- total_fair_hf_results[total_fair_hf_results$sg_key != "all_patients", ]
fair_hf_results <- fair_hf_results[fair_hf_results$sg_key == "race", ]

# Restrict to XGBoost
pred_hf_results <- pred_hf_results[pred_hf_results$model == "xgb", ] # xgb
fair_hf_results <- fair_hf_results[fair_hf_results$model == "xgb", ] # xgb
```
# PLOT comparison (line plot): AUROC vs. EO ratio.


(ONLY FOR XGBoost)
- Show scatter plot of 
    AUROC vs. EO ratio for:
  *all* sets (county_feat_sets, county_only_feats, tract_only_feats). XGB models.

county_feat_sets <- c("M1", 
        "M2",
        "M3_county",
        "M4",
        "M5",
        "total_ahrq_cty_nomissing")
county_only_feats <- c( "ahrq1_social_cty_nomissing",
        "ahrq2_economic_cty_nomissing",
        "ahrq3_education_cty_nomissing",
        "ahrq4_physicalinfrastructure_cty_nomissing",
        "ahrq5_healthcarequality_cty_nomissing",
        "total_ahrq_cty_nomissing")
tract_only_feats <- c("ahrq1_social_trct_nomissing",
        "ahrq2_economic_trct_nomissing",
        "ahrq3_education_trct_nomissing",
        "ahrq4_physicalinfrastructure_trct_nomissing",
        "ahrq5_healthcarequality_trct_nomissing",
        "total_ahrq_trct_nomissing"
        )

```{r}
# feat_sets <- c("Set 1", "Set 2", "Set 3", "Set 4", "Set 5", "Set 6")

# Pred performance

# M1: M1
# total_ahrq_cty_nomissing: M6

county_feat_sets <- c("M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm_demo",
        "M5_and_demo",
        "M6_total_ahrq_cty_DF1_nm")
county_only_feats <- c("M6_total_ahrq_cty_DF1_nm",
                       "DF1_nm_county_AHRQ_domain1",
        "DF1_nm_county_AHRQ_domain2",
        "DF1_nm_county_AHRQ_domain3",
        "DF1_nm_county_AHRQ_domain4",
        "DF1_nm_county_AHRQ_domain5"
        )
tract_only_feats <- c("M6_total_ahrq_trct_DF1_nm",
                      "DF1_nm_tract_AHRQ_domain1",
        "DF1_nm_tract_AHRQ_domain2",
        "DF1_nm_tract_AHRQ_domain3",
        "DF1_nm_tract_AHRQ_domain4",
        "DF1_nm_tract_AHRQ_domain5"
        )



# Remove duplicate feat:         "total_ahrq_cty_nomissing"
total_feats_used <- c("M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm_demo",
        "M5_and_demo",
        "M6_total_ahrq_cty_DF1_nm",
        
        "DF1_nm_county_AHRQ_domain1",
        "DF1_nm_county_AHRQ_domain2",
        "DF1_nm_county_AHRQ_domain3",
        "DF1_nm_county_AHRQ_domain4",
        "DF1_nm_county_AHRQ_domain5",
        
        "DF1_nm_tract_AHRQ_domain1",
        "DF1_nm_tract_AHRQ_domain2",
        "DF1_nm_tract_AHRQ_domain3",
        "DF1_nm_tract_AHRQ_domain4",
        "DF1_nm_tract_AHRQ_domain5",
        "total_ahrq_trct_nomissing"
        )

long_names <- c("Set 1",
                "Set 2",
                "Set 3",
                "Set 4",
                "Set 5",
                "Set 6",
                
                "County: Domain 1",
                "County: Domain 2",
                "County: Domain 3",
                "County: Domain 4",
                "County: Domain 5",
                
                "Tract: Domain 1",
                "Tract: Domain 2",
                "Tract: Domain 3",
                "Tract: Domain 4",
                "Tract: Domain 5",
                "Tract: All"
                )

long_county_feat_sets <- c("Set 1", 
        "Set 2",
        "Set 3",
        "Set 4",
        "Set 5",
        "Set 6")
long_county_only_feats <- c( "County: Domain 1",
        "County: Domain 2",
        "County: Domain 3",
        "County: Domain 4",
        "County: Domain 5",
        "County: All")
long_tract_only_feats <- c("Tract: Domain 1",
        "Tract: Domain 2",
        "Tract: Domain 3",
        "Tract: Domain 4",
        "Tract: Domain 5",
        "Tract: All"
        )

# F1: using feat 
f1_avg <- pred_hf_results[pred_hf_results$feat %in% total_feats_used, ] %>% group_by(feat) %>% summarise(mean_f1 = mean(f1)) %>% mutate(mean_f1 = round(mean_f1, 3)) 
f1_avg


# AUC: using feat 
auc_avg <- pred_hf_results[pred_hf_results$feat %in% total_feats_used, ]  %>% group_by(feat) %>% summarise(mean_auc = mean(auc)) %>% mutate(mean_auc = round(mean_auc, 3)) 
auc_avg



# Fairness
eo_ratio_avg <- fair_hf_results[fair_hf_results$feat %in% total_feats_used, ]  %>% group_by(feat) %>% summarise(mean_eo_ratio = mean(eo_ratio)) %>% mutate(mean_eo_ratio = round(mean_eo_ratio, 3)) 
eo_ratio_avg


# Temp rename for plotting.
# auc_avg <- f1_avg
# names(auc_avg)[names(auc_avg) == "mean_f1"] <- "mean_auc"
```

# Run tests:
```{r}
library(dplyr)
library(tidyr)
library(knitr)


fairness_feats_curr <- tract_only_feats # tract_only_feats, county_feat_sets

# Function to perform Wilcoxon test safely
perform_wilcoxon_test <- function(baseline_results, current_results, metric_col) {
  # Ensure we're working with the correct columns
  baseline_values <- baseline_results[[metric_col]]
  current_values <- current_results[[metric_col]]
  
  # Perform Wilcoxon test if possible
  if (length(baseline_values) == length(current_values)) {
    wilcox_test <- wilcox.test(
      baseline_values, 
      current_values, 
      paired = TRUE, 
      alternative = "two.sided"
    )
    return(wilcox_test$p.value)
  } else {
    warning(paste("Unequal lengths for", metric_col, 
                  "Baseline:", length(baseline_values), 
                  "Current:", length(current_values)))
    return(NA)
  }
}

# Prepare a dataframe to store Wilcoxon test results
wilcoxon_results <- data.frame(
  metric = character(),
  feature_set = character(),
  p_value = numeric(),
  stringsAsFactors = FALSE
)

# List of feature sets to compare
feature_sets <- county_only_feats[-1]  # Exclude baseline
baseline_feat <- "M6_total_ahrq_trct_DF1_nm" # M6_total_ahrq_trct_DF1_nm, M6_total_ahrq_cty_DF1_nm, M1

# Metrics to test
metrics_to_test <- c("auc", "f1", "eo_ratio")

# Determine which results dataframe to use for each metric
results_dfs <- list(
  auc = pred_hf_results,
  f1 = pred_hf_results,
  eo_ratio = fair_hf_results
)

# Perform Wilcoxon signed-rank test for each metric
for (metric_name in metrics_to_test) {
  # Get the full results dataframe for this metric
  full_results <- results_dfs[[metric_name]]
  
  # Baseline data
  baseline_data <- full_results %>% 
    filter(feat == baseline_feat)
  
  # Compare each feature set to baseline
  for (current_feat in feature_sets) {
    # Current feature set data
    current_data <- full_results %>% 
      filter(feat == current_feat)
    
    # Perform Wilcoxon test
    p_value <- perform_wilcoxon_test(baseline_data, current_data, metric_name)
    
    # Store results
    wilcoxon_results <- rbind(wilcoxon_results, data.frame(
      metric = metric_name,
      feature_set = current_feat,
      p_value = p_value
    ))
  }
}

# Prepare the table of Wilcoxon test results
wilcoxon_table <- wilcoxon_results %>%
  pivot_wider(
    names_from = metric,
    values_from = p_value,
    names_glue = "{metric}_p"
  ) %>%
  mutate(
    across(where(is.numeric), ~ sprintf("%.3f", .)),
    feature_set = fairness_feats_curr[-1]  # Use long names for readability
  )

# Print the Wilcoxon signed-rank test results
kable(
  wilcoxon_table,
  col.names = c(
    "Feature Set", 
    "AUC p-value", 
    "F1 p-value", 
    "EO Ratio p-value"
  ),
  caption = "Wilcoxon Signed-Rank Test P-values Comparing Feature Sets to Baseline"
)
```






# Change names.
```{r}

name_mapping <- setNames(long_names, total_feats_used)

# Replace values in auc_avg$feat
auc_avg$feat <- name_mapping[match(auc_avg$feat, total_feats_used)]
f1_avg$feat <- name_mapping[match(f1_avg$feat, total_feats_used)]
eo_ratio_avg$feat <- name_mapping[match(eo_ratio_avg$feat, total_feats_used)]

```


```{r}

# Calculate the balance score (product of x and y values)
balance_scores <- auc_avg$mean_auc * eo_ratio_avg$mean_eo_ratio

# Find the index of the maximum balance score
best_index <- which.max(balance_scores)

# Get the best feature set and its corresponding values
best_feat_set <- auc_avg$feat[best_index] # Order matches eo_ratio_avg$feat
best_auroc <- auc_avg$mean_auc[best_index]
best_eo <- eo_ratio_avg$mean_eo_ratio[best_index]

# Print the best feature set and its values
cat("Best feature set:", best_feat_set, "\n")
cat("AUROC:", best_auroc, "\n")
cat("EO ratio:", best_eo, "\n")

############################
# Plot the AUROC vs EO ratio
# Define the three color palettes
colors_county_feat_sets <- brewer.pal(length(long_county_feat_sets), "Set1")
colors_county_only_feats <- brewer.pal(length(long_county_only_feats), "Set2")
colors_tract_only_feats <- brewer.pal(length(long_tract_only_feats), "Set3")

# Create a named vector for colors
color_mapping <- c(setNames(colors_county_feat_sets, long_county_feat_sets),
                   setNames(colors_county_only_feats, long_county_only_feats),
                   setNames(colors_tract_only_feats, long_tract_only_feats))

# Assign colors to each point based on the feature set
point_colors <- color_mapping[auc_avg$feat]

# Adjust margins to make space for the legend
par(mar = c(5, 4, 4, 15) + 0.1)

# Plot
plot(eo_ratio_avg$mean_eo_ratio, auc_avg$mean_auc, type = "p", col = point_colors, pch = 19,
     xlab = "EO Ratio", ylab = "AUROC", xlim = c(0, 1), ylim = c(0, 1))

# Highlight the best point
points(best_eo, best_auroc, col = "red", pch = 19, cex = 1.5)

# Create a combined legend
legend_labels <- c(long_county_feat_sets, long_county_only_feats, long_tract_only_feats)
legend_colors <- color_mapping[legend_labels]

# Add legend outside the plot area
legend("topright", inset = c(-0.5, 0), legend = legend_labels, col = legend_colors, pch = 19, xpd = TRUE)




```

```{r}
# Define the three color palettes
colors_county_feat_sets <- brewer.pal(length(long_county_feat_sets), "Set1")
colors_county_only_feats <- brewer.pal(length(long_county_only_feats), "Set2")
colors_tract_only_feats <- brewer.pal(length(long_tract_only_feats), "Set3")

# Define the three shapes
shapes_county_feat_sets <- rep(19, length(long_county_feat_sets))  # circles
shapes_county_only_feats <- rep(17, length(long_county_only_feats))  # triangles
shapes_tract_only_feats <- rep(15, length(long_tract_only_feats))  # squares


# Find the point with the highest AUROC and EO ratio
max_auroc_index <- which.max(auc_avg$mean_auc)
max_eo_index <- which.max(eo_ratio_avg$mean_eo_ratio)

max_auroc <- auc_avg$mean_auc[max_auroc_index]
max_eo <- eo_ratio_avg$mean_eo_ratio[max_eo_index]

# Print the best feature set and its values
cat("Best feature set:", best_feat_set, "\n")
cat("AUROC:", best_auroc, "\n")
cat("EO ratio:", best_eo, "\n")



# Create named vectors for colors and shapes
color_mapping <- c(setNames(colors_county_feat_sets, long_county_feat_sets),
                   setNames(colors_county_only_feats, long_county_only_feats),
                   setNames(colors_tract_only_feats, long_tract_only_feats))

shape_mapping <- c(setNames(shapes_county_feat_sets, long_county_feat_sets),
                   setNames(shapes_county_only_feats, long_county_only_feats),
                   setNames(shapes_tract_only_feats, long_tract_only_feats))

# Assign colors and shapes to each point based on the feature set
point_colors <- color_mapping[auc_avg$feat]
point_shapes <- shape_mapping[auc_avg$feat]

# Adjust margins to make space for the legend
par(mar = c(5, 4, 4, 15) + 0.1)

# Plot
plot(eo_ratio_avg$mean_eo_ratio, auc_avg$mean_auc, type = "p", col = point_colors, pch = point_shapes,
     xlab = "EO Ratio", ylab = "AUROC", xlim = c(0, 1), ylim = c(0, 1))

# Highlight the best point
points(best_eo, best_auroc, col = "red", pch = 19, cex = 1.5)

# Add dashed black lines at the points with the highest f1 and EO ratio
abline(h = max_auroc, col = "black", lty = 2)  # Horizontal dashed line for max f1
abline(v = max_eo, col = "black", lty = 2)     # Vertical dashed line for max EO ratio


# Create a combined legend
legend_labels <- c(long_county_feat_sets, long_county_only_feats, long_tract_only_feats)
legend_colors <- color_mapping[legend_labels]
legend_shapes <- shape_mapping[legend_labels]

# Add legend outside the plot area
legend("topright", inset = c(-0.5, 0), legend = legend_labels, col = legend_colors, pch = legend_shapes, xpd = TRUE)


```

