---
title: "Untitled"
author: "Chase Fensore"
date: "2024-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Bringing in AHRQ data
# MERGE IN:
- County AHRQ data
- Census tract AHRQ data
- 
```{r}
library(caret)
library(data.table)
library(dplyr)
library(sets)
library(scales)
library(tidyr)
library(stringr)

# State plotting
library(usmap)
library(sf)
library(rio)  
library(sp) # Dependency
library(maps)

# Set wd when not in RStudio
source('data_cleaners.R')
```

# Load AHRQ data: county, census tract, zip
```{r}
library(stringi)


# County AHRQ
# cty_ahrq_df_2009 <- read.csv("./COUNTY_AHRQ/SDOH_2010_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2010 <- read.csv("./COUNTY_AHRQ/SDOH_2010_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2011 <- read.csv("./COUNTY_AHRQ/SDOH_2011_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2012 <- read.csv("./COUNTY_AHRQ/SDOH_2012_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2013 <- read.csv("./COUNTY_AHRQ/SDOH_2013_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2014 <- read.csv("./COUNTY_AHRQ/SDOH_2014_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2015 <- read.csv("./COUNTY_AHRQ/SDOH_2015_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2016 <- read.csv("./COUNTY_AHRQ/SDOH_2016_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2017 <- read.csv("./COUNTY_AHRQ/SDOH_2017_COUNTY_1_0.csv") # header=FALSE
cty_ahrq_df_2018 <- read.csv("./COUNTY_AHRQ/SDOH_2018_COUNTY_1_1.csv") # header=FALSE
# cty_ahrq_df_2019 <- read.csv("./COUNTY_AHRQ/SDOH_2019_COUNTY_1_0.csv") # header=FALSE
# cty_ahrq_df_2020 <- read.csv("./COUNTY_AHRQ/SDOH_2020_COUNTY_1_0.csv") # header=FALSE

# Use fill to ensured consistent number of columns, merge dfs from 2009-2020 (via mean).
# Merge dfs on the 6 items below in list.
# cty_ahrq_df_2009, 
cty_ahrq_total <- rbindlist(list(cty_ahrq_df_2010, cty_ahrq_df_2011, cty_ahrq_df_2012, cty_ahrq_df_2013,
                             cty_ahrq_df_2014, cty_ahrq_df_2015, cty_ahrq_df_2016, cty_ahrq_df_2017, cty_ahrq_df_2018), fill=TRUE)


# Save names WITHOUT suffix first.
cty_features_nosuffix <- colnames(cty_ahrq_total)
# Rename cty_ahrq_total FEATURE names: append "_countylevel"

# Columns to exclude from renaming
exclude_columns_cty <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "YEAR")

cty_ahrq_total <- cty_ahrq_total %>%
  rename_at(vars(-one_of(exclude_columns_cty)), ~paste0(., "_countylevel"))

# Import AHRQ data
# tr_ahrq_df_2009 <- read.csv("./TRACT_AHRQ/sdoh_2009_tract_1_0.csv")
tr_ahrq_df_2010 <- read.csv("./TRACT_AHRQ/sdoh_2010_tract_1_0.csv") # header=FALSE
tr_ahrq_df_2011 <- read.csv("./TRACT_AHRQ/sdoh_2011_tract_1_0.csv")
tr_ahrq_df_2012 <- read.csv("./TRACT_AHRQ/sdoh_2012_tract_1_0.csv")
tr_ahrq_df_2013 <- read.csv("./TRACT_AHRQ/sdoh_2013_tract_1_0.csv")
tr_ahrq_df_2014 <- read.csv("./TRACT_AHRQ/sdoh_2014_tract_1_0.csv")
tr_ahrq_df_2015 <- read.csv("./TRACT_AHRQ/sdoh_2015_tract_1_0.csv")
tr_ahrq_df_2016 <- read.csv("./TRACT_AHRQ/sdoh_2016_tract_1_0.csv")
tr_ahrq_df_2017 <- read.csv("./TRACT_AHRQ/sdoh_2017_tract_1_0.csv")
tr_ahrq_df_2018 <- read.csv("./TRACT_AHRQ/sdoh_2018_tract_1_0.csv")
# tr_ahrq_df_2019 <- read.csv("./TRACT_AHRQ/sdoh_2019_tract_1_0.csv")
# tr_ahrq_df_2020 <- read.csv("./TRACT_AHRQ/sdoh_2020_tract_1_0.csv")


# tr_ahrq_df_2009, 
tr_ahrq_total <- rbindlist(list(tr_ahrq_df_2010, tr_ahrq_df_2011, tr_ahrq_df_2012, tr_ahrq_df_2013,
                             tr_ahrq_df_2014, tr_ahrq_df_2015, tr_ahrq_df_2016, tr_ahrq_df_2017, tr_ahrq_df_2018), fill=TRUE)


# Save names WITHOUT suffix first.
trct_features_nosuffix <- colnames(tr_ahrq_total)


# Rename tr_ahrq_total FEATURE names: append "_census_tractlevel"
# Columns to exclude from renaming
exclude_columns_trct <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "TRACTFIPS", "YEAR") # "TRACTFIPS_padded",

tr_ahrq_total <- tr_ahrq_total %>%
  rename_at(vars(-one_of(exclude_columns_trct)), ~paste0(., "_census_tractlevel"))



# SAVE VAR NAMES: 
# COUNTY:
cty_features <- colnames(cty_ahrq_total)
trct_features <- colnames(tr_ahrq_total)


all_gran_feats_metadata <- readxl::read_excel("2010-18-all-granularities-AHRQ-dict.xlsx") # has 2010-2018 feats, metadata.

# -- todo: merge all AHRQ features w/ their metadata (from cty_features_metadata ) 
# length(set_intersection(cty_features, cty_feats_metadata$`Variable Name`))

########### COUNTY LEVEL: #################
# COUNTY: 1267 county features (2010-2018)
length(set_intersection(cty_features_nosuffix, all_gran_feats_metadata$`Variable Name`))
# Only select feats metadata which are present in "cty_features."
cty_features_and_metadata <- all_gran_feats_metadata[all_gran_feats_metadata$`Variable Name` %in% cty_features_nosuffix, ]

# Lower-case domain names.
cty_features_and_metadata <- cty_features_and_metadata %>%
  mutate(Domain = stri_trans_tolower(stri_trans_general(Domain, "Latin-ASCII")))

# Columns to exclude from renaming
exclude_columns_cty <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "YEAR")

cty_features_and_metadata <- cty_features_and_metadata %>%
  mutate(`Variable Name` = case_when(
    `Variable Name` %in% exclude_columns_cty ~ `Variable Name`,
    TRUE ~ paste0(`Variable Name`, "_countylevel")
  ))


# Print the modified dataframe
cat("COUNTY Num variables per Domain: ", "\n")
table(cty_features_and_metadata$Domain)

# print(cty_features_and_metadata)


########### TRACT LEVEL: #################
length(set_intersection(trct_features_nosuffix, all_gran_feats_metadata$`Variable Name`))
# TRACT: 519 county features (2010-2018)
# Only select feats metadata which are present in "trct_features_nosuffix"
trct_features_and_metadata <- all_gran_feats_metadata[all_gran_feats_metadata$`Variable Name` %in% trct_features_nosuffix, ]

# Lower-case domain names.
trct_features_and_metadata <- trct_features_and_metadata %>%
  mutate(Domain = stri_trans_tolower(stri_trans_general(Domain, "Latin-ASCII")))

# Append _census_tractlevel to all feats in this METADATA:

exclude_columns_trct <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "TRACTFIPS", "YEAR") # "TRACTFIPS_padded"

trct_features_and_metadata <- trct_features_and_metadata %>%
  mutate(`Variable Name` = case_when(
    `Variable Name` %in% exclude_columns_trct ~ `Variable Name`,
    TRUE ~ paste0(`Variable Name`, "_census_tractlevel")
  ))


# Print the modified dataframe
# print(trct_features_and_metadata)

cat("TRACT Num variables per Domain: ", "\n")
table(trct_features_and_metadata$Domain)


cat("OVERALL COUNTY (not geo or identifiers): 1240 \n")
cat("OVERALL TRACT (not geo or identifiers): 506 \n")

```


# Over 2010-2018 at county level: how many distinct datasets (Data Source) does AHRQ contain?
```{r}

# First, ignore geography variables.
all_cty_unique_databases <- subset(subset(cty_features_and_metadata, Domain != '6. geography'),  Domain != 'identifier')
nrow(cty_features_and_metadata)
nrow(all_cty_unique_databases)

# 1. 2010-2018 county: save unique 'Data Source' (excluding geography, identifier)
print("Unique 2010-2018 County data sources, excluding geography/identifier: ")
unique(all_cty_unique_databases$'Data Source')

# 2. 2010-2018 county: save variable names within each 'Data Souce'
all_cty_unique_databases



############
# First, ignore geography variables.
all_trct_unique_databases <- subset(subset(trct_features_and_metadata, Domain != '6. geography'), Domain != 'identifier')
nrow(trct_features_and_metadata)
nrow(all_trct_unique_databases) 

# 1. 2010-2018 tract: save unique 'Data Source' (excluding geography, identifier)
print("Unique 2010-2018 Tract data sources, excluding geography/identifier:")
unique(all_trct_unique_databases$'Data Source')


# 2. 2010-2018 tract: save variable names within each 'Data Souce'
# all_trct_unique_databases

```


# Save AHRQ census tract variables for LLM annotation:
```{r}
llm_tract_AHRQ_prompts <- trct_features_and_metadata

# 1.) Rename columns: 
# Domain, Variable Name, Variable Label, Data Source
# domain, variable_name, variable_label, data_source
llm_tract_AHRQ_prompts <- llm_tract_AHRQ_prompts %>%
  rename(
    domain = Domain,
    SDOH_topic = Topic,
    variable_name = `Variable Name`,  # Enclose the column name in backticks because it contains spaces
    variable_label = `Variable Label`, # Enclose the column name in backticks because it contains spaces
    data_source = `Data Source`       # Enclose the column name in backticks because it contains spaces
  )


# 2.) Don't include vars: drop these rows.
# exclude_columns_trct <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "TRACTFIPS", "YEAR") # "TRACTFIPS_padded"
exclude_columns_trct <- c("6. geography", "identifier")
llm_tract_AHRQ_prompts <- llm_tract_AHRQ_prompts %>%
  filter(!(domain %in% exclude_columns_trct))

# TODO: We *could* insert full names in data_source:

write.csv(llm_tract_AHRQ_prompts ,"AHRQ_tract_varnames_2010-2018.csv", row.names = FALSE )


nrow(llm_tract_AHRQ_prompts)
cat("Excluding geography/identifier variables: \n")
table(llm_tract_AHRQ_prompts$domain)
```




# Read in patient data (includes maximum possible number of geocoded patients)
```{r}

patient_cohort <- read.csv(file="./ML4H_Circ_Nonmissing_Cohort.csv") # Fully geocoded from patient_inclusion_Circ_2024.Rmd
cat("Number patients fully geocoded: ", nrow(patient_cohort), "\n")


# Pad censustract column to 11 characters with leading zeroes
patient_cohort$censustract <- str_pad(patient_cohort$censustract, width = 11, pad = "0") # Tract fips.
patient_cohort$fips <- str_pad(patient_cohort$fips, width = 5, pad = "0") # County fips.

```


# For "patient_cohort" df with missing ADI & SDI values (since newly geocoded via addresses), fill in ADI & SDI Columns: 
- "ADI_national_rank"
- "sdi"


CHALLENGE -- all of ADI, & some of SDI are not able to match with patient_cohort DF via "censustract"!


```{r}
# Read in ADI:

adi_tract <- read.csv("./adi-download-2020-tract/US_2020_ADI_Census_Block_Group_v4_0_1.csv")
# Convert to character and ensure all values have exactly 11 digits
adi_tract$FIPS <- ifelse(
  nchar(as.character(adi_tract$FIPS)) == 12,
  substr(as.character(adi_tract$FIPS), 1, 11),
  as.character(adi_tract$FIPS)
)
# Pad censustract column to 11 characters with leading zeroes
adi_tract$FIPS <- str_pad(adi_tract$FIPS, width = 11, pad = "0", side = "left")

# Read in SDI: 
sdi_censustract <- read.csv("./sdi-download-2019-tract/rgcsdi-2015-2019-censustract.csv")


sdi_censustract$CENSUSTRACT_FIPS <- as.character(sdi_censustract$CENSUSTRACT_FIPS) # Cast to <chr> for future join.
sdi_censustract$CENSUSTRACT_FIPS <- str_pad(sdi_censustract$CENSUSTRACT_FIPS, width = 11, pad = "0", side = "left")



cat("Before merging: \n")
# ###########
# ADI missing: 
# table(patient_cohort$ADI_national_rank)
sum(is.na(patient_cohort$ADI_national_rank))
# SDI missing:
# table(patient_cohort$sdi)
sum(is.na(patient_cohort$sdi))


# MERGE: 
# ADI, SDI values are shown in "patient_cohort" as NA.
# Join adi_tract data to patient_cohort on matching census tract FIPS codes
# 1.) ADI:
#patient_cohort <- patient_cohort %>%
#  left_join(adi_tract %>% select(FIPS, ADI_NATRANK), by = c("censustract" = "FIPS")) %>%
#  mutate(
#    ADI_national_rank = ifelse(is.na(ADI_national_rank), ADI_NATRANK, ADI_national_rank)
#  ) %>%
#  select(-ADI_NATRANK) # Remove the extra column after filling in

# First, sample a random ADI_NATRANK value per tract in adi_tract
adi_tract_random <- adi_tract %>%
  group_by(FIPS) %>%
  sample_n(1) %>%
  ungroup() %>%
  select(FIPS, ADI_NATRANK)

# Now perform the left join as before, only filling NA values in patient_cohort$ADI_national_rank
patient_cohort <- patient_cohort %>%
  left_join(adi_tract_random, by = c("censustract" = "FIPS")) %>%
  mutate(
    ADI_national_rank = ifelse(is.na(ADI_national_rank), ADI_NATRANK, ADI_national_rank)
  ) %>%
  select(-ADI_NATRANK) # Remove the extra column after filling in



# 2.) SDI:
patient_cohort <- patient_cohort %>%
  left_join(sdi_censustract %>% select(CENSUSTRACT_FIPS, SDI_score), by = c("censustract" = "CENSUSTRACT_FIPS")) %>%
  mutate(
    sdi = ifelse(is.na(sdi), SDI_score, sdi)
  ) %>%
  select(-SDI_score) # Remove the extra column after filling in

# SDI: Finally, merge other columns in:
# List of column names to merge into patient_cohort
# "sdi",
columns_to_merge <- c(
  "PovertyLT100_FPL_score", "Single_Parent_Fam_score", "Education_LT12years_score", 
  "HHNo_Vehicle_score", "HHRenter_Occupied_score", "HHCrowding_score", 
  "Nonemployed_score", "pct_Poverty_LT100", "pct_Single_Parent_Fam", 
  "pct_Education_LT12years", "pct_NonEmployed", "pctHH_No_Vehicle", 
  "pctHH_Renter_Occupied", "pctHH_Crowding"
)

# Merge columns into patient_cohort
patient_cohort <- patient_cohort %>%
  left_join(
    sdi_censustract %>% select(CENSUSTRACT_FIPS, all_of(columns_to_merge)), 
    by = c("censustract" = "CENSUSTRACT_FIPS")
  )

nrow(patient_cohort)

```


```{r}
# Check missing rows...
# patient_cohort[is.na(patient_cohort$ADI_national_rank), ]
# sum(table(sdi_censustract$SDI_score))
# nrow(sdi_censustract)


#table(sdi_censustract$SDI_score, useNA = "ifany") # 646 MISSING VALUES shown as: "<NA>"
#table(adi_tract$ADI_NATRANK, useNA = "ifany") # MISSING VALUES shown as: "GQ", "GQ-PH", "PH", "QDI". Rates: c(2803, 778, 2649, 215)


# For ADI, SDI: check how many censustract values do not have counterparts to match on within adi_tract df (has FIPS column).
unmatched_tract_ADI <- setdiff(patient_cohort$censustract, adi_tract$FIPS)
unmatched_tract_SDI <- setdiff(patient_cohort$censustract, sdi_censustract$CENSUSTRACT_FIPS)

length(unmatched_tract_ADI) # 1356
length(unmatched_tract_SDI) # 858


# Check ADI for these census tracts...

# Before merging: 
#[1] 6586
#[1] 4778
#After merging: 
#[1] 1338
#[1] 2155


# Breaking down missing values of deprivation indices...
cat("After merging: \n")
sum(is.na(patient_cohort$ADI_national_rank))
sum(is.na(patient_cohort$sdi))

# ADI: 
# 'ADI_national_rank' has 6586 missing before merging with ADI data, 1338 after. Among the 1338, 1356 do not have a matching census tract between patient_cohort & adi_tract.
# --> ___ don't match ??
# table(patient_cohort$ADI_national_rank, useNA = "ifany") # INSERT 'NA' for these values: 'GQ' 'GQ-PH'    'PH'   'QDI' 

# SDI: 
# 'sdi' has 4778 missing before merging with SDI data, 2155 after. Among the 2155, 858 do not have a matching census tract between patient_cohort & sdi_censustract.
# --> the other 1297: ____ unreported SDI from small areas, or poorly collected data. So there's no SDI from these. 

# Find rows where NO value for patient_cohort$sdi, and patient_cohort$censustract is NOT contained within the list unmatched_tract_SDI.
# Filter rows based on the conditions for `sdi` and `censustract`


#table(patient_cohort$sdi, useNA = "ifany")
table(patient_cohort$sdi.x, useNA = "ifany") # Has integers. 2155 <NA>
table(patient_cohort$sdi.y, useNA = "ifany") # Has decimals, +/-. How many <NA> ?

nrow(patient_cohort)
```





# We already have ADI, SDI in the dataset. But we need to merge in SVI.
```{r}
# SVI Year: 2014. SVI Geo level: census tract. All states in US included. 
SVI_2014_US <- read.csv("./TRACT_AHRQ/SVI_2014_US.csv") 

# Cast FIPS to 11 digit double. (before, some are 10, some are 11).
SVI_2014_US$FIPS <- sprintf("%011s", SVI_2014_US$FIPS)


keep_svi <- c(
              "FIPS",
              "SPL_THEMES",
              "EPL_POV", "EPL_UNEMP", "EPL_PCI", "EPL_NOHSDP",
              "EPL_AGE65", "EPL_AGE17", "EPL_DISABL", "EPL_SNGPNT",
              "EPL_MINRTY", "EPL_LIMENG", "EPL_MUNIT", "EPL_MOBILE",
              "EPL_CROWD", "EPL_NOVEH", "EPL_GROUPQ",
              "RPL_THEME1",
              "RPL_THEME2",
              "RPL_THEME3",
              "RPL_THEME4",
              "EPL_POV", "EPL_UNEMP", "EPL_PCI", "EPL_NOHSDP",
              "EPL_AGE65", "EPL_AGE17", "EPL_DISABL", "EPL_SNGPNT",
              "EPL_MINRTY", "EPL_LIMENG", "EPL_MUNIT", "EPL_MOBILE",
              "EPL_CROWD", "EPL_NOVEH", "EPL_GROUPQ",
              "RPL_THEMES",
              "F_TOTAL",
              "E_HU", "E_HH", "E_POV", "E_UNEMP", "E_PCI", 
                   "E_NOHSDP", "E_AGE65", "E_AGE17", "E_DISABL", "E_SNGPNT", 
                   "E_MINRTY", "E_LIMENG", "E_MUNIT", "E_MOBILE", "E_CROWD", 
                   "E_NOVEH", "E_GROUPQ",
              "EP_POV", "EP_UNEMP", "EP_PCI", "EP_NOHSDP", 
                            "EP_AGE65", "EP_AGE17", "EP_DISABL", "EP_SNGPNT", 
                            "EP_MINRTY", "EP_LIMENG", "EP_MUNIT", "EP_MOBILE", 
                            "EP_CROWD", "EP_NOVEH", "EP_GROUPQ")

# Subset only necessary SVI columns + FIPS (tract fips). Don't need county fips.
SVI_2014_US <- SVI_2014_US[, keep_svi]


head(SVI_2014_US)

# Merge SVI_2014_US:


# Finally, merge in SVI 2014 data:
# use cty code: ahrq_baseline ++ SVI_2014_US (by 'censustract', 'FIPS')
# Old approach: patient_cohort <- merge(patient_cohort, SVI_2014_US  , by.x="censustract", by.y="FIPS" )
# Make sure to keep all rows in patient_cohort.
patient_cohort <- patient_cohort %>%
  left_join(SVI_2014_US, by = c("censustract" = "FIPS"))
nrow(patient_cohort)

# SVI has 2155 missing:
sum(is.na(patient_cohort$SPL_THEMES)) # 2155 missing.

```
# Contains full data for AHRQ 2010-2018. Able to merge using these DFs. 
```{r}

# These have AHRQ variables as the COLUMNS.
# head(cty_ahrq_total) # May not show all columns & rows (too many to view here)
# head(tr_ahrq_total)
# These have AHRQ variables as the ROWS.
# cty_features_and_metadata
# trct_features_and_metadata

# PREPARE AHRQ spatial codes: 
# Prepare AHRQ join keys (county, tract datasets):
# cty_ahrq_total$COUNTYFIPS # 4-5 digit int (e.g., 4 digit min: 1001, 5 digit max: 78030)
# Coerce to 5 digit *chr*: 
cty_ahrq_total$COUNTYFIPS_padded <- sprintf("%05s", cty_ahrq_total$COUNTYFIPS)
# TRACT: state + county + tract
# tr_ahrq_total$TRACTFIPS # 10-11 digit *double* (e.g., 10 digit min: 1001020100, 11 digit max: 78030990000). Do NOT coerce to int, numbers will overflow.
# Coerce to 11 digit *chr*:
tr_ahrq_total$TRACTFIPS_padded <- sprintf("%011s", tr_ahrq_total$TRACTFIPS)

```


# 1.) Merge (cty_ahrq_total, tr_ahrq_total) <<->> patient_cohort
```{r}
# MERGE: both on subject's 'baselineyear' (admission year)
# Bring in County AHRQ:

# OLD: patient_cohort <- merge(patient_cohort, cty_ahrq_total, by.x=c("fips", "baselineyear"), by.y=c("COUNTYFIPS_padded", "YEAR"))
# Make sure to keep all rows in patient_cohort.
patient_cohort <- patient_cohort %>%
  left_join(cty_ahrq_total, by = c("fips" = "COUNTYFIPS_padded", "baselineyear" = "YEAR")) %>%
  mutate(across(ends_with(".x"), ~ coalesce(.x, get(sub(".x$", ".y", cur_column()))))) %>%
  rename_with(~ sub("\\.x$", "", .), ends_with(".x")) %>%
  select(-ends_with(".y"))


# Tract AHRQ baseline: use later for patient-level cenus tract missingness calculations. 
# OLD: patient_cohort <- merge(patient_cohort, tr_ahrq_total, by.x=c("censustract", "baselineyear"), by.y=c("TRACTFIPS_padded", "YEAR"))
patient_cohort <- patient_cohort %>%
  left_join(tr_ahrq_total, by = c("censustract" = "TRACTFIPS_padded", "baselineyear" = "YEAR")) %>%
  mutate(across(ends_with(".x"), ~ coalesce(.x, get(sub(".x$", ".y", cur_column()))))) %>%
  rename_with(~ sub("\\.x$", "", .), ends_with(".x")) %>%
  select(-ends_with(".y"))

head(patient_cohort) # Patient data + ADI + SDI + SVI + AHRQ matched at county & tract. AHRQ naming scheme: _countylevel, _census_tractlevel
nrow(patient_cohort)
# Patient cohort: 35964

```


# Final steps: only for Circ.

# Before feeding to classifiers, drop potential PHI from ahrq_baseline
```{r}

drop_vars <- c("Lon__Lat", "cxy_address", "cxy_matched_address", "cxy_tiger_line_id", "cxy_tract_id", "cxy_block_id")
ncol(patient_cohort)

# Remove drop_vars
patient_cohort <- patient_cohort[, !(names(patient_cohort) %in% drop_vars)]

ncol(patient_cohort)


# CREATE DEMOGRAPHIC VARS: 

patient_cohort <- patient_cohort %>%
  mutate(baselineyear = as.numeric(baselineyear), na.rm=FALSE)

# Then merge
# ahrq_AQI_baseline_final <- merge(ahrq_baseline, original_df, by.x=c("COUNTYFIPS", "YEAR"), by.y=c("COUNTY_FIPS", "YEAR"))
#sum(is.na(ahrq_baseline$ACS_PCT_PUB_COMMT_59MIN)) # Should be 0 for all columns.


############ Prepare subgroup flags:
# Subgroups to eval on: 
# 1. Race
# white <int>: 0/1
# black <int>: 0/1
# other <int>: 0/1
patient_cohort <- patient_cohort %>%
  mutate(blackorwhite = if_else(black == 1 | white == 1, 1, 0))

# 2. Sex (or binary gender)
# female <int>: 0/1
# ADD this field: male <int>: 0/1
patient_cohort <- patient_cohort %>%
  mutate(male = if_else(female == 0, 1, 0))

# 3. The 6 subgroup combinations: 
# c(white_male, black_male, other_male, white_female, black_female, other_female)

# Note that all 6 are mutually exclusive here.
patient_cohort <- patient_cohort %>%
  mutate(white_male = if_else(white == 1 & female == 0, 1, 0))
patient_cohort <- patient_cohort %>%
  mutate(black_male = if_else((black == 1 & female == 0), 1, 0))
patient_cohort <- patient_cohort %>%
  mutate(other_male = if_else(other == 1 & female == 0, 1, 0))
patient_cohort <- patient_cohort %>%
  mutate(white_female = if_else(white == 1 & female == 1, 1, 0))
patient_cohort <- patient_cohort %>%
  mutate(black_female = if_else(black == 1 & female == 1, 1, 0))
patient_cohort <- patient_cohort %>%
  mutate(other_female = if_else(other == 1 & female == 1, 1, 0))

#######################
# Hypertension
patient_cohort <- patient_cohort %>%
  mutate(no_htn = if_else(htn == 0, 1, 0))

#######################
# Coronary Artery Disease
patient_cohort <- patient_cohort %>%
  mutate(no_cad = if_else(cad == 0, 1, 0))

#######################
# Chronic Kidney Disease
patient_cohort <- patient_cohort %>%
  mutate(no_ckd = if_else(ckd == 0, 1, 0))




```

# Re-usable function for each of the 4 DF below:
```{r}


prepare_inputDF <- function(total_df) {
    
  #######################
  # SUBGROUP FOR AGEs: 

  quarts <- quantile(total_df$age_current, probs = seq(0, 1, 0.25))
  upper_Q1_age <- quarts[2]
  upper_Q2_age <- quarts[3]
  upper_Q3_age <- quarts[4]
  upper_Q4_age <- quarts[5]
  
  total_df <- total_df %>%
    mutate(age_Q1 = if_else(age_current <= upper_Q1_age, 1, 0))
  
  total_df <- total_df %>%
    mutate(age_Q2 = if_else(age_current <= upper_Q2_age & age_current > upper_Q1_age, 1, 0))
  
  total_df <- total_df %>%
    mutate(age_Q3 = if_else(age_current <= upper_Q3_age & age_current > upper_Q2_age, 1, 0))
  
  total_df <- total_df %>%
    mutate(age_Q4 = if_else(age_current <= upper_Q4_age & age_current > upper_Q3_age, 1, 0))
  
  # Charlson Comorbidity Index
  # Theoretical Max CCI: 24 (updated version) or 29 (older version).
  # Our dataset: Max CCI = 22.
  CCI_quarts <- quantile(total_df$cci, probs = seq(0, 1, 0.25))
  
  
  total_df <- total_df %>%
    mutate(CCI_0 = if_else(cci == 0, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_1 = if_else(cci == 1, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_2 = if_else(cci == 2, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_3 = if_else(cci == 3, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_4 = if_else(cci == 4, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_5 = if_else(cci == 5, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_6 = if_else(cci == 6, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_7up = if_else(cci >= 7, 1, 0))
  
  # 2.) CCI quartiles (second way to evaluate CCI)
  upper_Q1_CCI <- CCI_quarts[2]
  upper_Q2_CCI <- CCI_quarts[3]
  upper_Q3_CCI <- CCI_quarts[4]
  upper_Q4_CCI <- CCI_quarts[5]
  
  
  total_df <- total_df %>%
    mutate(CCI_Q1 = if_else(cci <= upper_Q1_CCI, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_Q2 = if_else(cci <= upper_Q2_CCI & cci > upper_Q1_CCI, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_Q3 = if_else(cci <= upper_Q3_CCI & cci > upper_Q2_CCI, 1, 0))
  
  total_df <- total_df %>%
    mutate(CCI_Q4 = if_else(cci > upper_Q3_CCI, 1, 0))
  
  # Create 10 folds.
  endpoints <- c("readmit30bin", "death30bin", "composite30")
  total_df <- create_folds(as.data.frame(total_df), endpoints) # NOTE: For nested k-fold CV.
  
  return(total_df)
}



# FUNCTION: Calculate missing rates for DF 1, DF 3 (where cols are not removed, we retain cols even if they have patients with misisng vals).
# Define the function to calculate missing rates and save to CSV
calculate_missing_rates <- function(input_df, cty_features_and_metadata, trct_features_and_metadata, output_file) {
  
  # Define lists of variable names and domains from cty_features_and_metadata and trct_features_and_metadata
  cty_var_names <- cty_features_and_metadata$`Variable Name`
  cty_domains <- cty_features_and_metadata$Domain
  trct_var_names <- trct_features_and_metadata$`Variable Name`
  trct_domains <- trct_features_and_metadata$Domain
  
  # Combine the lists of variable names and their corresponding domains into a data frame
  all_var_info <- data.frame(
    Variable = c(cty_var_names, trct_var_names),
    Domain = c(cty_domains, trct_domains)
  )
  
  # Calculate the missingness rates for each specified column
  missing_non_numeric_rates <- sapply(all_var_info$Variable, function(col) {
    if (col %in% names(input_df)) {
      # Calculate the rate of missing or non-numeric values
      mean(is.na(input_df[[col]]) | !is.numeric(as.numeric(as.character(input_df[[col]])))) * 100
    } else {
      NA
    }
  })
  
  # Add the missingness rates to the all_var_info data frame
  all_var_info$MissingRate <- missing_non_numeric_rates
  
  # Remove rows with NA rates (for variables not found in input_df)
  all_var_info <- all_var_info[!is.na(all_var_info$MissingRate), ]
  
  # Save the data frame with missingness rates and domain info to a CSV file
  write.csv(all_var_info, output_file, row.names = FALSE)
  
  # Print confirmation message and return all_var_info
  print(paste("Missing rates with domain information saved to", output_file))
  return(all_var_info)
}


#################

# Function to convert non-numeric to NA
convert_nonnumeric_to_na <- function(df, columns) {
  # Loop through specified columns
  for(col in columns) {
    if(col %in% names(df)) {
      # Convert column to numeric, non-numeric values will become NA
      df[[col]] <- suppressWarnings(as.numeric(df[[col]]))
    }
  }
  return(df)
}

# Define deprivation columns
deprivation_cols_total <- c("percnt_ltfpl100", "percnt_singlparntfly", "percnt_dropout", 
                           "percnt_hhnocar", "percnt_rentoccup", "percnt_crowding", 
                           "percnt_nonemp", "percnt_unemp", "percnt_highneeds", 
                           "percnt_hispanic", "percnt_frgnborn", "percnt_black", 
                           "sdi", "ADI_state_rank", "ADI_national_rank")

# Apply the conversion to patient_cohort
patient_cohort <- convert_nonnumeric_to_na(patient_cohort, deprivation_cols_total)


```

# Final DF (ignore DF1-4 below.) 
[This is equivalent to DF1_nm]
- Across nation: 35964 (we choose this)
- GA only: 33579.
```{r}

patient_cohort_final <- prepare_inputDF(patient_cohort)

# Continue with your existing code
missing_rates_df <- calculate_missing_rates(patient_cohort_final, 
                                          cty_features_and_metadata, 
                                          trct_features_and_metadata, 
                                          "missing_rates_Circ_final_allstates_modelinput.csv")

write.csv(patient_cohort_final, 
          file = "./Circ_final_allstates_modelinput.csv", 
          row.names = FALSE)

nrow(patient_cohort_final)

patient_cohort_final$Encounter

# Find patients not in the final cohort
patients_excluded = patient_cohort[~patient_cohort['Encounter'].isin(patient_cohort_final['Encounter'])]

```



Circ_final_allstates_modelinput.csv ****
vs. 
Circ_allstates_includemissing_modelinput_DF1.csv



# DF 1 [all states]: our final approach matches this above.

[Keeping missing deprivation, Y/N keep missing AHRQ.] produces 2 identical patient cohorts.
```{r}


# TODO: Save column names (tract, county).
# TODO: get missingness rates for each variable.

patient_cohort_A <- prepare_inputDF(patient_cohort)


# Continue with your existing code
missing_rates_df <- calculate_missing_rates(patient_cohort_A, 
                                          cty_features_and_metadata, 
                                          trct_features_and_metadata, 
                                          "missing_rates_Circ_allstates_includemissing_modelinput_DF1.csv")

write.csv(patient_cohort_A, 
          file = "./Circ_allstates_includemissing_modelinput_DF1.csv", 
          row.names = FALSE)


table(patient_cohort_A$readmit30bin_folds)
table(patient_cohort_A$death30bin_folds)
table(patient_cohort_A$composite30_folds)
nrow(patient_cohort_A)


# sum(is.na(patient_cohort_A$ADI_state_rank))/nrow(patient_cohort_A)

nrow(patient_cohort_A)

```
# DF 2 [all states]: 
[DROP missing deprivation, (Y/N) Don't use missing AHRQ columns.] -- produces 2 identical patient cohorts.

```{r}


# TODO: Save column names (tract, county).
# TODO: get missingness rates for each variable.

# 1. Only select patients (rows) with non-missing deprivation index vars (ADI, SDI, SVI).
# Drop row if there is a non-numerical or NA value for ANY of the following columns: 
# No column: "CensusTract_population", in patient_cohort...
# Define the deprivation columns as before
#complete_deprivation_cols_total <- c("ADI_national_rank", "ADI_state_rank", "sdi_score", "PovertyLT100_FPL_score", 
#                            "Single_Parent_Fam_score", "Education_LT12years_score", "HHNo_Vehicle_score", 
#                            "HHRenter_Occupied_score", "HHCrowding_score", "Nonemployed_score", "sdi", 
#                            "pct_Poverty_LT100", "pct_Single_Parent_Fam", "pct_Education_LT12years", 
#                            "pct_NonEmployed", "pctHH_No_Vehicle", "pctHH_Renter_Occupied", "pctHH_Crowding", 
#                            "SPL_THEMES", "EPL_POV", "EPL_UNEMP", "EPL_PCI", "EPL_NOHSDP", "EPL_AGE65", 
#                            "EPL_AGE17", "EPL_DISABL", "EPL_SNGPNT", "EPL_MINRTY", "EPL_LIMENG", "EPL_MUNIT", 
#                            "EPL_MOBILE", "EPL_CROWD", "EPL_NOVEH", "EPL_GROUPQ", "RPL_THEME1", "RPL_THEME2", 
#                            "RPL_THEME3", "RPL_THEME4", "RPL_THEMES", "F_TOTAL", "E_HU", "E_HH", "E_POV", 
#                            "E_UNEMP", "E_PCI", "E_NOHSDP", "E_AGE65", "E_AGE17", "E_DISABL", "E_SNGPNT", 
#                            "E_MINRTY", "E_LIMENG", "E_MUNIT", "E_MOBILE", "E_CROWD", "E_NOVEH", "E_GROUPQ", 
#                            "EP_POV", "EP_UNEMP", "EP_PCI", "EP_NOHSDP", "EP_AGE65", "EP_AGE17", "EP_DISABL", 
#                            "EP_SNGPNT", "EP_MINRTY", "EP_LIMENG", "EP_MUNIT", "EP_MOBILE", "EP_CROWD", 
#                            "EP_NOVEH", "EP_GROUPQ")


# Only drop patients if they have missing/non-numerical values for any of the the 15 "Limited SDOH" variables from Circ.
deprivation_cols_total <- c("percnt_ltfpl100", "percnt_singlparntfly", "percnt_dropout", "percnt_hhnocar", "percnt_rentoccup", 
                            "percnt_crowding", "percnt_nonemp", "percnt_unemp", "percnt_highneeds", "percnt_hispanic", "percnt_frgnborn",
                            "percnt_black", "sdi", "ADI_state_rank", "ADI_national_rank")



# Define variable names from cty_features_and_metadata and trct_features_and_metadata
cty_var_names <- cty_features_and_metadata$`Variable Name`
trct_var_names <- trct_features_and_metadata$`Variable Name`

# Combine lists of variable names and keep only columns present in patient_cohort
all_var_names <- union(cty_var_names, trct_var_names)
all_var_names_in_df <- all_var_names[all_var_names %in% names(patient_cohort)]



# Part 1: Filter patient_cohort_B based on deprivation columns with non-numerical or NA values
deprivation_cols_in_df <- deprivation_cols_total[deprivation_cols_total %in% names(patient_cohort)]
patient_cohort_B <- patient_cohort %>%
  filter(across(all_of(deprivation_cols_in_df), ~ !is.na(.) & !is.na(as.numeric(as.character(.)))))

cat("Number of rows dropped due to missing deprivation variables: ", (nrow(patient_cohort) - nrow(patient_cohort_B)), "\n")
cat("Rows before: ", nrow(patient_cohort), "\n")
cat("Rows after: ", nrow(patient_cohort_B), "\n")


# Part 2: Calculate missingness rates for each column in all_var_names
missing_non_numeric_rates <- sapply(all_var_names_in_df, function(col) {
  mean(is.na(patient_cohort_B[[col]]) | !is.numeric(as.numeric(as.character(patient_cohort_B[[col]])))) * 100
})

# Convert the results to a data frame
missing_rates_df <- data.frame(
  Variable = names(missing_non_numeric_rates),
  MissingRate = missing_non_numeric_rates
)

# Add the "Domain" column from the metadata
missing_rates_df <- merge(missing_rates_df, 
                          rbind(cty_features_and_metadata[, c("Variable Name", "Domain")], 
                                trct_features_and_metadata[, c("Variable Name", "Domain")]),
                          by.x = "Variable", by.y = "Variable Name", all.x = TRUE)

# Save the missing rates data frame to a CSV file
write.csv(missing_rates_df, "missing_rates_Circ_allstates_EXCLUDEmissing_modelinput_DF2.csv", row.names = FALSE)

# Confirm the file is saved
print("Initial missing rates with domain information saved to missing_rates_Circ_allstates_initial.csv")






# 3. To prepare subgroup quartile vars, & create folds:
patient_cohort_B <- prepare_inputDF(patient_cohort_B)


# ML4H_Circ_Nonmissing_Cohort
write.csv(patient_cohort_B, file="./Circ_allstates_EXCLUDEmissing_modelinput_DF2.csv", row.names=FALSE) 

table(patient_cohort_B$readmit30bin_folds)
table(patient_cohort_B$death30bin_folds)
table(patient_cohort_B$composite30_folds)
nrow(patient_cohort_B)



```

























#
OLD CODE BELOW -- restricts to GA, 1st keeps patients w missing deprivation, 2nd drops patients w missing deprivation.
#
# DF 3 [GA only]: include all columns (regardless of how many patients have values missing).
```{r}


# TODO: Save column names (tract, county).
# TODO: get missingness rates for each variable.
patient_cohort_C <- patient_cohort[patient_cohort$STATEFIPS == 13, ]

# table(patient_cohort_C$STATEFIPS, useNA = "ifany") # Patient count in Georgia (FIPS "13"): 33579

missing_rates_df <- calculate_missing_rates(patient_cohort_A, cty_features_and_metadata, trct_features_and_metadata, "missing_rates_Circ_GA_only_includemissing_modelinput_DF3.csv")

patient_cohort_C <- prepare_inputDF(patient_cohort_C)


# ML4H_Circ_Nonmissing_Cohort
write.csv(patient_cohort_C, file="./Circ_GA_only_includemissing_modelinput_DF3.csv", row.names=FALSE) 

table(patient_cohort_C$readmit30bin_folds)
table(patient_cohort_C$death30bin_folds)
table(patient_cohort_C$composite30_folds)
nrow(patient_cohort_C)


```

# DF 4 [GA only]: only include columns with 0% missing across patients. 
```{r}


# TODO: Save column names (tract, county).
# TODO: get missingness rates for each variable.
patient_cohort_D <- patient_cohort[patient_cohort$STATEFIPS == 13, ]

# table(patient_cohort_D$STATEFIPS, useNA = "ifany") # Patient count in Georgia (FIPS "13"): 33579
# 1. Only select pat ients (rows) with non-missing deprivation index vars (ADI, SDI, SVI).
# 2. Only selct AHRQ cols with 0% missing patients..


# Only drop patients if they have missing/non-numerical values for any of the the 15 "Limited SDOH" variables from Circ.
deprivation_cols_total <- c("percnt_ltfpl100", "percnt_singlparntfly", "percnt_dropout", "percnt_hhnocar", "percnt_rentoccup", 
                            "percnt_crowding", "percnt_nonemp", "percnt_unemp", "percnt_highneeds", "percnt_hispanic", "percnt_frgnborn",
                            "percnt_black", "sdi", "ADI_state_rank", "ADI_national_rank")



# Define variable names from cty_features_and_metadata and trct_features_and_metadata
cty_var_names <- cty_features_and_metadata$`Variable Name`
trct_var_names <- trct_features_and_metadata$`Variable Name`

# Combine lists of variable names and keep only columns present in patient_cohort
all_var_names <- union(cty_var_names, trct_var_names)
all_var_names_in_df <- all_var_names[all_var_names %in% names(patient_cohort)]



# Part 1: Filter patient_cohort_D based on deprivation columns with non-numerical or NA values
deprivation_cols_in_df <- deprivation_cols_total[deprivation_cols_total %in% names(patient_cohort)]
patient_cohort_D <- patient_cohort %>%
  filter(across(all_of(deprivation_cols_in_df), ~ !is.na(.) & !is.na(as.numeric(as.character(.)))))

cat("Number of rows dropped due to missing deprivation variables: ", (nrow(patient_cohort) - nrow(patient_cohort_D)), "\n")
cat("Rows before: ", nrow(patient_cohort), "\n")
cat("Rows after: ", nrow(patient_cohort_D), "\n")


# Part 2: Calculate missingness rates for each column in all_var_names
missing_non_numeric_rates <- sapply(all_var_names_in_df, function(col) {
  mean(is.na(patient_cohort_D[[col]]) | !is.numeric(as.numeric(as.character(patient_cohort_D[[col]])))) * 100
})

# Convert the results to a data frame
missing_rates_df <- data.frame(
  Variable = names(missing_non_numeric_rates),
  MissingRate = missing_non_numeric_rates
)

# Add the "Domain" column from the metadata
missing_rates_df <- merge(missing_rates_df, 
                          rbind(cty_features_and_metadata[, c("Variable Name", "Domain")], 
                                trct_features_and_metadata[, c("Variable Name", "Domain")]),
                          by.x = "Variable", by.y = "Variable Name", all.x = TRUE)

# Save the missing rates data frame to a CSV file
write.csv(missing_rates_df, "missing_rates_Circ_GA_only_EXCLUDEmissing_modelinput_DF4.csv", row.names = FALSE)

# Confirm the file is saved
print("Initial missing rates with domain information saved to missing_rates_Circ_allstates_initial.csv")








# 3. To prepare subgroup quartile vars, & create folds:
patient_cohort_D <- prepare_inputDF(patient_cohort_D)


# ML4H_Circ_Nonmissing_Cohort
write.csv(patient_cohort_D, file="./Circ_GA_only_EXCLUDEmissing_modelinput_DF4.csv", row.names=FALSE) 

table(patient_cohort_D$readmit30bin_folds)
table(patient_cohort_D$death30bin_folds)
table(patient_cohort_D$composite30_folds)
nrow(patient_cohort_D)


```










## Save all AHRQ feat names (County level, Tract level.)
```{r}

```


# 2.) County level: Save AHRQ feature names which are non-missing wrt ALL patient rows 
```{r}

```

# 2.) Tract level: Save AHRQ feature names which are non-missing wrt ALL patient rows 
```{r}

```




# AHRQ+EHR Missingness Calculation (Tract) : 
- EHR + Tract AHRQ : (given AHRQ features from 2010-2018 across all tract feats)
- across all counties, years. Not stratified at all.
- INTERPRETATION: for each AHRQ feature, "Number_Missing" describes the number of subjects in the dataset which do not have a value for that feature.
- INTERPRETATION: for each AHRQ feature, "Percent_Missing" describes the percent of subjects (out of n= 30702) with a missing value for that feature.
```{r}
# ahrq_baseline_tract

# Calculate missing values

# Num AHRQ county-level features we're interested in
# print(length(colnames(ga_cty_total)))

missing_values <- sapply(ahrq_baseline_tract, function(x) sum(is.na(x)))
percent_missing <- (missing_values / nrow(ahrq_baseline_tract)) * 100

# Combine the summary and missing value information
summary_info <- data.frame(
  Variable = names(ahrq_baseline_tract),
  Type = sapply(ahrq_baseline_tract, class),
  Number_Subjects_Missing_Feature = missing_values,
  Percent_Subjects_Missing_Feature = percent_missing
)

# Rename vars in 'summary_info'


# Merge summary_info with feature metadata. So we have variable descriptions.

composite_summary_info <- merge(summary_info, trct_features_and_metadata, by.x="Variable", by.y="Variable Name")

# composite_summary_info$Percent_Subjects_Missing_Feature <- format(composite_summary_info$Percent_Subjects_Missing_Feature, scientific = # FALSE)

trct_features_and_metadata

# Print the summary information
print(composite_summary_info) 

# OF 1267 features, at county-level, 506 have 0% missingness, 526 have <= 20% missingness
filter(composite_summary_info , Percent_Subjects_Missing_Feature <= 0) # %>% count(Domain)

hist(filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0.2)$Percent_Subjects_Missing_Feature)

hist(composite_summary_info$Percent_Subjects_Missing_Feature)


# Feats that have 0% of patients missing.
tract_non_missing_feats <- filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0)[ , c("Variable", "Domain")] 
tract_somepatients_missing_feats <- filter(composite_summary_info, Percent_Subjects_Missing_Feature > 0)$Variable 


write.csv(composite_summary_info, "./HF_patients_AHRQ_tractlevel_missingness.csv")


```

# AHRQ+EHR Missingness Calculation (County): 
- EHR + County AHRQ : (given AHRQ features from 2010-2018 across all county feats)
- across all counties, years. Not stratified at all.
- INTERPRETATION: for each AHRQ feature, "Number_Missing" describes the number of subjects in the dataset which do not have a value for that feature.
- INTERPRETATION: for each AHRQ feature, "Percent_Missing" describes the percent of subjects (out of n= 30702) with a missing value for that feature.
```{r}
# Calculate missing values

# Num AHRQ county-level features we're interested in
# print(length(colnames(ga_cty_total)))

missing_values <- sapply(ahrq_baseline, function(x) sum(is.na(x)))
percent_missing <- (missing_values / nrow(ahrq_baseline)) * 100

# Combine the summary and missing value information
summary_info <- data.frame(
  Variable = names(ahrq_baseline),
  Type = sapply(ahrq_baseline, class),
  Number_Subjects_Missing_Feature = missing_values,
  Percent_Subjects_Missing_Feature = percent_missing
)

# Merge summary_info with feature metadata. So we have variable descriptions.
composite_summary_info <- merge(summary_info, cty_features_and_metadata, by.x="Variable", by.y="Variable Name")

#composite_summary_info$Percent_Subjects_Missing_Feature <- format(composite_summary_info$Percent_Subjects_Missing_Feature, scientific = # FALSE)

# Print the summary information
print(composite_summary_info)


# OF 1267 features, at county-level, 506 have 0% missingness, 526 have <= 20% missingness
filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0) # %>% count(Domain)

hist(filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0.2)$Percent_Subjects_Missing_Feature )

hist(composite_summary_info$Percent_Subjects_Missing_Feature)


# Feats that have 0% of patients missing.
county_non_missing_feats <- filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0)[ , c("Variable", "Domain")] 
county_somepatients_missing_feats <- filter(composite_summary_info, Percent_Subjects_Missing_Feature > 0)$Variable 

write.csv(composite_summary_info, "./HF_patients_AHRQ_countylevel_missingness.csv")


```


# From 'county_non_missing_feats', 'tract_non_missing_feats', drop AHRQ features from category 6 (geography).
These are not relevant variables to prediction.
```{r}

# Tract feats non-missing:
nrow(tract_non_missing_feats)
tract_non_missing_feats <- subset(tract_non_missing_feats, Domain != '6. geography')
nrow(tract_non_missing_feats)


# County feats non-missing:
nrow(county_non_missing_feats)
county_non_missing_feats <- subset(county_non_missing_feats, Domain != '6. geography')
nrow(county_non_missing_feats)


```


# Remove AHRQ feats with >0% of patients missing them ("county_somepatients_missing_feats")
- can adjust to keep different feats here. 
- Then, feed "county_non_missing_feats" as model AHRQ feats in JSON feature file.
```{r}

# keep all AHRQ feats here. Can choose in python which ones to keep.
# ahrq_baseline <- ahrq_baseline[, !(names(ahrq_baseline) %in% county_non_missing_feats)]


# Domain column: shows domain each feat belongs to.
# AHRQ county-level feats we KEPT:
write.csv(county_non_missing_feats, "./AHRQ_featlist_county_nonmissing.csv", row.names=FALSE)
# AHRQ tract-level feats we KEPT:
write.csv(tract_non_missing_feats, "./AHRQ_featlist_tract_nonmissing.csv", row.names=FALSE)

```


- add feats to json file (diff domains, then all domains total)
- adjust MongoCol. 
- preprocess/scale feats!!


# Build subgroup variables.
```{r}
###################################
# Merge: ahrq_baseline + baseline_AQI_imp_1 (both already imputed on orig datasets. Possible issue is high missingness for several features.)
###################################


# First, match types of merge columns

#ahrq_baseline # COUNTYFIPS: <int>, 'YEAR' <int> here is baselineyear from orig_df (i.e. tractmlm)
#baseline_AQI_imp_1 # County_FIPS: <int>, YEAR: <chr>... cast year to <int>
#original_df <- original_df %>%
#  mutate(YEAR = as.numeric(YEAR), na.rm=FALSE)

ahrq_baseline <- ahrq_baseline %>%
  mutate(baselineyear = as.numeric(baselineyear), na.rm=FALSE)

# Then merge
# ahrq_AQI_baseline_final <- merge(ahrq_baseline, original_df, by.x=c("COUNTYFIPS", "YEAR"), by.y=c("COUNTY_FIPS", "YEAR"))
#sum(is.na(ahrq_baseline$ACS_PCT_PUB_COMMT_59MIN)) # Should be 0 for all columns.


############ Prepare subgroup flags:
# Subgroups to eval on: 
# 1. Race
# white <int>: 0/1
# black <int>: 0/1
# other <int>: 0/1
ahrq_baseline <- ahrq_baseline %>%
  mutate(blackorwhite = if_else(black == 1 | white == 1, 1, 0))

# 2. Sex (or binary gender)
# female <int>: 0/1
# ADD this field: male <int>: 0/1
ahrq_baseline <- ahrq_baseline %>%
  mutate(male = if_else(female == 0, 1, 0))

# 3. The 6 subgroup combinations: 
# c(white_male, black_male, other_male, white_female, black_female, other_female)

# Note that all 6 are mutually exclusive here.
ahrq_baseline <- ahrq_baseline %>%
  mutate(white_male = if_else(white == 1 & female == 0, 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(black_male = if_else((black == 1 & female == 0), 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(other_male = if_else(other == 1 & female == 0, 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(white_female = if_else(white == 1 & female == 1, 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(black_female = if_else(black == 1 & female == 1, 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(other_female = if_else(other == 1 & female == 1, 1, 0))



#######################
# SUBGROUP FOR AGEs: 

quarts <- quantile(ahrq_baseline$age_current, probs = seq(0, 1, 0.25))
upper_Q1_age <- quarts[2]
upper_Q2_age <- quarts[3]
upper_Q3_age <- quarts[4]
upper_Q4_age <- quarts[5]

ahrq_baseline <- ahrq_baseline %>%
  mutate(age_Q1 = if_else(age_current <= upper_Q1_age, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(age_Q2 = if_else(age_current <= upper_Q2_age & age_current > upper_Q1_age, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(age_Q3 = if_else(age_current <= upper_Q3_age & age_current > upper_Q2_age, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(age_Q4 = if_else(age_current <= upper_Q4_age & age_current > upper_Q3_age, 1, 0))



#######################
# HF: Sex & age quartiles

#######################
# Hypertension
ahrq_baseline <- ahrq_baseline %>%
  mutate(no_htn = if_else(htn == 0, 1, 0))

#######################
# Coronary Artery Disease
ahrq_baseline <- ahrq_baseline %>%
  mutate(no_cad = if_else(cad == 0, 1, 0))

#######################
# Chronic Kidney Disease
ahrq_baseline <- ahrq_baseline %>%
  mutate(no_ckd = if_else(ckd == 0, 1, 0))


#######################
# Charlson Comorbidity Index
  # Theoretical Max CCI: 24 (updated version) or 29 (older version).
  # Our dataset: Max CCI = 22.
CCI_quarts <- quantile(ahrq_baseline$cci, probs = seq(0, 1, 0.25))

# 1.) CCI steps 0-6, 7+ groups.
# 


ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_0 = if_else(cci == 0, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_1 = if_else(cci == 1, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_2 = if_else(cci == 2, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_3 = if_else(cci == 3, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_4 = if_else(cci == 4, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_5 = if_else(cci == 5, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_6 = if_else(cci == 6, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_7up = if_else(cci >= 7, 1, 0))





# 2.) CCI quartiles (second way to evaluate CCI)
upper_Q1_CCI <- CCI_quarts[2]
upper_Q2_CCI <- CCI_quarts[3]
upper_Q3_CCI <- CCI_quarts[4]
upper_Q4_CCI <- CCI_quarts[5]


ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_Q1 = if_else(cci <= upper_Q1_CCI, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_Q2 = if_else(cci <= upper_Q2_CCI & cci > upper_Q1_CCI, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_Q3 = if_else(cci <= upper_Q3_CCI & cci > upper_Q2_CCI, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_Q4 = if_else(cci > upper_Q3_CCI, 1, 0))
  

nrow(ahrq_baseline) # Sample size: 30702
############

endpoints <- c("readmit30bin", "death30bin", "composite30")
ahrq_baseline <- create_folds(as.data.frame(ahrq_baseline), endpoints) # NOTE: For nested k-fold CV.

hist(ahrq_baseline$sdi)


ahrq_baseline


# ML4H_Circ_Nonmissing_Cohort
write.csv(ahrq_baseline, file="./ahrq_ML4H_Circ_Nonmissing_Cohort.csv", row.names=FALSE) 
# Old: CHIL_no_missingfeats_county_ahrq_HF_modelinputs.csv

table(ahrq_baseline$readmit30bin_folds)

table(ahrq_baseline$death30bin_folds)
table(ahrq_baseline$composite30_folds)

# Ex run: 80-20 train-test split --> then k-fold s
#set.seed(42)
#sample <- sample(c(TRUE, FALSE), nrow(ahrq_baseline), replace=TRUE, prob=c(0.8,0.2))
#train  <- ahrq_baseline[sample, ]
#test   <- ahrq_baseline[!sample, ]

```

# ###############################
# Read in model input data
```{r}

input_ahrq_patient_feats <- read.csv(file="./ahrq_ML4H_Circ_Nonmissing_Cohort.csv") # CHIL_no_missingfeats_county_ahrq_HF_modelinputs.csv

# ML4H_Circ_Nonmissing_Cohort

# Check readmit30bin positive rates by "race" subgroup, fold.

input_ahrq_patient_feats %>% group_by(Race, readmit30bin_folds) %>% summarize(newvar = mean(readmit30bin, na.rm = TRUE))


########################

# What races are there?
table(input_ahrq_patient_feats$Race)


# All: should be ONLY black or white patients.
table(input_ahrq_patient_feats$blackorwhite)

# Count black vs. white.
table(input_ahrq_patient_feats$white)
table(input_ahrq_patient_feats$black)




```








































# County: GA state missing calculations
- across all counties, years. Not stratifed at all.
```{r}
# Calculate missing values
missing_values <- sapply(ga_cty_total, function(x) sum(is.na(x)))
percent_missing <- missing_values / nrow(ga_cty_total) * 100

# Combine the summary and missing value information
summary_info <- data.frame(
  Variable = names(ga_cty_total),
  Type = sapply(ga_cty_total, class),
  Number_Missing = missing_values,
  Percent_Missing = percent_missing
)

# Merge summary_info with feature metadata. So we have variable descriptions.
composite_summary_info <- merge(summary_info, cty_features_and_metadata, by.x="Variable", by.y="Variable Name")

# Print the summary information
print(composite_summary_info)

write.csv(composite_summary_info, "./AHRQ_GA_countylevel_missingness_composite2010-2018.csv")

```
# Feats missingness: by year (2010-2018).
- Note: there are 159 counties in Georgia.
```{r}
# Num counties in Georgia:
# length(unique(ga_cty_total$COUNTYFIPS))
GA_num_counties <- length(unique(ga_cty_total$COUNTYFIPS))

# Calculate missing values stratified by year
summary_info <- ga_cty_total %>%
  group_by(YEAR) %>%
  summarize(
    across(everything(), ~sum(is.na(.))),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = -YEAR, names_to = "Variable", values_to = "Number_GA_Ctys_Missing") %>%
  mutate(
    Percent_GA_Ctys_Missing = Number_GA_Ctys_Missing / GA_num_counties * 100
  )

# Print the modified summary information
print(summary_info) 


```

# Feats missingness: by year (in wide format)
```{r}

# Calculate missing values stratified by YEAR
summary_info <- ga_cty_total %>%
  group_by(YEAR) %>%
  summarize(
    across(everything(), ~sum(is.na(.))),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = -YEAR, names_to = "Variable", values_to = "Number_GA_Ctys_Missing") %>%
  mutate(
    Percent_GA_Ctys_Missing = Number_GA_Ctys_Missing / GA_num_counties * 100
  ) %>%
  pivot_wider(names_from = YEAR, values_from = c("Number_GA_Ctys_Missing", "Percent_GA_Ctys_Missing"), names_glue = "{.value}_{YEAR}")



# Merge summary_info with feature metadata. So we have variable descriptions.
yr_composite_summary_info <- merge(summary_info, cty_features_and_metadata, by.x="Variable", by.y="Variable Name")

# Print 
print(yr_composite_summary_info)

write.csv(yr_composite_summary_info, "./AHRQ_GA_countylevel_missingness_yearly2010-2018.csv")


```

# AHRQ missigness by Domain: 
- Hypothesis: some AHRQ domains (1-5) have much larger missingness rates than other AHRQ domains.
- NOTE: These missing counts & rates are still JUST for Georgia.
```{r}
# For colors scheme
library(ggsci)


# First, count num variables in each domain. 


domain_counts <- yr_composite_summary_info %>%
  group_by(Domain) %>%
  summarize(Num_Variables = n_distinct(Variable))

table(domain_counts)


#ggplot(domain_counts, aes(x = Domain, y = Num_Variables)) +
#  geom_bar(stat = "identity", color = "black", fill = "#BA5C4E") +  # Set fill to gray for no coloring
#  labs(title = "Number of Features in Each AHRQ Domain (Georgia, County)",
#       x = "AHRQ Domain",
#       y = "Number of Variables") +
#  theme_minimal() +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels


# * GA_Ctys

# Select relevant columns for 2010 through 2018
missing_columns <- colnames(yr_composite_summary_info)[grep("^Number_GA_Ctys_Missing_201[0-9]{1}$", colnames(yr_composite_summary_info))]

# Calculate the number of missing values for each Domain
missing_by_domain <- yr_composite_summary_info %>%
  group_by(Domain) %>%
  summarize(
    across(all_of(missing_columns), ~sum(.)),
    .groups = "drop"
  )


# Calculate the percentage of missing values for each Domain
missing_percentage_by_domain <- yr_composite_summary_info %>%
  group_by(Domain) %>%
  summarize(
    across(all_of(missing_columns), ~ round(sum(.))),
    .groups = "drop"
  ) %>%
  rename_with(~ gsub("^Number_GA_Ctys_Missing_(\\d{4})$", "Percent_GA_Ctys_Missing_\\1", .), .cols = contains("Number_GA_Ctys_Missing"))


# Create missing proportions from missing counts.
numeric_missing_percentage_by_domain <- sweep(missing_percentage_by_domain[, -1], MARGIN = 1, STATS = domain_counts$Num_Variables * GA_num_counties, FUN = "/")

# Create percentages from missing proportions.
numeric_missing_percentage_by_domain <- numeric_missing_percentage_by_domain * 100

# Prepend domain names back in.
final_missing_percentage_by_domain <- cbind(missing_percentage_by_domain[1], numeric_missing_percentage_by_domain)

# ----> IN DEMO WHEN CALCULATING compositie precentage of counties missing features values per domain: domain_counts$Domain
# .... check on this calculation. What do we want the percent to show?

# Print the result
print(missing_by_domain)
print(final_missing_percentage_by_domain)


```

# Dummy figure: self-calculated data for AHRQ domain dims.
```{r}
library(ggplot2)

dum_dims <- data.frame(
  AHRQ_dim_name = c("1. Social Context", "2. Economic Context", "3. Education", "4. Physical Infrastructure", "5. Healthcare Context"),
  Var_Count = c(109, 62, 11, 106, 207)
)

barplot(dum_dims$Var_Count, names.arg = dum_dims$AHRQ_dim_name, col = "#BA5C4E", xlab = "AHRQ Domain", ylab = "Number of Features", las=3)


# Create a ggplot bar chart with slanted x-axis labels
ggplot(dum_dims, aes(x = AHRQ_dim_name, y = Var_Count, fill = AHRQ_dim_name)) +
  geom_bar(stat = "identity", position = "dodge",  fill = "#BA5C4E", color = "white") +
  labs(x = "AHRQ Domain",
       y = "Number of Features") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=11),
        axis.text.y = element_text(size = 11))

```


# ##########################################
# TRACT CALCULATIONS (Geogia): 
# ##########################################

```{r}
# (b) GA Level: census tracts

```



# (1) MISSING ANALYSIS: Total AHRQ data (2010-2018)
```{r}
#library(naniar)

# -- For each, plot missing for (i) each year 2010-2018
# Col 1: var name, Col 2: ahrq dimension of 5 (& orig data source), 

# (a) GA Level: counties
ga_tract_total <-  tr_ahrq_total[tr_ahrq_total$STATEFIPS == 13, ]

```

# Tract: GA state missing calculations
- across all tracts, years. Not stratifed at all.
```{r}
# Calculate missing values
missing_values <- sapply(ga_tract_total, function(x) sum(is.na(x)))
percent_missing <- missing_values / nrow(ga_tract_total) * 100

# Combine the summary and missing value information
summary_info <- data.frame(
  Variable = names(ga_tract_total),
  Type = sapply(ga_tract_total, class),
  Number_Missing = missing_values,
  Percent_Missing = percent_missing
)

# Merge summary_info with feature metadata. So we have variable descriptions.
composite_summary_info <- merge(summary_info, trct_features_and_metadata, by.x="Variable", by.y="Variable Name")

# Print the summary information
print(composite_summary_info)

write.csv(composite_summary_info, "./AHRQ_GA_tractlevel_missingness_composite2010-2018.csv")

```


# Feats missingness (count, % of GA tracts): by year (2010-2018).
- Note: there are 1969 census tracts in Georgia. 
```{r}
# Num counties in Georgia:
# length(unique(ga_cty_total$COUNTYFIPS))
GA_num_tracts <- length(unique(ga_tract_total$TRACTFIPS ))

# Calculate missing values stratified by year
summary_info <- ga_tract_total %>%
  group_by(YEAR) %>%
  summarize(
    across(everything(), ~sum(is.na(.))),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = -YEAR, names_to = "Variable", values_to = "Number_GA_Tracts_Missing") %>%
  mutate(
    Percent_GA_Tracts_Missing = Number_GA_Tracts_Missing / GA_num_tracts * 100
  )

# Print the modified summary information
print(summary_info)

```



# Feats missingness: by year (in wide format)
```{r}

# Calculate missing values stratified by YEAR
summary_info <- ga_tract_total %>%
  group_by(YEAR) %>%
  summarize(
    across(everything(), ~sum(is.na(.))),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = -YEAR, names_to = "Variable", values_to = "Number_GA_Tracts_Missing") %>%
  mutate(
    Percent_GA_Tracts_Missing = Number_GA_Tracts_Missing / GA_num_tracts * 100
  ) %>%
  pivot_wider(names_from = YEAR, values_from = c("Number_GA_Tracts_Missing", "Percent_GA_Tracts_Missing"), names_glue = "{.value}_{YEAR}")



# Merge summary_info with feature metadata. So we have variable descriptions.
yr_composite_summary_info <- merge(summary_info, trct_features_and_metadata, by.x="Variable", by.y="Variable Name")

# Print 
print(yr_composite_summary_info)

write.csv(yr_composite_summary_info, "./AHRQ_GA_tractlevel_missingness_yearly2010-2018.csv")


```



# AHRQ missigness by Domain: 
- Hypothesis: some AHRQ domains (1-5) have much larger missingness rates than other AHRQ domains.
- NOTE: These missing counts & rates are still JUST for Georgia.
```{r}
# For colors scheme
library(ggsci)


# First, count num variables in each domain. 


domain_counts <- yr_composite_summary_info %>%
  group_by(Domain) %>%
  summarize(Num_Variables = n_distinct(Variable))




ggplot(domain_counts, aes(x = Domain, y = Num_Variables)) +
  geom_bar(stat = "identity", color = "black", fill = "#BA5C4E") +  # Set fill to gray for no coloring
  labs(title = "Number of Features in Each AHRQ Domain (Georgia, Tract)",
       x = "AHRQ Domain",
       y = "Number of Variables") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels


# * GA_Ctys

# Select relevant columns for 2010 through 2018
missing_columns <- colnames(yr_composite_summary_info)[grep("^Number_GA_Tracts_Missing_201[0-9]{1}$", colnames(yr_composite_summary_info))]

# Calculate the number of missing values for each Domain
missing_by_domain <- yr_composite_summary_info %>%
  group_by(Domain) %>%
  summarize(
    across(all_of(missing_columns), ~sum(.)),
    .groups = "drop"
  )


# Calculate the percentage of missing values for each Domain
missing_percentage_by_domain <- yr_composite_summary_info %>%
  group_by(Domain) %>%
  summarize(
    across(all_of(missing_columns), ~ round(sum(.))),
    .groups = "drop"
  ) %>%
  rename_with(~ gsub("^Number_GA_Tracts_Missing_(\\d{4})$", "Percent_GA_Tracts_Missing_\\1", .), .cols = contains("Number_GA_Tracts_Missing"))


# Create missing proportions from missing counts.
numeric_missing_percentage_by_domain <- sweep(missing_percentage_by_domain[, -1], MARGIN = 1, STATS = domain_counts$Num_Variables * GA_num_tracts, FUN = "/")

# Create percentages from missing proportions.
numeric_missing_percentage_by_domain <- numeric_missing_percentage_by_domain * 100

# Prepend domain names back in.
final_missing_percentage_by_domain <- cbind(missing_percentage_by_domain[1], numeric_missing_percentage_by_domain)

# ----> IN DEMO WHEN CALCULATING compositie precentage of counties missing features values per domain: domain_counts$Domain
# .... check on this calculation. What do we want the percent to show?

# Print the result
print(missing_by_domain)
print(final_missing_percentage_by_domain)


```






# ####################
# EHR data below
# ####################


# Merge: sectry_orig_df <> (cty_ahrq_total & tr_ahrq_total)

```{r}
final_df <- sectry_orig_df

# Variables for Census block groups:
# Not included in this analysis
# Entire 15 digit number: final_df$censusblock, 
# Substrings of that number: final_df$blockgroup, final_df$blocknum, final_df$cxy_block_id, final_df$censusblockgroup

# First, remove rows in final_df with an NA for any of the above: censustract, countynum, OR statenum
# --- TODO - also in other Rmd
final_df <- final_df %>% drop_na(censustract)
final_df <- final_df %>% drop_na(tractnum)

final_df <- final_df %>% drop_na(countynum)
final_df <- final_df %>% drop_na(statenum)


### 1.) CHECK COUNTY FIPS value consistency
# Note: 35 subjects have different "county FIPS" values (countynum) from orig tractmlm values vs. newer (cxy_county_id) queried values.
nrow(final_df)
#final_df$cxy_county_id[final_df$countynum != final_df$cxy_county_id]
#final_df$countynum[final_df$countynum != final_df$cxy_county_id]


### 2.) CHECK TRACT FIPS value consistency
```

# Cast Census Tract (i.e., censustract variable), County Fips (i.e., fips variable) to proper length for merge with AHRQ (keep these 2 cols as type *chr*)
```{r}
#
#final_df$tractnum # 6 dig: str: just tract 
#final_df$censustract # Full number (must be 11 characters long)
#final_df$cxy_tract_id # 6 dig: Newer queried values: just tract
# final_df$cxy_tract_id[final_df$tractnum != final_df$cxy_tract_id]

# County: to 5 chars. Note: 'fips' has state[2 chr]+county[3 chr]
#final_df$fips
final_df$fips <- sprintf("%05s", final_df$fips)

# Tract: to 11 chars
#final_df$censustract
final_df$censustract <- sprintf("%011s", final_df$censustract)



head(final_df )
```


```{r}
# Prepare AHRQ join keys (county, tract datasets):
# cty_ahrq_total$COUNTYFIPS # 4-5 digit int (e.g., 4 digit min: 1001, 5 digit max: 78030)
# Coerce to 5 digit *chr*: 
ga_cty_total$COUNTYFIPS_padded <- sprintf("%05s", ga_cty_total$COUNTYFIPS)

# TRACT: state + county + tract
# tr_ahrq_total$TRACTFIPS # 10-11 digit *double* (e.g., 10 digit min: 1001020100, 11 digit max: 78030990000). Do NOT coerce to int, numbers will overflow.
# Coerce to 11 digit *chr*:
ga_tract_total$TRACTFIPS_padded <- sprintf("%011s", ga_tract_total$TRACTFIPS)


# 2. Rename feats suffix in ga_cty_total, ga_tract_total (_countylevel, _census_tractlevel):



#ga_cty_total <- ga_cty_total %>%
#  rename_at(vars(-one_of(c("COUNTY", "COUNTYFIPS_padded", "REGION", "STATE", "STATEFIPS", "TERRITORY", "YEAR"))), ~paste0(., #"_countylevel"))


# Tract:
#exclude_columns_trct <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "TRACTFIPS_padded", "YEAR")


#ga_tract_total <- ga_tract_total %>%
#  rename_at(vars(-one_of(exclude_columns_trct)), ~paste0(., "_census_tractlevel"))


# MERGE: both on subject's 'baselineyear' (admission year)
# Bring in County AHRQ:

ahrq_baseline_cty <- merge(final_df, ga_cty_total, by.x=c("fips", "baselineyear"), by.y=c("COUNTYFIPS_padded", "YEAR"))
# TODO: -- use "ga_tract_total"
# Tract AHRQ baseline: use later for patient-level cenus tract missingness calculations. 
ahrq_baseline_tract <- merge(final_df, ga_tract_total, by.x=c("censustract", "baselineyear"), by.y=c("TRACTFIPS_padded", "YEAR"))

# both_ahrq_baseline <- merge(ga_tract_total, ga_cty_total, by.x=c("TRACTFIPS_padded", "YEAR"), by.y=c("censustract", "YEAR"))

ahrq_baseline <- merge(ahrq_baseline_cty, ga_tract_total, by.x=c("censustract", "baselineyear"), by.y=c("TRACTFIPS_padded", "YEAR"))

# Finally, merge in SVI 2014 data:
# use cty code: ahrq_baseline ++ SVI_2014_US (by 'censustract', 'FIPS')

ahrq_baseline <- merge(ahrq_baseline, SVI_2014_US  , by.x="censustract", by.y="FIPS" )

head(ahrq_baseline) # Patient data + matched county & tract. Names: _countlevel, _census_tractlevel

```
# Before feeding to classifiers, drop potential PHI from ahrq_baseline
```{r}

drop_vars <- c("Lon__Lat", "cxy_address", "cxy_matched_address", "cxy_tiger_line_id", "cxy_tract_id", "cxy_block_id")
ncol(ahrq_baseline)

# Remove drop_vars
ahrq_baseline <- ahrq_baseline[, !(names(ahrq_baseline) %in% drop_vars)]

ncol(ahrq_baseline)

```



# AHRQ+EHR Missingness Calculation (Tract) : 
- EHR + Tract AHRQ : (given AHRQ features from 2010-2018 across all tract feats)
- across all counties, years. Not stratified at all.
- INTERPRETATION: for each AHRQ feature, "Number_Missing" describes the number of subjects in the dataset which do not have a value for that feature.
- INTERPRETATION: for each AHRQ feature, "Percent_Missing" describes the percent of subjects (out of n= 30702) with a missing value for that feature.
```{r}
# ahrq_baseline_tract

# Calculate missing values

# Num AHRQ county-level features we're interested in
# print(length(colnames(ga_cty_total)))

missing_values <- sapply(ahrq_baseline_tract, function(x) sum(is.na(x)))
percent_missing <- (missing_values / nrow(ahrq_baseline_tract)) * 100

# Combine the summary and missing value information
summary_info <- data.frame(
  Variable = names(ahrq_baseline_tract),
  Type = sapply(ahrq_baseline_tract, class),
  Number_Subjects_Missing_Feature = missing_values,
  Percent_Subjects_Missing_Feature = percent_missing
)

# Rename vars in 'summary_info'


# Merge summary_info with feature metadata. So we have variable descriptions.

composite_summary_info <- merge(summary_info, trct_features_and_metadata, by.x="Variable", by.y="Variable Name")

# composite_summary_info$Percent_Subjects_Missing_Feature <- format(composite_summary_info$Percent_Subjects_Missing_Feature, scientific = # FALSE)

trct_features_and_metadata

# Print the summary information
print(composite_summary_info) 

# OF 1267 features, at county-level, 506 have 0% missingness, 526 have <= 20% missingness
filter(composite_summary_info , Percent_Subjects_Missing_Feature <= 0) # %>% count(Domain)

hist(filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0.2)$Percent_Subjects_Missing_Feature)

hist(composite_summary_info$Percent_Subjects_Missing_Feature)


# Feats that have 0% of patients missing.
tract_non_missing_feats <- filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0)[ , c("Variable", "Domain")] 
tract_somepatients_missing_feats <- filter(composite_summary_info, Percent_Subjects_Missing_Feature > 0)$Variable 


write.csv(composite_summary_info, "./HF_patients_AHRQ_tractlevel_missingness.csv")


```

# AHRQ+EHR Missingness Calculation (County): 
- EHR + County AHRQ : (given AHRQ features from 2010-2018 across all county feats)
- across all counties, years. Not stratified at all.
- INTERPRETATION: for each AHRQ feature, "Number_Missing" describes the number of subjects in the dataset which do not have a value for that feature.
- INTERPRETATION: for each AHRQ feature, "Percent_Missing" describes the percent of subjects (out of n= 30702) with a missing value for that feature.
```{r}
# Calculate missing values

# Num AHRQ county-level features we're interested in
# print(length(colnames(ga_cty_total)))

missing_values <- sapply(ahrq_baseline, function(x) sum(is.na(x)))
percent_missing <- (missing_values / nrow(ahrq_baseline)) * 100

# Combine the summary and missing value information
summary_info <- data.frame(
  Variable = names(ahrq_baseline),
  Type = sapply(ahrq_baseline, class),
  Number_Subjects_Missing_Feature = missing_values,
  Percent_Subjects_Missing_Feature = percent_missing
)

# Merge summary_info with feature metadata. So we have variable descriptions.
composite_summary_info <- merge(summary_info, cty_features_and_metadata, by.x="Variable", by.y="Variable Name")

#composite_summary_info$Percent_Subjects_Missing_Feature <- format(composite_summary_info$Percent_Subjects_Missing_Feature, scientific = # FALSE)

# Print the summary information
print(composite_summary_info)


# OF 1267 features, at county-level, 506 have 0% missingness, 526 have <= 20% missingness
filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0) # %>% count(Domain)

hist(filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0.2)$Percent_Subjects_Missing_Feature )

hist(composite_summary_info$Percent_Subjects_Missing_Feature)


# Feats that have 0% of patients missing.
county_non_missing_feats <- filter(composite_summary_info, Percent_Subjects_Missing_Feature <= 0)[ , c("Variable", "Domain")] 
county_somepatients_missing_feats <- filter(composite_summary_info, Percent_Subjects_Missing_Feature > 0)$Variable 

write.csv(composite_summary_info, "./HF_patients_AHRQ_countylevel_missingness.csv")


```

# Mutate metadata's AHRQ feature names: ..._countylevel, ..._census_tractlevel. 
```{r}

head(county_non_missing_feats)
length(trct_features_and_metadata$`Variable Name`)
length(cty_features_and_metadata$`Variable Name`)
length(set_intersection(trct_features_and_metadata$`Variable Name`, cty_features_and_metadata$`Variable Name`))
#####
# Mutate:
# County:
# Create new columns for county-level features with a suffix

# Columns to exclude from renaming
#exclude_columns_cty <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "YEAR")


#cty_features_and_metadata <- cty_features_and_metadata %>%
#  mutate(`Variable Name` = case_when(
#    `Variable Name` %in% c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "YEAR") ~ `Variable Name`,
#    TRUE ~ paste0(`Variable Name`, "_countylevel")
#  ))

# Tract:
#exclude_columns_trct <- c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "TRACTFIPS", "YEAR")

#trct_features_and_metadata <- trct_features_and_metadata %>%
#  mutate(`Variable Name` = case_when(
#    `Variable Name` %in% c("COUNTY", "COUNTYFIPS", "REGION", "STATE", "STATEFIPS", "TERRITORY", "TRACTFIPS", "YEAR") ~ `Variable Name`,
#    TRUE ~ paste0(`Variable Name`, "_census_tractlevel")
#  ))

```
# From 'county_non_missing_feats', 'tract_non_missing_feats', drop AHRQ features from category 6 (geography).
These are not relevant variables to prediction.
```{r}

# Tract feats non-missing:
nrow(tract_non_missing_feats)
tract_non_missing_feats <- subset(tract_non_missing_feats, Domain != '6. geography')
nrow(tract_non_missing_feats)


# County feats non-missing:
nrow(county_non_missing_feats)
county_non_missing_feats <- subset(county_non_missing_feats, Domain != '6. geography')
nrow(county_non_missing_feats)


```


# Remove AHRQ feats with >0% of patients missing them ("county_somepatients_missing_feats")
- can adjust to keep different feats here. 
- Then, feed "county_non_missing_feats" as model AHRQ feats in JSON feature file.
```{r}

# keep all AHRQ feats here. Can choose in python which ones to keep.
# ahrq_baseline <- ahrq_baseline[, !(names(ahrq_baseline) %in% county_non_missing_feats)]


# Domain column: shows domain each feat belongs to.
# AHRQ county-level feats we KEPT:
write.csv(county_non_missing_feats, "./AHRQ_featlist_county_nonmissing.csv", row.names=FALSE)
# AHRQ tract-level feats we KEPT:
write.csv(tract_non_missing_feats, "./AHRQ_featlist_tract_nonmissing.csv", row.names=FALSE)

```


- add feats to json file (diff domains, then all domains total)
- adjust MongoCol. 
- preprocess/scale feats!!


# Build subgroup variables.
```{r}
###################################
# Merge: ahrq_baseline + baseline_AQI_imp_1 (both already imputed on orig datasets. Possible issue is high missingness for several features.)
###################################


# First, match types of merge columns

#ahrq_baseline # COUNTYFIPS: <int>, 'YEAR' <int> here is baselineyear from orig_df (i.e. tractmlm)
#baseline_AQI_imp_1 # County_FIPS: <int>, YEAR: <chr>... cast year to <int>
#original_df <- original_df %>%
#  mutate(YEAR = as.numeric(YEAR), na.rm=FALSE)

ahrq_baseline <- ahrq_baseline %>%
  mutate(baselineyear = as.numeric(baselineyear), na.rm=FALSE)

# Then merge
# ahrq_AQI_baseline_final <- merge(ahrq_baseline, original_df, by.x=c("COUNTYFIPS", "YEAR"), by.y=c("COUNTY_FIPS", "YEAR"))
#sum(is.na(ahrq_baseline$ACS_PCT_PUB_COMMT_59MIN)) # Should be 0 for all columns.


############ Prepare subgroup flags:
# Subgroups to eval on: 
# 1. Race
# white <int>: 0/1
# black <int>: 0/1
# other <int>: 0/1
ahrq_baseline <- ahrq_baseline %>%
  mutate(blackorwhite = if_else(black == 1 | white == 1, 1, 0))

# 2. Sex (or binary gender)
# female <int>: 0/1
# ADD this field: male <int>: 0/1
ahrq_baseline <- ahrq_baseline %>%
  mutate(male = if_else(female == 0, 1, 0))

# 3. The 6 subgroup combinations: 
# c(white_male, black_male, other_male, white_female, black_female, other_female)

# Note that all 6 are mutually exclusive here.
ahrq_baseline <- ahrq_baseline %>%
  mutate(white_male = if_else(white == 1 & female == 0, 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(black_male = if_else((black == 1 & female == 0), 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(other_male = if_else(other == 1 & female == 0, 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(white_female = if_else(white == 1 & female == 1, 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(black_female = if_else(black == 1 & female == 1, 1, 0))
ahrq_baseline <- ahrq_baseline %>%
  mutate(other_female = if_else(other == 1 & female == 1, 1, 0))



#######################
# SUBGROUP FOR AGEs: 

quarts <- quantile(ahrq_baseline$age_current, probs = seq(0, 1, 0.25))
upper_Q1_age <- quarts[2]
upper_Q2_age <- quarts[3]
upper_Q3_age <- quarts[4]
upper_Q4_age <- quarts[5]

ahrq_baseline <- ahrq_baseline %>%
  mutate(age_Q1 = if_else(age_current <= upper_Q1_age, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(age_Q2 = if_else(age_current <= upper_Q2_age & age_current > upper_Q1_age, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(age_Q3 = if_else(age_current <= upper_Q3_age & age_current > upper_Q2_age, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(age_Q4 = if_else(age_current <= upper_Q4_age & age_current > upper_Q3_age, 1, 0))



#######################
# HF: Sex & age quartiles

#######################
# Hypertension
ahrq_baseline <- ahrq_baseline %>%
  mutate(no_htn = if_else(htn == 0, 1, 0))

#######################
# Coronary Artery Disease
ahrq_baseline <- ahrq_baseline %>%
  mutate(no_cad = if_else(cad == 0, 1, 0))

#######################
# Chronic Kidney Disease
ahrq_baseline <- ahrq_baseline %>%
  mutate(no_ckd = if_else(ckd == 0, 1, 0))


#######################
# Charlson Comorbidity Index
  # Theoretical Max CCI: 24 (updated version) or 29 (older version).
  # Our dataset: Max CCI = 22.
CCI_quarts <- quantile(ahrq_baseline$cci, probs = seq(0, 1, 0.25))

# 1.) CCI steps 0-6, 7+ groups.
# 


ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_0 = if_else(cci == 0, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_1 = if_else(cci == 1, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_2 = if_else(cci == 2, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_3 = if_else(cci == 3, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_4 = if_else(cci == 4, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_5 = if_else(cci == 5, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_6 = if_else(cci == 6, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_7up = if_else(cci >= 7, 1, 0))





# 2.) CCI quartiles (second way to evaluate CCI)
upper_Q1_CCI <- CCI_quarts[2]
upper_Q2_CCI <- CCI_quarts[3]
upper_Q3_CCI <- CCI_quarts[4]
upper_Q4_CCI <- CCI_quarts[5]


ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_Q1 = if_else(cci <= upper_Q1_CCI, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_Q2 = if_else(cci <= upper_Q2_CCI & cci > upper_Q1_CCI, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_Q3 = if_else(cci <= upper_Q3_CCI & cci > upper_Q2_CCI, 1, 0))

ahrq_baseline <- ahrq_baseline %>%
  mutate(CCI_Q4 = if_else(cci > upper_Q3_CCI, 1, 0))
  

nrow(ahrq_baseline) # Sample size: 30702
############

endpoints <- c("readmit30bin", "death30bin", "composite30")
ahrq_baseline <- create_folds(as.data.frame(ahrq_baseline), endpoints) # NOTE: For nested k-fold CV.

hist(ahrq_baseline$sdi)


ahrq_baseline


# ML4H_Circ_Nonmissing_Cohort
write.csv(ahrq_baseline, file="./ahrq_ML4H_Circ_Nonmissing_Cohort.csv", row.names=FALSE) 
# Old: CHIL_no_missingfeats_county_ahrq_HF_modelinputs.csv

table(ahrq_baseline$readmit30bin_folds)

table(ahrq_baseline$death30bin_folds)
table(ahrq_baseline$composite30_folds)

# Ex run: 80-20 train-test split --> then k-fold s
#set.seed(42)
#sample <- sample(c(TRUE, FALSE), nrow(ahrq_baseline), replace=TRUE, prob=c(0.8,0.2))
#train  <- ahrq_baseline[sample, ]
#test   <- ahrq_baseline[!sample, ]

```

# ###############################
# Read in model input data
```{r}
# setwd('/Users/chasefensore/Documents/Emory/Ho Lab/SU23 Rotation/sdoh_summer/aequitus/data/CHIL Total/')


input_ahrq_patient_feats <- read.csv(file="./ahrq_ML4H_Circ_Nonmissing_Cohort.csv") # CHIL_no_missingfeats_county_ahrq_HF_modelinputs.csv

# ML4H_Circ_Nonmissing_Cohort

# Check readmit30bin positive rates by "race" subgroup, fold.

input_ahrq_patient_feats %>% group_by(Race, readmit30bin_folds) %>% summarize(newvar = mean(readmit30bin, na.rm = TRUE))


########################

# What races are there?
table(input_ahrq_patient_feats$Race)


# All: should be ONLY black or white patients.
table(input_ahrq_patient_feats$blackorwhite)

# Count black vs. white.
table(input_ahrq_patient_feats$white)
table(input_ahrq_patient_feats$black)




```





```{r}
nrow(input_ahrq_patient_feats)









```



# Table 1: WIHOUT "other"

--> TODO: 
- Col 1: "Subgroups"
- Col 2: "# of Subjects"
- Col 3: "% of Subjects"
- Col 4: % of Subjects in Positive Class

```{r}
## tableone package itself
library(tableone)
library(kableExtra)

# Make strata column: white, black, other (1,2,3)
# make using cols:
# Drop "Race" == "other"


# ahrq_ML4H_Circ_Nonmissing_Cohort.csv
input_ahrq_patient_feats <- read.csv(file="./ahrq_ML4H_Circ_Nonmissing_Cohort.csv") # CHIL_no_missingfeats_county_ahrq_HF_modelinputs.csv

input_feats <- input_ahrq_patient_feats
trans_input_feats <- input_feats[! (input_feats$Race == "other"), ]


trans_input_feats <- trans_input_feats <- transform(
  input_feats, "Race"= ifelse(white==1, "White", ifelse(black==1, "Black", NA)))
  

# Make insurance column: "privinsurance","Medicare", "Medicaid",
trans_input_feats <- data_frame <- transform(
  trans_input_feats, "Insurance_type"= ifelse(privinsurance==1, "Private", ifelse(Medicare==1, "Medicare", ifelse(Medicaid==1, "Medicaid", "Not recorded"))))
  

# Make HF type: hfsystolic, hfdiastolic, hfother
trans_input_feats <- data_frame <- transform(
  trans_input_feats, "HF_classification"= ifelse(hfsystolic==1, "Systolic (HFrEF)", ifelse(hfdiastolic==1, "Diastolic (HFpEF)", ifelse(hfother==1, "Other", "Not recorded"))))



# vars to summarize (i.e demographics)
# htn: Hypertension
# diabetes: ___not available___
# cad: Coronary Artery Disease
# ckd: Chronic Kidney Disease
# sdi: Social deprivation index score
# ADI_national_rank: 
# ADI_state_rank: 
# svi: i.e. RPL_THEMES
# cci: Charlson Comorbidity Index
# Systolic_BP: Systolic blood pressure, mm Hg
# Heart_Rate: Heart Rate, beats per minute
# Resp: Respiratory rate, breaths per minute

# Rename feats not needing spaces: 
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_current"] ="Age, years"
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_Q1"] ="Age, Q1"
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_Q2"] ="Age, Q2"
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_Q3"] ="Age, Q3"
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_Q4"] ="Age, Q4"

colnames(trans_input_feats)[colnames(trans_input_feats) == "female"] = "Female"
colnames(trans_input_feats)[colnames(trans_input_feats) == "male"] = "Male"

colnames(trans_input_feats)[colnames(trans_input_feats) == "black"] = "Black"
colnames(trans_input_feats)[colnames(trans_input_feats) == "white"] = "White"

colnames(trans_input_feats)[colnames(trans_input_feats) == "htn"] = "Hypertension"
colnames(trans_input_feats)[colnames(trans_input_feats) == "cad"] = "Coronary Artery Disease"
colnames(trans_input_feats)[colnames(trans_input_feats) == "ckd"] = "Chronic Kidney Disease"
colnames(trans_input_feats)[colnames(trans_input_feats) == "cci"] = "Charlson Comorbidity Index"
colnames(trans_input_feats)[colnames(trans_input_feats) == "Systolic_BP"] = "Systolic blood pressure, mm Hg"
colnames(trans_input_feats)[colnames(trans_input_feats) == "Heart_Rate"] = "Heart Rate, beats per minute"
colnames(trans_input_feats)[colnames(trans_input_feats) == "Resp"] = "Respiratory rate, breaths per minute"
colnames(trans_input_feats)[colnames(trans_input_feats) == "sdi"] = "Social deprivation index (percentile)"
colnames(trans_input_feats)[colnames(trans_input_feats) == "ADI_national_rank"] = "Area deprivation index, national rank (percentile)"
colnames(trans_input_feats)[colnames(trans_input_feats) == "ADI_state_rank"] = "Area deprivation index, state rank (percentile)"
colnames(trans_input_feats)[colnames(trans_input_feats) == "RPL_THEMES"] = "Social vulnerability index score (composite score)"

# New feats:
#colnames(trans_input_feats)[colnames(trans_input_feats) == "readmit30bin"] = "30-day Binary HF Readmission"
#colnames(trans_input_feats)[colnames(trans_input_feats) == "death30bin"] = "30-day Binary HF Death"

# Referring specialties: 
colnames(trans_input_feats)[colnames(trans_input_feats) == "specialcvd"] = "CVD Specialist" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "specialint"] = "Int. Med. Specialist"
colnames(trans_input_feats)[colnames(trans_input_feats) == "specialeme"] = "EME Specialist"
colnames(trans_input_feats)[colnames(trans_input_feats) == "specialoth"] = "Other Specialist"

# Location treated: 
colnames(trans_input_feats)[colnames(trans_input_feats) == "loceuhfloor"] = "UH Floor" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "loceclhicu"] = "CLH ICU"
colnames(trans_input_feats)[colnames(trans_input_feats) == "loceclhfloor"] = "CLH Floor"
colnames(trans_input_feats)[colnames(trans_input_feats) == "loceuhicu"] = "UH ICU"

# SDOH feats (comp -- other Circulation HF paper)

colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_ltfpl100"] = "percnt_ltfpl100" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_singlparntfly"] = "percnt_singlparntfly" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_dropout"] = "percnt_dropout" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_rentoccup"] = "percnt_rentoccup" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_crowding"] = "percnt_crowding" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_nonemp"] = "percnt_nonemp" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_unemp"] = "percnt_unemp" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_highneeds"] = "percnt_highneeds" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_hispanic"] = "percnt_hispanic" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_frgnborn"] = "percnt_frgnborn" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_black"] = "percnt_black" 
# sdi
# ADI_state_rank
# ADI_national_rank


# Rename feats with spaces: 
colnames(trans_input_feats)[colnames(trans_input_feats) == "Insurance_type"] = "Insurance type"
colnames(trans_input_feats)[colnames(trans_input_feats) == "HF_classification"] = "HF classification"




# Length of stay, days: LOScheck
#myVars <- c("Age, years", "female", "Insurance type", "HF classification", "htn", "cad", "ckd", "cci", 
#            "Systolic_BP", "Heart_Rate", "Resp", 
#            "LOScheck", "sdi", "ADI_national_rank", "ADI_state_rank", "RPL_THEMES")

myVars <- c("Age, years","Age, Q1","Age, Q2","Age, Q3","Age, Q4", "Male", "Female", "Black", "White",
            "CVD Specialist", "Int. Med. Specialist", "EME Specialist", "Other Specialist",
            "UH Floor", "CLH ICU", "CLH Floor", "UH ICU",
            
            "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease", 
            "Charlson Comorbidity Index", 
            "Systolic blood pressure, mm Hg", "Heart Rate, beats per minute", "Respiratory rate, breaths per minute", 
            "LOScheck", "Social deprivation index (percentile)", "Area deprivation index, national rank (percentile)", "Area deprivation index, state rank (percentile)", "Social vulnerability index score (composite score)",
            "percnt_ltfpl100", "percnt_singlparntfly", "percnt_dropout", "percnt_rentoccup", "percnt_crowding", "percnt_nonemp", "percnt_unemp", "percnt_highneeds", "percnt_hispanic", "percnt_frgnborn", "percnt_black"
            
            )


## Vector of categorical variables that need transformation
catVars <- c("Women", "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease")

tab1 <- CreateTableOne(vars = myVars, strata = "readmit30bin", data = trans_input_feats, addOverall = TRUE) 
# For race strata instead of outcome strata: strata = "Race"
# For p-values: test=TRUE, smd = TRUE

# typeof(tab1["test"])



# If you just say nonnormal = TRUE, all variables are summarized the “nonnormal” way.
#print(tab1, formatOptions = list(big.mark = ","))
# kable(p, format = "latex")
p <- print(tab1, printToggle = FALSE, noSpaces = TRUE)
p <- p[,1:3] # Remove "test" column.

kable(p, format = "markdown", booktabs = TRUE)
```



# For tableone missingness:
```{r}
summary(tab1)
```





