---
title: "new_HF_performance_2024"
output: html_notebook
---

# Load libraries
```{r}
library(caret)
library(data.table)
library(dplyr)
library(sets)
library(scales)
library(tidyr)
library(stringr)
library(tibble)
library(tidygeocoder)
library(stringi)
library(ggplot2)


library(psych)
library(tidyverse)
library(irr)


library(readr) # For read_tsv
```


# Model input files: 
- input file:  
    - Circ_allstates_includemissing_modelinput_DF1.csv
- output file:



# Model Circ HF results: 
- input file: DF1_results_individualfolds_Circ_2024.csv
- 

- train file: private_CHIL_repo/scripts/bal_allfeats_MLHC_nosmote_sgs_evaluate_baselines_nestKfold.py
- output file: private_CHIL_repo/scripts/3-21-ACC/Circ_HF_results_individualfolds.csv

```{r}

fairness_full_hf_results <- read.csv("../scripts/individualfolds_Circ_HF_results_fairness_final.csv") # individualfolds_Circ_HF_results_fairness_final

full_hf_results <- read.csv("../scripts/final_results_individualfolds_Circ_2024.csv") # 

```

# Subset results: 
```{r}

logr_hf_results <- full_hf_results[full_hf_results$model == "logr", ]
xgb_hf_results <- full_hf_results[full_hf_results$model == "xgb", ]
rf_hf_results <- full_hf_results[full_hf_results$model == "rf", ]



fairness_full_hf_results <- fairness_full_hf_results[fairness_full_hf_results$sg_key == "race", ]

fair_logr_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "logr", ]
fair_xgb_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "xgb", ]
fair_rf_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "rf", ]



full_hf_results
fairness_full_hf_results
```

# Print md table of all feature set results: 
- AUC
- F1
# Subgroups: blackorwhite, black, white.
```{r}
# Required library for markdown tables

#curr_model_results <- xgb_hf_results # logr_hf_results, xgb_hf_results

library(dplyr)
library(tidyr)
library(knitr)

#     "M1",
#        "M2",
#        "M3_county_DF1_nm",
#        "M4_DF1_nm",
#        "M5",
#        "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
#    
#        "DF1_nm_county_AHRQ_domain1",
#        "DF1_nm_county_AHRQ_domain2",
#        "DF1_nm_county_AHRQ_domain3",
#        "DF1_nm_county_AHRQ_domain4",
#        "DF1_nm_county_AHRQ_domain5",
#
#        "DF1_nm_tract_AHRQ_domain1",
#        "DF1_nm_tract_AHRQ_domain2",
#        "DF1_nm_tract_AHRQ_domain3",
#        "DF1_nm_tract_AHRQ_domain4",
#        "DF1_nm_tract_AHRQ_domain5"

feat_sets <- c(
   "M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm",
        "M5",
        "M6_total_ahrq_cty_DF1_nm")


sdoh_feat_sets <- c(
    "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
    
        "DF1_nm_county_AHRQ_domain1",
        "DF1_nm_county_AHRQ_domain2",
        "DF1_nm_county_AHRQ_domain3",
        "DF1_nm_county_AHRQ_domain4",
        "DF1_nm_county_AHRQ_domain5",

        "DF1_nm_tract_AHRQ_domain1",
        "DF1_nm_tract_AHRQ_domain2",
        "DF1_nm_tract_AHRQ_domain3",
        "DF1_nm_tract_AHRQ_domain4",
        "DF1_nm_tract_AHRQ_domain5")


# Ensure all required subgroups are present in the data
grouped_results <- curr_model_results %>%
  filter(feat %in% sdoh_feat_sets, subgroup %in% c("blackorwhite", "black", "white")) %>%
  group_by(feat, subgroup) %>%
  summarize(
    auc = mean(auc, na.rm = TRUE),
    f1 = mean(f1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = subgroup,
    values_from = c(auc, f1),
    names_glue = "{subgroup}_{.value}"
  )

# Reorder columns in grouped_results
grouped_results <- grouped_results %>%
  select(
    feat, 
    blackorwhite_auc, black_auc, white_auc, 
    blackorwhite_f1, black_f1, white_f1
  )

# Format numeric columns to three decimal places
grouped_results <- grouped_results %>%
  mutate(
    across(
      starts_with(c("blackorwhite_", "black_", "white_")),
      ~ sprintf("%.3f", .) # Format as a string with 3 decimal places
    )
  )

# Create the markdown table
kable(
  grouped_results,
  col.names = c(
    "Feature Set", 
    "AUROC (All Patients)", "AUROC (Black)", "AUROC (White)", 
    "F1 (All Patients)", "F1 (Black)", "F1 (White)"
  ),
  caption = "Summary of AUROC and F1 Score by Feature Set for Each Subgroup"
)


colnames(logr_hf_results)
```
# [PERFORMANCE] Print md table: for all 3 models:
```{r}
# Combine the logr, rf, and xgb results
curr_model_results <- bind_rows(
  logr_hf_results %>% mutate(model = "logr"),
  rf_hf_results %>% mutate(model = "rf"),
  xgb_hf_results %>% mutate(model = "xgb")
)


feat_sets <- c(
   "M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm",
        "M5",
        "M6_total_ahrq_cty_DF1_nm")

sdoh_feat_sets <- c(
  "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
  "DF1_nm_county_AHRQ_domain1", "DF1_nm_county_AHRQ_domain2", "DF1_nm_county_AHRQ_domain3",
  "DF1_nm_county_AHRQ_domain4", "DF1_nm_county_AHRQ_domain5", "DF1_nm_tract_AHRQ_domain1",
  "DF1_nm_tract_AHRQ_domain2", "DF1_nm_tract_AHRQ_domain3", "DF1_nm_tract_AHRQ_domain4",
  "DF1_nm_tract_AHRQ_domain5"
)

# Ensure all required subgroups are present in the data
grouped_results <- curr_model_results %>%
  filter(feat %in% sdoh_feat_sets, subgroup %in% c("blackorwhite", "black", "white")) %>%
  group_by(feat, model) %>%
  summarize(
    auroc_all = mean(auc[subgroup == "blackorwhite"], na.rm = TRUE),
    auroc_black = mean(auc[subgroup == "black"], na.rm = TRUE), 
    auroc_white = mean(auc[subgroup == "white"], na.rm = TRUE),
    f1_all = mean(f1[subgroup == "blackorwhite"], na.rm = TRUE),
    f1_black = mean(f1[subgroup == "black"], na.rm = TRUE),
    f1_white = mean(f1[subgroup == "white"], na.rm = TRUE),
    .groups = "drop"
  )

# Format numeric columns to three decimal places
grouped_results <- grouped_results %>%
  mutate(
    across(
      starts_with(c("auroc_", "f1_")),
      ~ sprintf("%.3f", .)
    )
  )

# Create the markdown table
kable(
  grouped_results,
  col.names = c(
    "Feature Set", "Model",
    "AUROC (All Patients)", "AUROC (Black)", "AUROC (White)",
    "F1 (All Patients)", "F1 (Black)", "F1 (White)"
  ),
  caption = "Summary of AUROC and F1 Score by Feature Set for Each Subgroup", format = "markdown"
)

```

# Fairness results:
```{r}
# Required libraries
library(dplyr)
library(tidyr)
library(knitr)

# Define feature sets
feat_sets <- c(
    "M1",
    "M2",
    "M3_county_DF1_nm",
    "M4_DF1_nm",
    "M5",
    "M6_total_ahrq_cty_DF1_nm"
)

sdoh_feat_sets <- c(
    "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
    "DF1_nm_county_AHRQ_domain1",
    "DF1_nm_county_AHRQ_domain2",
    "DF1_nm_county_AHRQ_domain3",
    "DF1_nm_county_AHRQ_domain4",
    "DF1_nm_county_AHRQ_domain5",
    "DF1_nm_tract_AHRQ_domain1",
    "DF1_nm_tract_AHRQ_domain2",
    "DF1_nm_tract_AHRQ_domain3",
    "DF1_nm_tract_AHRQ_domain4",
    "DF1_nm_tract_AHRQ_domain5"
)

# Filter, group, and summarize results
grouped_results <- curr_model_results %>%
  filter(feat %in% sdoh_feat_sets) %>%
  group_by(feat, fold) %>%
  summarize(
    eo_ratio = mean(eo_ratio, na.rm = TRUE),
    fpr_parity = mean(fpr_parity, na.rm = TRUE),
    tpr_parity = mean(tpr_parity, na.rm = TRUE),
    fnr_parity = mean(fnr_parity, na.rm = TRUE),
    dp_ratio = mean(dp_ratio, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(feat) %>%
  summarize(
    eo_ratio = round(mean(eo_ratio, na.rm = TRUE), 3),
    fpr_parity = round(mean(fpr_parity, na.rm = TRUE), 3),
    tpr_parity = round(mean(tpr_parity, na.rm = TRUE), 3),
    fnr_parity = round(mean(fnr_parity, na.rm = TRUE), 3),
    dp_ratio = round(mean(dp_ratio, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  select(feat, eo_ratio, fpr_parity, tpr_parity, fnr_parity, dp_ratio) # Reorder columns

# Create the markdown table
kable(
  grouped_results,
  col.names = c(
    "Feature Set", 
    "EO Ratio", "FPR Parity", "TPR Parity", 
    "FNR Parity", "DP Ratio"
  ),
  caption = "Summary of fairness by Feature Set for all patients (Black vs. White)"
)


```

# [FAIRNESS] Print md table: for all 3 models:
```{r}


fairness_full_hf_results <- fairness_full_hf_results[fairness_full_hf_results$sg_key == "race", ]

fair_logr_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "logr", ]
fair_xgb_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "xgb", ]
fair_rf_hf_results <- fairness_full_hf_results[fairness_full_hf_results$model == "rf", ]


# Combine the logr, rf, and xgb results
curr_model_results_fair <- bind_rows(
  fair_logr_hf_results %>% mutate(model = "logr"),
  fair_xgb_hf_results %>% mutate(model = "rf"),
  fair_rf_hf_results %>% mutate(model = "xgb")
)


feat_sets <- c(
   "M1",
        "M2",
        "M3_county_DF1_nm",
        "M4_DF1_nm",
        "M5",
        "M6_total_ahrq_cty_DF1_nm")

sdoh_feat_sets <- c(
  "M6_total_ahrq_cty_DF1_nm", "M6_total_ahrq_trct_DF1_nm",
  "DF1_nm_county_AHRQ_domain1", "DF1_nm_county_AHRQ_domain2", "DF1_nm_county_AHRQ_domain3",
  "DF1_nm_county_AHRQ_domain4", "DF1_nm_county_AHRQ_domain5", "DF1_nm_tract_AHRQ_domain1",
  "DF1_nm_tract_AHRQ_domain2", "DF1_nm_tract_AHRQ_domain3", "DF1_nm_tract_AHRQ_domain4",
  "DF1_nm_tract_AHRQ_domain5"
)

fairness_full_hf_results$dp_ratio

# Ensure all required subgroups are present in the data
grouped_results <- curr_model_results_fair %>%
  filter(feat %in% sdoh_feat_sets) %>%
  group_by(feat, model) %>%
  summarize(
    eo_ratio = mean(eo_ratio[sg_key == "race"], na.rm = TRUE),
    fpr_parity = mean(fpr_parity[sg_key == "race"], na.rm = TRUE), 
    tpr_parity = mean(tpr_parity[sg_key == "race"], na.rm = TRUE), 
    fnr_ratio = mean(fnr_parity[sg_key == "race"], na.rm = TRUE), 
    dp_ratio = mean(dp_ratio[sg_key == "race"], na.rm = TRUE), 
    .groups = "drop"
  )

# Format numeric columns to three decimal places
grouped_results <- grouped_results %>%
  mutate(
    across(
      starts_with(c("eo_", "fpr_", "tpr_", "fnr_", "dp_")),
      ~ sprintf("%.3f", .)
    )
  )

# Create the markdown table
kable(
  grouped_results,
  col.names = c(
    "Feature Set", "Model",
    "EO Ratio", "FPR Ratio", "TPR Ratio", "FNR Ratio",
    "DP Ratio"
  ),
  caption = "Summary of fairness metrics by model and feature set.", format = "markdown"
)

```

