---
title: "patient_inclusion_Circ_2024"
author: "Chase Fensore"
date: "2024-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tibble)
library(dplyr)
library(caret)

library(censusxy)
library(tidygeocoder)

```

# Read in tractmlm
```{r}
source('data_cleaners.R')

orig_df <- read.csv("./tractmlm.csv") # Input file of patients with HF with their encounters.
# Before restricting with inclusion criteria.
cat("Raw from tractmlm: \n")
nrow(orig_df)

```


# Inclusion criteria:
```{r}
# 1.) Clean up the end-point related stuff
# Restrict to endpoints: 
endpoints <- c("readmit30bin", "death30bin", "composite30")
for (colx in endpoints) {
	filtered_df <- remove_na(orig_df, colx)
}
cat("After endpoints inclusion: ", nrow(filtered_df), "\n") # After endpoints inclusion:  41614 



# 2.) Street address
# Useful columns: Patient_Home_Street_Address, Patient_Address (less full), fips, censustract, 
# Drop if: no value/missing/empty string like NA,"", ".", "..", "..." for *ALL* of the useful columns. Keep if at least 1 column is valid.
# Note: Patient_Home_Street_Address (listed in EHR) vs. Patient_Address (from matching)
# Define a vector of invalid values
invalid_values <- c(NA, "", ".", "..", "...")

# Filter rows where at least one of the useful columns has a valid value to describe patient address.
filtered_df <- filtered_df %>%
  filter(
    !(Patient_Home_Street_Address %in% invalid_values &
      Patient_Address %in% invalid_values &
      fips %in% invalid_values &
      censustract %in% invalid_values)
  )

# To find bad values: head(sort(table(orig_df$Patient_Home_Street_Address), decreasing = TRUE), 10)

# View the filtered data
cat("After invalid address exclusion: ", nrow(filtered_df), "\n") # After: 


# [Do not restrict to GA] 3.) Next, include only patients where "Patient_Home_State" == "GA". Excluded because (i) lived in another state, (ii) had "??" or has (iii) "NOT". 62 patients had "??" and 33 had "NOT"
# filtered_df <- filtered_df[filtered_df$Patient_Home_State == "GA", ]
# cat("After home state exclusion: ", nrow(filtered_df), "\n") # After: 38449


# 4.) Keep black or white patients only:
filtered_df <- filtered_df[(filtered_df$black == 1 | filtered_df$white == 1),]
cat("After keeping only Black or White patients: ", nrow(filtered_df), "\n") # After:

# 5.) Keep only adults aged >= 18
filtered_df <- filtered_df[(filtered_df$age_current >= 18), ]
cat("After adult inclusion: ", nrow(filtered_df), "\n") # After:

# Write out:
write.csv(filtered_df$Encounter, "encouterIDs_validaddress.csv", row.names = FALSE )

# After endpoints inclusion:  41614 [44442 - 41614: 2828 did not have endpoint data.]
# After invalid address exclusion:  41614 [all 44442 had some form of address... not necessarily able to be geocoded.]
# After keeping only Black or White patients:  39067 
# After adult inclusion:  39066 



```

# Among these 39066, how many have a valid "censustract" or "fips" value?
```{r}

invalid_values <- c(NA, "", ".", "..", "...")


# strict_filtered_df: Filter rows where at least one of the useful columns has a valid value to describe patient address.
strict_filtered_df <- filtered_df %>%
  filter(
    !(fips %in% invalid_values &
      censustract %in% invalid_values)
  )

cat("After we exclude patients who do not have at least 1 valid value between fips or censustract: ", nrow(strict_filtered_df)) # 31804

# Write out: 
write.csv(strict_filtered_df$Encounter, "encouterIDs_nonmissingFIPS_censustract.csv", row.names = FALSE )
nrow(strict_filtered_df)


sum(is.na(strict_filtered_df$ADI_national_rank)) # 2426
```

# Starting with n=39,066 patients across USA... how many geocodes are we missing? (censustract, fips, Patient_Home_Zip_Code)
```{r}
# Frequency of each "invalid value":
invalid_values <- c(NA, "", ".", "..", "...")


patient_cohort <- read.csv(file="./Circ_final_allstates_modelinput.csv") # Fully geocoded from patient_inclusion_Circ_2024.Rmd, ML4H_Circ_Nonmissing_Cohort.csv
cat("Number patients fully geocoded: ", nrow(patient_cohort), "\n")



# Convert to numeric without scientific notation
filtered_df$Encounter <- as.numeric(as.character(filtered_df$Encounter))
patient_cohort$Encounter <- as.numeric(as.character(patient_cohort$Encounter))


sum(duplicated(filtered_df$Encounter) | duplicated(filtered_df$Encounter, fromLast = TRUE))



length(unique(patient_cohort$Encounter)) # 35964
length(unique(filtered_df$Encounter)) # 39066
merged_df <- merge(filtered_df, patient_cohort, by = "Encounter")
nrow(merged_df) # 35890
length(unique(merged_df$Encounter)) # 35890




# Check merge conditions
merged_df <- merge(filtered_df, patient_cohort, by = "Encounter")
nrow(merged_df)  # Should show number of matching rows
length(unique(filtered_df$Encounter))  # Unique Encounters in filtered_df
length(unique(patient_cohort$Encounter))  # Unique Encounters in patient_cohort


# Find Encounters in filtered_df not in patient_cohort
unmatched_encounters <- filtered_df$Encounter[!filtered_df$Encounter %in% patient_cohort$Encounter]
length(unmatched_encounters)
head(unmatched_encounters)


# Detailed investigation of unmatched Encounters
unmatched_df <- filtered_df[filtered_df$Encounter %in% unmatched_encounters, ]
unmatched_df


```
# DELETE: Find patients in filtered_df (n=39066) which are NOT in patient_cohort (i.e., n=35964, final model input: Circ_final_allstates_modelinput.csv)
```{r}
options(scipen = 999)  # Disable scientific notation

#head(filtered_df)
#head(patient_cohort)


# NOTE: patients which are in patient_cohort but NOT in filtered_df: means that they were geocoded on 2nd try. 
# These were not dropped (n=74), they were included in final cohort, so we don't care about them now.
#missing_geocode_df <- anti_join(filtered_df, patient_cohort, by = "Encounter")

#nrow(missing_geocode_df) # 3176

# Then remove the 74 present in patient_cohort but NOT in filtered_df.
#to_remove <- anti_join(patient_cohort, filtered_df, by = "Encounter")
#nrow(to_remove)


# Save for anlaysis in table1.Rmd: n=3102
#missing_to_analyze <- missing_geocode_df[!missing_geocode_df$patient_name %in% to_remove$patient, ]
#nrow(missing_to_analyze)
```

# DELETE
```{r}
# CAST: 
#to_remove$Encounter <- format(to_remove$Encounter, scientific = FALSE)
#missing_to_analyze$Encounter <- format(missing_to_analyze$Encounter, scientific = FALSE)
# Remove leading whitespace from Encounter column
#to_remove$Encounter <- trimws(to_remove$Encounter, which = "left")
#missing_to_analyze$Encounter <- trimws(missing_to_analyze$Encounter, which = "left")

# Find common Encounters 
#common_encounters <- intersect(
#  missing_to_analyze %>% select(followup), 
#  to_remove %>% select(followup)
#)

#length(common_encounters)


# Find columns without duplicates
#unique_cols <- sapply(missing_to_analyze, function(x) length(unique(x)) == nrow(missing_to_analyze))
#names(which(unique_cols))

#to_remove$followup

```







```{r}
# Now, select ONLY the patients without geocodes.
missing_geocode_df <- filtered_df %>%
  filter(
    (fips %in% invalid_values |
      censustract %in% invalid_values)
  )
cat("Select all patients with INVALID fips or censustract values... but VALID addresses: ", nrow(missing_geocode_df), "\n") 


# Here, get the rows that are SUCCESSFULLY geocoded previously.
# Filter out rows in filtered_df that are in missing_geocode_df based on the "Encounter" column
non_missing_geocode_df <- anti_join(filtered_df, missing_geocode_df, by = "Encounter")
cat("Filtered df is before final geocoding attempt: ", nrow(filtered_df), "\n")
missing_geocode_df <- anti_join(filtered_df, patient_cohort, by = "Encounter")
cat("Missing geocode df is patients who met inclusion criteria (i.e., in filtered_df) but could not be geocoded after final attempt (i.e., not in patient_cohort): ", nrow(missing_geocode_df), "\n")
cat("Patient cohort is after all final patients are selected: ", nrow(patient_cohort), "\n")


# Adds up to 39066. 
# Among  7262 (from 39066) ,  4160  were correctly geocoded from address. 
# We want to select 7262 - 4160 = 3102 to analyze in our "missing geocode" study.




cat("Had successful geocodes previously: ", nrow(non_missing_geocode_df), "\n") # 31,804
cat("Missing geocode i.e., fips or censustract: ", nrow(missing_geocode_df), "\n") # 7,262

# Save patients with unsuccessful geocodes (but who meet all rest of inclusion criteria):
fname <- paste0("./odd-nonmissing_geocodedf_n-",  nrow(missing_geocode_df), ".csv")
drop_vars <- c("Patient", "patient_name", "Lon__Lat", "cxy_address", "cxy_matched_address", "cxy_tiger_line_id", "cxy_tract_id", "cxy_block_id")
# Remove drop_vars
missing_geocode_df <- missing_geocode_df[, !(names(missing_geocode_df) %in% drop_vars)]

write.csv(missing_geocode_df, fname)
cat("Printed patients with missing geocode (n=", nrow(missing_geocode_df), ") to csv.\n")


# there are 74 patients in patient_cohort (final 36k df) that are NOT in the original filtered_df...
anti_join(patient_cohort, filtered_df, by = "Encounter")


```


# ######################################################################

# Geocode the extra patients:
start with "missing_geocode_df" (n=7262)
```{r}
library(tibble)
library(dplyr)
library(tidygeocoder)
library(censusxy)
library(tidyr)
library(tidycensus) # For getting FIPS codes
library(purrr) # For mapping functions


cat("Select all patients with INVALID fips or censustract values... but VALID addresses: ", nrow(missing_geocode_df)) 
# [TOTAL] Complete address: Patient_Address

# USEFUL GEOCODING COLUMNS: 
# Patient_Home_Street_Address
# Patient_Home_City
# Patient_Home_State
# Patient_Home_Zip_Code
# Patient_Home_County

# Baseline year: use "baselineyear" to decide census boundary.


# Check adresses:
```


```{r}
# If needed
# census_api_key("5abbb66bb9e0d9f9b616e41d56776e6aca616005")

```





# Helper functions: using full address --> get tract, county FIPS values.
```{r}
library(httr)
library(jsonlite)

# API Docs: https://geocoding.geo.census.gov/geocoder/Geocoding_Services_API.html
get_fips_from_address <- function(address) {
  # Build the API URL
  base_url <- "https://geocoding.geo.census.gov/geocoder/geographies/onelineaddress"
  params <- list(
    address = address,
    benchmark = 4,# Uses "Public_AR_Current". To get 2020 values only, use benchmark "2020". Summary here: https://geocoding.geo.census.gov/geocoder/benchmarks
    vintage = 4 # Uses "Public_AR_Current"
  )
  
  # Make the API request
  response <- GET(base_url, query = params)
  
  # Check if the request was successful
  if (response$status_code == 200) {
    # Parse the response
    result <- fromJSON(content(response, "text"), flatten = TRUE)
    
    # Extract the address matches
    address_matches <- result$result$addressMatches
    
    # Check if there are any address matches
    if (length(address_matches) > 0) {
      # Extract the geographies
      geographies <- address_matches$geographies$`Census Tracts` # [["Census Tracts"]] #.Census.Tracts
      county_fips <- address_matches$geographies.Counties
      
      # cat("ADDRESS MATCHES: ", paste(address_matches, collapse = ", "), "\n")

      
      # Extract FIPS codes from the geographies
      # tract_fips <- ifelse(!is.null(geographies), sub("^(\\d{1,3})\\d*", "\\1", geographies), NA)
      county_fips <- ifelse(!is.null(county_fips), sub("^(\\d{1,3})\\d*", "\\1", county_fips), NA)

      
      
      # Get County FIPS list: 
      # tract_fips <- sub('^".*?([^"]+)".*$', '\\1', tract_fips[0]) # [["Census Tracts"]]
      # tract_fips <- sub(".*\\s(\\d{11})[,].*", "\\1", tract_fips)
      county_fips <- sub(".*GEOID = \"([^\"]+)\".*", "\\1", county_fips)
      
      # Get Tract fips:
      # Use sub to extract the 11-digit GEOID value
      tract_fips <- sub('.*GEOID = "(\\d{11})".*', '\\1', paste(address_matches, collapse = ", "))
      

        
      return(list(tract_fips = tract_fips, county_fips = county_fips))
    } else {
      return(list(tract_fips = NA, county_fips = NA))
    }
  } else {
    stop("Error in API request: ", response$status_code)
  }
}

# Example usage
#address <- "_Some one-line address string here_"
#fips_codes <- get_fips_from_address(address)

#print(paste("Tract FIPS:", fips_codes$tract_fips))
#print(paste("County FIPS:", fips_codes$county_fips))



```

# Run function for all rows with lat, lon.
```{r}
# Set initial values for results of geocoding.
fips_w_geocoded_data$tract_fips <- NA
fips_w_geocoded_data$county_fips <- NA


for (i in 1:nrow(fips_w_geocoded_data)) {
  
  # Use raw address: "Patient_Address"
  # Run geocode funct.
  tractcounty_both <- get_fips_from_address(fips_w_geocoded_data$Patient_Address[i])
  tract_result <- tractcounty_both$tract_fips
  county_result <- tractcounty_both$county_fips
  
  cat("Row: ", i, " of ", nrow(fips_w_geocoded_data), "\n")
  cat("Tract FIPS: ", tract_result, "\n")
  cat("County FIPS: ", county_result, "\n")
  
  fips_w_geocoded_data$tract_fips[i] <- as.character(tract_result) # Assign value for column
  fips_w_geocoded_data$county_fips[i] <-as.character(county_result) # Assign value for column
}

# Done
write.csv(fips_w_geocoded_data, "ViaAddresses_FIPS_DONE_allstates_geocoded_patient_data.csv", row.names = FALSE)

```


# Check results of re-coding:
```{r}

fips_w_geocoded_data <- read.csv("ViaAddresses_FIPS_DONE_allstates_geocoded_patient_data.csv")


# Frequency of each "invalid value":
invalid_values <- c(NA, "", ".", "..", "...")

# table(fips_w_geocoded_data$tract_fips) #

# table(fips_w_geocoded_data$county_fips) 


# Now, select ONLY the patients WITH geocodes.
nonmissing_fips_w_geocoded_data <- fips_w_geocoded_data %>%
  filter(
    !(tract_fips %in% invalid_values |
      county_fips %in% invalid_values)
  )


table(nonmissing_fips_w_geocoded_data$tract_fips)
table(nonmissing_fips_w_geocoded_data$county_fips)

nrow(nonmissing_fips_w_geocoded_data) # 4160 correctly geocoded


# HANDLE COLUMN NAMING: 

# First, enforce "nonmissing_fips_w_geocoded_data" columns to be named like the bigger df.
# Copy values from tract_fips to censustract, then remove tract_fips column
nonmissing_fips_w_geocoded_data$censustract <- nonmissing_fips_w_geocoded_data$tract_fips
nonmissing_fips_w_geocoded_data$tract_fips <- NULL
nonmissing_fips_w_geocoded_data$fips <- nonmissing_fips_w_geocoded_data$county_fips
nonmissing_fips_w_geocoded_data$county_fips <- NULL

# Rename columns to match
# Select the common columns between both data frames
common_columns <- intersect(names(non_missing_geocode_df), names(nonmissing_fips_w_geocoded_data))

# Rename corresponding columns in nonmissing_fips_w_geocoded_data to match non_missing_geocode_df
names(nonmissing_fips_w_geocoded_data)[names(nonmissing_fips_w_geocoded_data) == "tract_fips"] <- "censustract"
names(nonmissing_fips_w_geocoded_data)[names(nonmissing_fips_w_geocoded_data) == "county_fips"] <- "fips"


nrow(nonmissing_fips_w_geocoded_data$censustract) # 4160
nrow(non_missing_geocode_df$censustract) # 31804


```



# Merge "nonmissing_fips_w_geocoded_data" (newly completed correctly; 4160) and "non_missing_geocode_df" (orig completed: ~31k)
```{r}
# Originally geocoded ~31k: non_missing_geocode_df
# Geocoded via address ~4k: nonmissing_fips_w_geocoded_data

table(fips_w_geocoded_data$Patient_Home_State)

# Above we read 7k address geocoded back in: some geocoded, some not.
cat("Among ", nrow(fips_w_geocoded_data), ", ", nrow(nonmissing_fips_w_geocoded_data), " were correctly geocoded from address. \n")


# 1. Use our 31k df of nonmissing patients from above called 

# 2. For joining on VALID address-coded rows, Remove rows with NA in county_fips or tract_fips
 # Use: "nonmissing_fips_w_geocoded_data"


# New df
# nonmissing_ViaAddresses_fips_w_geocoded_data$county_fips # County
# nonmissing_ViaAddresses_fips_w_geocoded_data$tract_fips # Tract
# Orig big df + newly geocoded non-missing df:
# non_missing_geocode_df$censustract # Tract
# non_missing_geocode_df$fips # County
# Find non-overlapping column names between two data frames
#unique_cols <- list(
#  only_in_df1 = setdiff(names(non_missing_geocode_df), names(nonmissing_ViaAddresses_fips_w_geocoded_data)),
#  only_in_df2 = setdiff(names(nonmissing_ViaAddresses_fips_w_geocoded_data), names(non_missing_geocode_df))
#)
#unique_cols



# ###########################
# Originally geocoded ~31k: non_missing_geocode_df
# Geocoded via address ~4k: nonmissing_fips_w_geocoded_data


# Keep only the common columns in both dataframes 
non_missing_geocode_df <- non_missing_geocode_df[, common_columns]
nonmissing_fips_w_geocoded_data <- nonmissing_fips_w_geocoded_data[, common_columns]

# Stack the dataframes
# final_merged_df <- rbind(non_missing_geocode_df, nonmissing_fips_w_geocoded_data)


# Merge dataframes, keeping all unique Encounter values
# Merge dataframes without suffixes
final_merged_df <- merge(non_missing_geocode_df, nonmissing_fips_w_geocoded_data, 
                         by = "Encounter", 
                         all = TRUE,
                         suffixes = c("", "_y"))

# Identify columns to coalesce (excluding Encounter)
cols_to_coalesce <- names(final_merged_df)[grep("_y$", names(final_merged_df))]
orig_cols <- sub("_y$", "", cols_to_coalesce)

# Coalesce columns
for (i in seq_along(orig_cols)) {
  final_merged_df[[orig_cols[i]]] <- ifelse(is.na(final_merged_df[[orig_cols[i]]]), 
                                             final_merged_df[[cols_to_coalesce[i]]], 
                                             final_merged_df[[orig_cols[i]]])
  final_merged_df[[cols_to_coalesce[i]]] <- NULL
}

nrow(final_merged_df)



# Optional: Check for any remaining duplicates
sum(duplicated(final_merged_df$Encounter))

# View the result
final_merged_df


cat("Total final patients geocoded: ", nrow(final_merged_df))

final_merged_df[final_merged_df$censustract == "..", ]
final_merged_df[final_merged_df$fips == "..", ]

# Among  7262 ,  4160  were correctly geocoded from address. 
# Total final patients geocoded:  35964



```




# Compare: final_merged_df (35964) vs. nonmissing_fips_w_geocoded_data (4160)


```{r}
cat("To begin, we had ", nrow(filtered_df), "patients meeting inclusion criteria (Black/White, have endpoints, adults). But not all have been geocoded from their address string. \n")
cat("Originally, among these we had ", nrow(non_missing_geocode_df), " correctly geocoded.\n")
cat("Among ", nrow(fips_w_geocoded_data), ", ", nrow(nonmissing_fips_w_geocoded_data), "extra were correctly geocoded from address. \n")
cat(">>>> Finally, we have ", nrow(final_merged_df), " correctly geocoded.\n")


#fips_w_geocoded_data$Encounter # Double. 7262
#nonmissing_fips_w_geocoded_data$Encounter # Double. 4160

# To find patients who failed at all geocoding: Take all values in fips_w_geocoded_data but NOT in nonmissing_fips_w_geocoded_data
#fips_w_geocoded_data
#nonmissing_fips_w_geocoded_data

# Find rows in fips_w_geocoded_data not in nonmissing_fips_w_geocoded_data
failed_second_geocoding <- fips_w_geocoded_data[!fips_w_geocoded_data$Encounter %in% nonmissing_fips_w_geocoded_data$Encounter, ]
cat(">>>> We can explore the patients which were never able to be geocoded: ", nrow(failed_second_geocoding), ".\n")


# Saved for analysis in table1.Rmd
filename_failed_second_geocoding <- paste0("nonmissing_geocodedf_n-", nrow(failed_second_geocoding), ".csv")
write.csv(failed_second_geocoding, filename_failed_second_geocoding)
cat("Saved for analysis in table1.Rmd for ", nrow(failed_second_geocoding), "patients. [Filename: ", filename_failed_second_geocoding, "]\n")

```

















# Remove PHI columns
```{r}

remove_cols <- c("patient_name", "Patient", "Match_Address", "Patient_Address",
                 "Patient_Birth_Date", "Patient_Home_Street_Address", "Patient_Home_City", "Patient_Home_State", "dob",
                 "Patient_Death_Day")
                 

final_merged_df = final_merged_df[,!(names(final_merged_df) %in% remove_cols)]

nrow(final_merged_df)
```

# Fill in corresponding "STATE" column for newly geocoded patients.
```{r}

# Ensure 'censustract' column is a character type to allow substring operations
final_merged_df$censustract <- as.character(final_merged_df$censustract)
final_merged_df$censustract <- str_pad(final_merged_df$censustract, width = 11, pad = "0")
final_merged_df$fips <- str_pad(final_merged_df$fips, width = 5, pad = "0")


# Extract the first two digits from 'censustract' and assign to 'statenum' column
final_merged_df$statenum <- substr(final_merged_df$censustract, 1, 2)

table(final_merged_df$statenum, useNA = "ifany")

```

# Finally, write final result
```{r}

write.csv(final_merged_df, "ML4H_Circ_Nonmissing_Cohort.csv", row.names = FALSE)



# For quick checks.
ck_df <- read.csv("ML4H_Circ_Nonmissing_Cohort.csv")

ck_df[ck_df$Encounter == 8033890056, ]$censustract

```







