---
title: "table1_Circ_2024"
output: html_notebook
---

# Load libraries
```{r}
library(caret)
library(data.table)
library(dplyr)
library(sets)
library(scales)
library(tidyr)
library(stringr)
library(tibble)
library(tidygeocoder)
library(stringi)
library(ggplot2)


library(psych)
library(tidyverse)
library(irr)


library(readr) # For read_tsv
```


```{r}
## tableone package itself
library(tableone)
library(kableExtra)

# Make strata column: white, black, other (1,2,3)
# make using cols:
# Drop "Race" == "other"


input_ahrq_patient_feats <- read.csv(file="../data/Circ_final_allstates_modelinput.csv") # DO NOT USE: Circ_allstates_includemissing_modelinput_DF1.csv

nrow(input_ahrq_patient_feats[input_ahrq_patient_feats$white == 1, ])


#Print missing rates for each of the AHRQ SDOHD vars && also SDOH *limited* vars.
# Print for AHRQ SDOHD vars:
sum(is.na(input_ahrq_patient_feats$ADI_national_rank))/nrow(input_ahrq_patient_feats) # 4% miss.

# Print for SODH *limited vars*

# To csv:

sum(is.na(input_ahrq_patient_feats$ADI_state_rank))/nrow(input_ahrq_patient_feats) # 18% miss


table(input_ahrq_patient_feats$ADI_state_rank, useNA = "ifany")
```


```{r}
input_feats <- input_ahrq_patient_feats
trans_input_feats <- input_feats[! (input_feats$Race == "other"), ]


trans_input_feats <- trans_input_feats <- transform(
  input_feats, "Race"= ifelse(white==1, "White", ifelse(black==1, "Black", NA)))
  

# Make insurance column: "privinsurance","Medicare", "Medicaid",
trans_input_feats <- data_frame <- transform(
  trans_input_feats, "Insurance_type"= ifelse(privinsurance==1, "Private", ifelse(Medicare==1, "Medicare", ifelse(Medicaid==1, "Medicaid", "Not recorded"))))
  

# Make HF type: hfsystolic, hfdiastolic, hfother
trans_input_feats <- data_frame <- transform(
  trans_input_feats, "HF_classification"= ifelse(hfsystolic==1, "Systolic (HFrEF)", ifelse(hfdiastolic==1, "Diastolic (HFpEF)", ifelse(hfother==1, "Other", "Not recorded"))))



# vars to summarize (i.e demographics)
# htn: Hypertension
# diabetes: _not available_
# cad: Coronary Artery Disease
# ckd: Chronic Kidney Disease
# sdi: Social deprivation index score
# ADI_national_rank: 
# ADI_state_rank: 
# svi: i.e. RPL_THEMES
# cci: Charlson Comorbidity Index
# Systolic_BP: Systolic blood pressure, mm Hg
# Heart_Rate: Heart Rate, beats per minute
# Resp: Respiratory rate, breaths per minute

# Rename feats not needing spaces: 
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_current"] ="Age, years"
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_Q1"] ="Age, Q1"
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_Q2"] ="Age, Q2"
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_Q3"] ="Age, Q3"
colnames(trans_input_feats)[colnames(trans_input_feats) == "age_Q4"] ="Age, Q4"

colnames(trans_input_feats)[colnames(trans_input_feats) == "female"] = "Female"
colnames(trans_input_feats)[colnames(trans_input_feats) == "male"] = "Male"

colnames(trans_input_feats)[colnames(trans_input_feats) == "black"] = "Black"
colnames(trans_input_feats)[colnames(trans_input_feats) == "white"] = "White"

colnames(trans_input_feats)[colnames(trans_input_feats) == "death30bin"] = "30-day Mortality"

colnames(trans_input_feats)[colnames(trans_input_feats) == "htn"] = "Hypertension"
colnames(trans_input_feats)[colnames(trans_input_feats) == "cad"] = "Coronary Artery Disease"
colnames(trans_input_feats)[colnames(trans_input_feats) == "ckd"] = "Chronic Kidney Disease"
colnames(trans_input_feats)[colnames(trans_input_feats) == "cci"] = "Charlson Comorbidity Index"
colnames(trans_input_feats)[colnames(trans_input_feats) == "Systolic_BP"] = "Systolic blood pressure, mm Hg"
colnames(trans_input_feats)[colnames(trans_input_feats) == "Heart_Rate"] = "Heart Rate, beats per minute"
colnames(trans_input_feats)[colnames(trans_input_feats) == "Resp"] = "Respiratory rate, breaths per minute"

colnames(trans_input_feats)[colnames(trans_input_feats) == "Diastolic_BP"] = "Diastolic Blood Pressure (20-200), mmHg"
colnames(trans_input_feats)[colnames(trans_input_feats) == "BNP"] = "B-type natriuretic peptide (BNP), pg/mL"
colnames(trans_input_feats)[colnames(trans_input_feats) == "BUN"] = "Blood urea nitrogen (BUN) test, mg/dL"
colnames(trans_input_feats)[colnames(trans_input_feats) == "ALB"] = "Albumin"
colnames(trans_input_feats)[colnames(trans_input_feats) == "Hb"] = "Hemoglobin, g/dL"
colnames(trans_input_feats)[colnames(trans_input_feats) == "HbA1C"] = "Hemoglobin A1C (0-20), %"
colnames(trans_input_feats)[colnames(trans_input_feats) == "K"] = "Potassium, mEq/L"
colnames(trans_input_feats)[colnames(trans_input_feats) == "Creat"] = "Serum Creatinine, mg/dL"
colnames(trans_input_feats)[colnames(trans_input_feats) == "Na"] = "Sodium, mEq/L"


colnames(trans_input_feats)[colnames(trans_input_feats) == "sdi"] = "Social deprivation index (percentile)"
colnames(trans_input_feats)[colnames(trans_input_feats) == "ADI_national_rank"] = "Area deprivation index, national rank (percentile)"
colnames(trans_input_feats)[colnames(trans_input_feats) == "ADI_state_rank"] = "Area deprivation index, state rank (percentile)"
colnames(trans_input_feats)[colnames(trans_input_feats) == "RPL_THEMES"] = "Social vulnerability index score (composite score)"

# New feats:
#colnames(trans_input_feats)[colnames(trans_input_feats) == "readmit30bin"] = "30-day Binary HF Readmission"
#colnames(trans_input_feats)[colnames(trans_input_feats) == "death30bin"] = "30-day Binary HF Death"

# Referring specialties: 
colnames(trans_input_feats)[colnames(trans_input_feats) == "specialcvd"] = "CVD Specialist" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "specialint"] = "Int. Med. Specialist"
colnames(trans_input_feats)[colnames(trans_input_feats) == "specialeme"] = "EME Specialist"
colnames(trans_input_feats)[colnames(trans_input_feats) == "specialoth"] = "Other Specialist"

# Location treated: 
colnames(trans_input_feats)[colnames(trans_input_feats) == "loceuhfloor"] = "UH Floor" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "loceclhicu"] = "CLH ICU"
colnames(trans_input_feats)[colnames(trans_input_feats) == "loceclhfloor"] = "CLH Floor"
colnames(trans_input_feats)[colnames(trans_input_feats) == "loceuhicu"] = "UH ICU"

# SDOH feats (comp -- other Circulation HF paper)

colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_ltfpl100"] = "percnt_ltfpl100" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_singlparntfly"] = "percnt_singlparntfly" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_dropout"] = "percnt_dropout" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_rentoccup"] = "percnt_rentoccup" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_crowding"] = "percnt_crowding" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_nonemp"] = "percnt_nonemp" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_unemp"] = "percnt_unemp" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_highneeds"] = "percnt_highneeds" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_hispanic"] = "percnt_hispanic" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_frgnborn"] = "percnt_frgnborn" 
colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_black"] = "percnt_black" 
# sdi
# ADI_state_rank
# ADI_national_rank


# Rename feats with spaces: 
colnames(trans_input_feats)[colnames(trans_input_feats) == "Insurance_type"] = "Insurance type"
colnames(trans_input_feats)[colnames(trans_input_feats) == "HF_classification"] = "HF classification"




# Length of stay, days: LOScheck
#myVars <- c("Age, years", "female", "Insurance type", "HF classification", "htn", "cad", "ckd", "cci", 
#            "Systolic_BP", "Heart_Rate", "Resp", 
#            "LOScheck", "sdi", "ADI_national_rank", "ADI_state_rank", "RPL_THEMES")

myVars <- c("Age, years","Age, Q1","Age, Q2","Age, Q3","Age, Q4", "Male", "Female", "Black", "White",
            "30-day Mortality",
            "CVD Specialist", "Int. Med. Specialist", "EME Specialist", "Other Specialist",
            "UH Floor", "CLH ICU", "CLH Floor", "UH ICU",
            
            "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease", 
            "Charlson Comorbidity Index", 
            "Systolic blood pressure, mm Hg", "Heart Rate, beats per minute", "Respiratory rate, breaths per minute", 
                        "Diastolic Blood Pressure (20-200), mmHg",
            "B-type natriuretic peptide (BNP), pg/mL",
            "Blood urea nitrogen (BUN) test, mg/dL",
            "Albumin",
            "Hemoglobin, g/dL",
            "Hemoglobin A1C (0-20), %",
            "Potassium, mEq/L",
            "Serum Creatinine, mg/dL",
            "Sodium, mEq/L",

            "LOScheck", "Social deprivation index (percentile)", "Area deprivation index, national rank (percentile)", "Area deprivation index, state rank (percentile)", "Social vulnerability index score (composite score)",
            "percnt_ltfpl100", "percnt_singlparntfly", "percnt_dropout", "percnt_rentoccup", "percnt_crowding", "percnt_nonemp", "percnt_unemp", "percnt_highneeds", "percnt_hispanic", "percnt_frgnborn", "percnt_black"
            
            )


# Vector of categorical variables that need transformation
catVars <- c("Women", "30-day Mortality", "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease")

tab1 <- CreateTableOne(vars = myVars, strata = "readmit30bin", data = trans_input_feats, addOverall = TRUE) 
# For race strata instead of outcome strata: strata = "Race"
# For p-values: test=TRUE, smd = TRUE

# typeof(tab1["test"])



# If you just say nonnormal = TRUE, all variables are summarized the “nonnormal” way.
#print(tab1, formatOptions = list(big.mark = ","))
# kable(p, format = "latex")
p <- print(tab1, printToggle = FALSE, noSpaces = TRUE)
#p <- p[,1:3] # Remove "test" column.

kable(p, format = "markdown", booktabs = TRUE)
```



# For tableone missingness:
```{r}
summary(tab1)
```


# Print with cont, categorical proper:

```{r}
# Vector of continuous variables
contVars <- c("Age, years", 
              "Systolic blood pressure, mm Hg", 
              "Heart Rate, beats per minute", 
              "Respiratory rate, breaths per minute",
              "Diastolic Blood Pressure (20-200), mmHg",
              "B-type natriuretic peptide (BNP), pg/mL",
              "Blood urea nitrogen (BUN) test, mg/dL",
              "Albumin",
              "Hemoglobin, g/dL",
              "Hemoglobin A1C (0-20), %",
              "Potassium, mEq/L",
              "Serum Creatinine, mg/dL",
              "Sodium, mEq/L",
              "Social deprivation index (percentile)", 
              "Area deprivation index, national rank (percentile)", 
              "Area deprivation index, state rank (percentile)", 
              "Social vulnerability index score (composite score)")

# Vector of categorical variables
catVars <- c("Male", "Female", "Black", "White", 
             "30-day Mortality", 
             "Insurance type", 
             "HF classification", 
             "Hypertension", 
             "Coronary Artery Disease", 
             "Chronic Kidney Disease",
             "CVD Specialist", 
             "Int. Med. Specialist", 
             "EME Specialist", 
             "Other Specialist",
             "UH Floor", 
             "CLH ICU", 
             "CLH Floor", 
             "UH ICU")

# Convert these columns to factors
for (col in catVars) {
  trans_input_feats[[col]] <- as.factor(trans_input_feats[[col]])
}

# Create table with custom summary functions
# Create the table with more explicit parameters
tab1 <- CreateTableOne(vars = c(contVars, catVars), 
                       strata = "readmit30bin", 
                       data = trans_input_feats, 
                       addOverall = TRUE,
                       test = TRUE,
                       smd = TRUE)

# Print the table with simplified formatting
p <- print(tab1, 
           printToggle = FALSE, 
           noSpaces = TRUE)

# Convert to markdown table
kable(p, format = "markdown", booktabs = TRUE)
#summary(tab1)

```









































Total: Circ_final_allstates_modelinput.csv
- df: trans_input_feats
Unsuccessful geocoded: nonmissing_geocodedf_n-3102.csv
- df: nonmissing_df_input_ahrq_patient_feats


# Read in: missing patients.
... then Compare non-missing patients (above) vs. missing patients who were dropped:
```{r}

# 1. Prepare nonmissing df:
nonmissing_df_input_ahrq_patient_feats <- read.csv(file="../data/nonmissing_geocodedf_n-3102.csv")


# CREATE DEMOGRAPHIC VARS: 

nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(baselineyear = as.numeric(baselineyear), na.rm=FALSE)

############ Prepare subgroup flags:
# Subgroups to eval on: 
# 1. Race
# white <int>: 0/1
# black <int>: 0/1
# other <int>: 0/1
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(blackorwhite = if_else(black == 1 | white == 1, 1, 0))

# 2. Sex (or binary gender)
# female <int>: 0/1
# ADD this field: male <int>: 0/1
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(male = if_else(female == 0, 1, 0))

# 3. The 6 subgroup combinations: 
# c(white_male, black_male, other_male, white_female, black_female, other_female)

# Note that all 6 are mutually exclusive here.
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(white_male = if_else(white == 1 & female == 0, 1, 0))
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(black_male = if_else((black == 1 & female == 0), 1, 0))
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(other_male = if_else(other == 1 & female == 0, 1, 0))
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(white_female = if_else(white == 1 & female == 1, 1, 0))
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(black_female = if_else(black == 1 & female == 1, 1, 0))
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(other_female = if_else(other == 1 & female == 1, 1, 0))

#######################
# Hypertension
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(no_htn = if_else(htn == 0, 1, 0))

#######################
# Coronary Artery Disease
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(no_cad = if_else(cad == 0, 1, 0))

#######################
# Chronic Kidney Disease
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats %>%
  mutate(no_ckd = if_else(ckd == 0, 1, 0))



# Make insurance column: "privinsurance","Medicare", "Medicaid",
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats <- transform(
  nonmissing_df_input_ahrq_patient_feats, "Insurance_type"= ifelse(privinsurance==1, "Private", ifelse(Medicare==1, "Medicare", ifelse(Medicaid==1, "Medicaid", "Not recorded"))))
  

# Make HF type: hfsystolic, hfdiastolic, hfother
nonmissing_df_input_ahrq_patient_feats <- nonmissing_df_input_ahrq_patient_feats <- transform(
  nonmissing_df_input_ahrq_patient_feats, "HF_classification"= ifelse(hfsystolic==1, "Systolic (HFrEF)", ifelse(hfdiastolic==1, "Diastolic (HFpEF)", ifelse(hfother==1, "Other", "Not recorded"))))



########################
# Then, convert missing/invalid values in "nonmissing_df_input_ahrq_patient_feats" to NAs:

# Function to convert non-numeric to NA
convert_nonnumeric_to_na <- function(df, columns) {
  # Loop through specified columns
  for(col in columns) {
    if(col %in% names(df)) {
      # Convert column to numeric, non-numeric values will become NA
      df[[col]] <- suppressWarnings(as.numeric(df[[col]]))
    }
  }
  return(df)
}

# Define deprivation columns
deprivation_cols_total <- c("percnt_ltfpl100", "percnt_singlparntfly", "percnt_dropout", 
                           "percnt_hhnocar", "percnt_rentoccup", "percnt_crowding", 
                           "percnt_nonemp", "percnt_unemp", "percnt_highneeds", 
                           "percnt_hispanic", "percnt_frgnborn", "percnt_black", 
                           "sdi", "ADI_state_rank", "ADI_national_rank")

# Apply the conversion to nonmissing_df_input_ahrq_patient_feats
nonmissing_df_input_ahrq_patient_feats <- convert_nonnumeric_to_na(nonmissing_df_input_ahrq_patient_feats, deprivation_cols_total)

```

# Table labels: Prepare missing patients for tableone:
```{r}

# Rename feats not needing spaces: 
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "age_current"] ="Age, years"

colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "female"] = "Female"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "male"] = "Male"

colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "black"] = "Black"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "white"] = "White"

colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "death30bin"] = "30-day Mortality"

colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "htn"] = "Hypertension"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "cad"] = "Coronary Artery Disease"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "ckd"] = "Chronic Kidney Disease"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "cci"] = "Charlson Comorbidity Index"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "Systolic_BP"] = "Systolic blood pressure, mm Hg"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "Heart_Rate"] = "Heart Rate, beats per minute"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "Resp"] = "Respiratory rate, breaths per minute"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "Diastolic_BP"] = "Diastolic Blood Pressure (20-200), mmHg"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "BNP"] = "B-type natriuretic peptide (BNP), pg/mL"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "BUN"] = "Blood urea nitrogen (BUN) test, mg/dL"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "ALB"] = "Albumin"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "Hb"] = "Hemoglobin, g/dL"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "HbA1C"] = "Hemoglobin A1C (0-20), %"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "K"] = "Potassium, mEq/L"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "Creat"] = "Serum Creatinine, mg/dL"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "Na"] = "Sodium, mEq/L"


#colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "sdi"] = "Social deprivation index (percentile)"
#colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "ADI_national_rank"] = "Area deprivation index, national rank (percentile)"
#colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "ADI_state_rank"] = "Area deprivation index, state rank (percentile)"
#colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "RPL_THEMES"] = "Social vulnerability index score (composite score)"

# New feats:
#colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(trans_input_feats) == "readmit30bin"] = "30-day Binary HF Readmission"
#colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(trans_input_feats) == "death30bin"] = "30-day Binary HF Death"

# Referring specialties: 
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "specialcvd"] = "CVD Specialist" 
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "specialint"] = "Int. Med. Specialist"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "specialeme"] = "EME Specialist"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "specialoth"] = "Other Specialist"

# Location treated: 
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "loceuhfloor"] = "UH Floor" 
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "loceclhicu"] = "CLH ICU"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "loceclhfloor"] = "CLH Floor"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "loceuhicu"] = "UH ICU"

# SDOH feats (comp -- other Circulation HF paper)
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_ltfpl100"] = "percnt_ltfpl100" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_singlparntfly"] = "percnt_singlparntfly" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_dropout"] = "percnt_dropout" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_rentoccup"] = "percnt_rentoccup" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_crowding"] = "percnt_crowding" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_nonemp"] = "percnt_nonemp" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_unemp"] = "percnt_unemp" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_highneeds"] = "percnt_highneeds" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_hispanic"] = "percnt_hispanic" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_frgnborn"] = "percnt_frgnborn" 
#colnames(trans_input_feats)[colnames(trans_input_feats) == "percnt_black"] = "percnt_black" 
# sdi
# ADI_state_rank
# ADI_national_rank


# Rename feats with spaces: 
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "Insurance_type"] = "Insurance type"
colnames(nonmissing_df_input_ahrq_patient_feats)[colnames(nonmissing_df_input_ahrq_patient_feats) == "HF_classification"] = "HF classification"




# Length of stay, days: LOScheck
#myVars <- c("Age, years", "female", "Insurance type", "HF classification", "htn", "cad", "ckd", "cci", 
#            "Systolic_BP", "Heart_Rate", "Resp", 
#            "LOScheck", "sdi", "ADI_national_rank", "ADI_state_rank", "RPL_THEMES")

myVars <- c("Age, years", "Male", "Female", "Black", "White",
            "30-day Mortality",
            "CVD Specialist", "Int. Med. Specialist", "EME Specialist", "Other Specialist",
            "UH Floor", "CLH ICU", "CLH Floor", "UH ICU",
            
            "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease", 
            "Charlson Comorbidity Index", 
            "Systolic blood pressure, mm Hg", "Heart Rate, beats per minute", "Respiratory rate, breaths per minute", 
            "Diastolic Blood Pressure (20-200), mmHg",
            "B-type natriuretic peptide (BNP), pg/mL",
            "Blood urea nitrogen (BUN) test, mg/dL",
            "Albumin",
            "Hemoglobin, g/dL",
            "Hemoglobin A1C (0-20), %",
            "Potassium, mEq/L",
            "Serum Creatinine, mg/dL",
            "Sodium, mEq/L",
            
            "LOScheck"
            )


## Vector of categorical variables that need transformation
catVars <- c("Women", "30-day Mortality", "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease")

tab1_miss <- CreateTableOne(vars = myVars, strata = "readmit30bin", data = nonmissing_df_input_ahrq_patient_feats, addOverall = TRUE) 
# For race strata instead of outcome strata: strata = "Race"
# For p-values: test=TRUE, smd = TRUE

# typeof(tab1["test"])



# If you just say nonnormal = TRUE, all variables are summarized the “nonnormal” way.
#print(tab1, formatOptions = list(big.mark = ","))
# kable(p, format = "latex")
p_miss <- print(tab1_miss, printToggle = FALSE, noSpaces = TRUE)
#p <- p[,1:3] # Remove "test" column.

kable(p_miss, format = "markdown", booktabs = TRUE)
```
# Summary: for missing patients.
```{r}
summary(tab1_miss)

```


```{r}

# 1. State names: Invalid values for states: c("NOT", "AE", NA). "VI" is virgin islands. "PR" is puerto rico
# GA (6470)
# table(nonmissing_df_input_ahrq_patient_feats$Patient_Home_State)
# 2. County names: all full.
# table(nonmissing_df_input_ahrq_patient_feats$Patient_Home_County)


```



# Compare total vs. unsuccessful geocoded: 
```{r}
combined_myVars <- c("Age, years", "Male", "Female", "Black", "White",
            "30-day Mortality",
            "CVD Specialist", "Int. Med. Specialist", "EME Specialist", "Other Specialist",
            "UH Floor", "CLH ICU", "CLH Floor", "UH ICU",
            
            "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease", 
            "Charlson Comorbidity Index", 
            "Systolic blood pressure, mm Hg", "Heart Rate, beats per minute", "Respiratory rate, breaths per minute", 
            "Diastolic Blood Pressure (20-200), mmHg",
            "B-type natriuretic peptide (BNP), pg/mL",
            "Blood urea nitrogen (BUN) test, mg/dL",
            "Albumin",
            "Hemoglobin, g/dL",
            "Hemoglobin A1C (0-20), %",
            "Potassium, mEq/L",
            "Serum Creatinine, mg/dL",
            "Sodium, mEq/L",
            
            "LOScheck"
            )


## Vector of categorical variables that need transformation
combined_catVars <- c("Male", "Female", "Black", "White", "30-day Mortality", "CVD Specialist", "Int. Med. Specialist", "EME Specialist", "Other Specialist", "UH Floor", "CLH ICU", "CLH Floor", "UH ICU", "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease")

########################
tab1 <- CreateTableOne(vars = combined_myVars, strata = "readmit30bin", data = trans_input_feats, addOverall = TRUE) 
# For race strata instead of outcome strata: strata = "Race"
# For p-values: test=TRUE, smd = TRUE

# If you just say nonnormal = TRUE, all variables are summarized the “nonnormal” way.
#print(tab1, formatOptions = list(big.mark = ","))
# kable(p, format = "latex")
p <- print(tab1, printToggle = FALSE, noSpaces = TRUE)
kable(p, format = "markdown", booktabs = TRUE)


########################
tab1_miss <- CreateTableOne(vars = combined_myVars, strata = "readmit30bin", data = nonmissing_df_input_ahrq_patient_feats, addOverall = TRUE) 
# For race strata instead of outcome strata: strata = "Race"


# If you just say nonnormal = TRUE, all variables are summarized the “nonnormal” way.
#print(tab1, formatOptions = list(big.mark = ","))
# kable(p, format = "latex")
p_miss <- print(tab1_miss, printToggle = FALSE, noSpaces = TRUE)

kable(p_miss, format = "markdown", booktabs = TRUE)

```

# Comparison table output:
```{r}
library(tableone)
library(knitr)
library(dplyr)

# Define variables
combined_myVars <- c("Age, years", "Male", "Female", "Black", "White",
            "30-day Mortality",
            "CVD Specialist", "Int. Med. Specialist", "EME Specialist", "Other Specialist",
            "UH Floor", "CLH ICU", "CLH Floor", "UH ICU",
            
            "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease", 
            "Charlson Comorbidity Index", 
            "Systolic blood pressure, mm Hg", "Heart Rate, beats per minute", "Respiratory rate, breaths per minute", 
            "Diastolic Blood Pressure (20-200), mmHg",
            "B-type natriuretic peptide (BNP), pg/mL",
            "Blood urea nitrogen (BUN) test, mg/dL",
            "Albumin",
            "Hemoglobin, g/dL",
            "Hemoglobin A1C (0-20), %",
            "Potassium, mEq/L",
            "Serum Creatinine, mg/dL",
            "Sodium, mEq/L",
            
            "LOScheck"
)

# Vector of categorical variables that need transformation
combined_catVars <- c("Male", "Female", "Black", "White", "30-day Mortality", 
                      "CVD Specialist", "Int. Med. Specialist", "EME Specialist", 
                      "Other Specialist", "UH Floor", "CLH ICU", "CLH Floor", "UH ICU", 
                      "Insurance type", "HF classification", "Hypertension", 
                      "Coronary Artery Disease", "Chronic Kidney Disease")



library(tableone)
library(knitr)
library(dplyr)

# First, identify the common columns between the two dataframes
common_columns <- intersect(names(trans_input_feats), names(nonmissing_df_input_ahrq_patient_feats))

# Subset both dataframes to only these common columns
trans_input_feats_subset <- trans_input_feats[, common_columns]
nonmissing_df_input_ahrq_patient_feats_subset <- nonmissing_df_input_ahrq_patient_feats[, common_columns]

# Add the cohort identifier
trans_input_feats_subset$cohort <- "Geocoded"
nonmissing_df_input_ahrq_patient_feats_subset$cohort <- "Non-Geocoded"

# Now combine the dataframes
combined_df <- rbind(trans_input_feats_subset, nonmissing_df_input_ahrq_patient_feats_subset)

# Modify the variable lists to only include columns in the combined dataframe
combined_myVars <- intersect(combined_myVars, names(combined_df))
combined_catVars <- intersect(combined_catVars, names(combined_df))

# Create a comprehensive comparison table
comprehensive_table <- CreateTableOne(
    vars = combined_myVars, 
    strata = "cohort", 
    data = combined_df, 
    addOverall = TRUE,
    test = TRUE,  # Enable statistical testing
    smd = TRUE    # Standardized Mean Difference
)

# Print the table
print_comprehensive <- print(comprehensive_table, 
                              printToggle = FALSE, 
                              noSpaces = TRUE, 
                              showAllLevels = TRUE)

# Generate a markdown table
kable(print_comprehensive, format = "markdown", booktabs = TRUE)

# Optional: Save detailed statistical comparison to a CSV
write.csv(print_comprehensive, "cohort_comparison.csv", row.names = TRUE)

# Additional detailed statistical tests output
print(comprehensive_table)

# Print table as tab-separated text
print_comprehensive_tab <- print(comprehensive_table, 
                                  printToggle = FALSE, 
                                  noSpaces = TRUE, 
                                  showAllLevels = TRUE)

# Also print to console for immediate viewing
print(print_comprehensive_tab)
```

# Unsuccessful geocoded vs. total -- with cont, cat.
```{r}


# Define variables
combined_myVars <- c("Age, years", "Male", "Female", "Black", "White",
            "30-day Mortality",
            "CVD Specialist", "Int. Med. Specialist", "EME Specialist", "Other Specialist",
            "UH Floor", "CLH ICU", "CLH Floor", "UH ICU",
            
            "Insurance type", "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease", 
            "Charlson Comorbidity Index", 
            "Systolic blood pressure, mm Hg", "Heart Rate, beats per minute", "Respiratory rate, breaths per minute", 
            "Diastolic Blood Pressure (20-200), mmHg",
            "B-type natriuretic peptide (BNP), pg/mL",
            "Blood urea nitrogen (BUN) test, mg/dL",
            "Albumin",
            "Hemoglobin, g/dL",
            "Hemoglobin A1C (0-20), %",
            "Potassium, mEq/L",
            "Serum Creatinine, mg/dL",
            "Sodium, mEq/L",
            
            "LOScheck"
)

# Vector of categorical variables that need transformation
combined_catVars <- c("Male", "Female", "Black", "White", "30-day Mortality", 
                      "CVD Specialist", "Int. Med. Specialist", "EME Specialist", 
                      "Other Specialist", "UH Floor", "CLH ICU", "CLH Floor", "UH ICU", 
                      "Insurance type", "HF classification", "Hypertension", 
                      "Coronary Artery Disease", "Chronic Kidney Disease")

combined_contVars <- c("Age, years", 
              "Systolic blood pressure, mm Hg", 
              "Heart Rate, beats per minute", 
              "Respiratory rate, breaths per minute",
              "Diastolic Blood Pressure (20-200), mmHg",
              "B-type natriuretic peptide (BNP), pg/mL",
              "Blood urea nitrogen (BUN) test, mg/dL",
              "Albumin",
              "Hemoglobin, g/dL",
              "Hemoglobin A1C (0-20), %",
              "Potassium, mEq/L",
              "Serum Creatinine, mg/dL",
              "Sodium, mEq/L",
              "LOScheck")

# First, identify the common columns between the two dataframes
common_columns <- intersect(names(trans_input_feats), names(nonmissing_df_input_ahrq_patient_feats))

# Subset both dataframes to only these common columns
trans_input_feats_subset <- trans_input_feats[, common_columns]
nonmissing_df_input_ahrq_patient_feats_subset <- nonmissing_df_input_ahrq_patient_feats[, common_columns]

# Add the cohort identifier
trans_input_feats_subset$cohort <- "Geocoded"
nonmissing_df_input_ahrq_patient_feats_subset$cohort <- "Non-Geocoded"

# Now combine the dataframes
combined_df <- rbind(trans_input_feats_subset, nonmissing_df_input_ahrq_patient_feats_subset)

# Modify the variable lists to only include columns in the combined dataframe
combined_myVars <- intersect(combined_myVars, names(combined_df))
combined_catVars <- intersect(combined_catVars, names(combined_df))


##############

# Convert these columns to factors
for (col in combined_catVars) {
  combined_df[[col]] <- as.factor(combined_df[[col]])
}


# Create a comprehensive comparison table
comprehensive_table <- CreateTableOne(
    vars = c(combined_contVars, combined_catVars), 
    strata = "cohort", 
    data = combined_df, 
    addOverall = TRUE,
    test = TRUE,  # Enable statistical testing
    smd = TRUE    # Standardized Mean Difference
)


# Print the table
print_comprehensive <- print(comprehensive_table, 
                              printToggle = FALSE, 
                              noSpaces = TRUE, 
                              showAllLevels = TRUE)

# Generate a markdown table
kable(print_comprehensive, format = "markdown", booktabs = TRUE)

# Additional detailed statistical tests output
print(comprehensive_table)

# Print table as tab-separated text
print_comprehensive_tab <- print(comprehensive_table, 
                                  printToggle = FALSE, 
                                  noSpaces = TRUE, 
                                  showAllLevels = TRUE)

# Also print to console for immediate viewing
print(print_comprehensive_tab)

```

