---
title: "Heart Failure Performance & Fairness Analysis"
author: "Chase Fensore"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load all required libraries once
library(caret)
library(data.table)
library(dplyr)
library(tidyr)
library(stringr)
library(stringi)
library(ggplot2)
library(knitr)
library(RColorBrewer)
library(readr)
```

## 1. Global Configurations & Feature Sets
Define all feature sets globally so they can be reused across chunks without repetition.

```{r config-feature-sets}
model_order <- c("logr", "rf", "xgb")

county_feat_sets <- c("M1", "M2", "M3_county_DF1_nm", "M4_DF1_nm", "M5", "M6_total_ahrq_cty_DF1_nm")

county_only_feats <- c(
  "M6_total_ahrq_cty_DF1_nm", 
  "demo_DF1_nm_county_AHRQ_domain1", "demo_DF1_nm_county_AHRQ_domain2",
  "demo_DF1_nm_county_AHRQ_domain3", "demo_DF1_nm_county_AHRQ_domain4",
  "demo_DF1_nm_county_AHRQ_domain5"
)

tract_only_feats <- c(
  "M6_total_ahrq_trct_DF1_nm",
  "demo_DF1_nm_tract_AHRQ_domain1", "demo_DF1_nm_tract_AHRQ_domain2",
  "demo_DF1_nm_tract_AHRQ_domain3", "demo_DF1_nm_tract_AHRQ_domain4",
  "demo_DF1_nm_tract_AHRQ_domain5"
)

# Combined SDOH sets for filtering
sdoh_feat_sets <- unique(c(county_only_feats, tract_only_feats))

# Long names for plotting mapping
total_feats_used <- c(
  "M1", "M2", "M3_county_DF1_nm", "M4_DF1_nm_demo", "M5_and_demo", "M6_total_ahrq_cty_DF1_nm",
  "demo_DF1_nm_county_AHRQ_domain1", "demo_DF1_nm_county_AHRQ_domain2", "demo_DF1_nm_county_AHRQ_domain3",
  "demo_DF1_nm_county_AHRQ_domain4", "demo_DF1_nm_county_AHRQ_domain5",
  "demo_DF1_nm_tract_AHRQ_domain1", "demo_DF1_nm_tract_AHRQ_domain2", "demo_DF1_nm_tract_AHRQ_domain3",
  "demo_DF1_nm_tract_AHRQ_domain4", "demo_DF1_nm_tract_AHRQ_domain5", "M6_total_ahrq_trct_DF1_nm"
)

long_names <- c(
  "Set 1", "Set 2", "Set 3", "Set 4", "Set 5", "Set 6",
  "County: Domain 1", "County: Domain 2", "County: Domain 3", "County: Domain 4", "County: Domain 5",
  "Tract: Domain 1", "Tract: Domain 2", "Tract: Domain 3", "Tract: Domain 4", "Tract: Domain 5", "Tract: All"
)
name_mapping <- setNames(long_names, total_feats_used)
```

## 2. Load and Prepare Data

```{r load-data}
# Load prediction and fairness results
full_hf_results <- read.csv("../scripts/<your_output_filename>.csv") 
fairness_full_hf_results <- read.csv("../scripts/<your_fairness_output_filename>.csv") 

# Subset Prediction Results
logr_hf_results <- full_hf_results %>% filter(model == "logr")
xgb_hf_results  <- full_hf_results %>% filter(model == "xgb")
rf_hf_results   <- full_hf_results %>% filter(model == "rf")

# Subset Fairness Results (Restricting to Race Subgroups)
fairness_race_results <- fairness_full_hf_results %>% filter(sg_key == "race")

fair_logr_hf_results <- fairness_race_results %>% filter(model == "logr")
fair_xgb_hf_results  <- fairness_race_results %>% filter(model == "xgb")
fair_rf_hf_results   <- fairness_race_results %>% filter(model == "rf")

# Bind full dataframes for cross-model comparisons
curr_model_results <- bind_rows(
  logr_hf_results %>% mutate(model = "logr"),
  rf_hf_results %>% mutate(model = "rf"),
  xgb_hf_results %>% mutate(model = "xgb")
)

curr_model_results_fair <- bind_rows(
  fair_logr_hf_results %>% mutate(model = "logr"),
  fair_rf_hf_results %>% mutate(model = "rf"),
  fair_xgb_hf_results %>% mutate(model = "xgb")
)

# Select a primary model for specific model tables (e.g., XGBoost)
primary_model_results <- xgb_hf_results 
```

## 3. Predictive Performance Analysis

### Performance by Subgroup (Primary Model)

```{r table-performance-subgroup}
grouped_results <- primary_model_results %>%
  filter(feat %in% sdoh_feat_sets, subgroup %in% c("blackorwhite", "black", "white")) %>%
  group_by(feat, subgroup) %>%
  summarize(
    auc = mean(auc, na.rm = TRUE),
    f1 = mean(f1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = subgroup,
    values_from = c(auc, f1),
    names_glue = "{subgroup}_{.value}"
  ) %>%
  select(feat, blackorwhite_auc, black_auc, white_auc, blackorwhite_f1, black_f1, white_f1) %>%
  mutate(across(where(is.numeric), ~ sprintf("%.3f", .)))

kable(grouped_results, 
      col.names = c("Feature Set", "AUROC (All)", "AUROC (Black)", "AUROC (White)", "F1 (All)", "F1 (Black)", "F1 (White)"),
      caption = "Primary Model: AUROC and F1 Score by Subgroup")
```

### Performance Across All 3 Models (Black vs White)

```{r table-performance-allmodels}
grouped_results_all <- curr_model_results %>%
  filter(feat %in% sdoh_feat_sets, subgroup %in% c("blackorwhite", "black", "white")) %>%
  group_by(feat, model) %>%
  summarize(
    auroc_all = mean(auc[subgroup == "blackorwhite"], na.rm = TRUE),
    auroc_black = mean(auc[subgroup == "black"], na.rm = TRUE), 
    auroc_white = mean(auc[subgroup == "white"], na.rm = TRUE),
    f1_all = mean(f1[subgroup == "blackorwhite"], na.rm = TRUE),
    f1_black = mean(f1[subgroup == "black"], na.rm = TRUE),
    f1_white = mean(f1[subgroup == "white"], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(across(starts_with(c("auroc_", "f1_")), ~ sprintf("%.3f", .)))

kable(grouped_results_all,
      col.names = c("Feature Set", "Model", "AUROC (All)", "AUROC (Black)", "AUROC (White)", "F1 (All)", "F1 (Black)", "F1 (White)"),
      caption = "Summary of AUROC and F1 Score by Model and Subgroup")
```

## 4. Fairness Analysis

### Fairness Metrics Across All 3 Models

```{r table-fairness-allmodels}
grouped_fair_all <- curr_model_results_fair %>%
  filter(feat %in% sdoh_feat_sets) %>%
  group_by(feat, model) %>%
  summarize(
    eo_ratio = mean(eo_ratio, na.rm = TRUE),
    fpr_parity = mean(fpr_parity, na.rm = TRUE), 
    tpr_parity = mean(tpr_parity, na.rm = TRUE), 
    fnr_parity = mean(fnr_parity, na.rm = TRUE), 
    dp_ratio = mean(dp_ratio, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric), ~ sprintf("%.3f", .)))

kable(grouped_fair_all,
      col.names = c("Feature Set", "Model", "EO Ratio", "FPR Parity", "TPR Parity", "FNR Parity", "DP Ratio"),
      caption = "Summary of Fairness Metrics by Model and Feature Set (Black vs. White)")
```

### Extract Fairness Data for Excel Export
Generates a cleanly formatted CSV of metrics for the primary model rather than printing messy console output.

```{r export-fairness-excel}
curr_feat <- tract_only_feats

excel_fairness_export <- curr_model_results_fair %>%
  filter(feat %in% curr_feat) %>%
  group_by(feat, model) %>%
  summarize(
    mean_eo_ratio = round(mean(eo_ratio, na.rm = TRUE), 3),
    mean_fpr_parity = round(mean(fpr_parity, na.rm = TRUE), 3),
    mean_tpr_parity = round(mean(tpr_parity, na.rm = TRUE), 3),
    mean_fnr_parity = round(mean(fnr_parity, na.rm = TRUE), 3),
    mean_dp_ratio = round(mean(dp_ratio, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  arrange(feat, factor(model, levels = model_order))

write.csv(excel_fairness_export, "Fairness_Metrics_Export.csv", row.names = FALSE)
cat("Fairness metrics successfully exported to 'Fairness_Metrics_Export.csv'\n")
```

## 5. Statistical Comparisons (Wilcoxon Tests)

```{r wilcoxon-tests}
# Helper to safely perform Wilcoxon test, aligning by fold
perform_wilcoxon_test <- function(baseline_df, current_df, metric_col) {
  baseline_values <- baseline_df %>% arrange(fold) %>% pull(!!sym(metric_col))
  current_values <- current_df %>% arrange(fold) %>% pull(!!sym(metric_col))
  
  if (length(baseline_values) == length(current_values) && length(baseline_values) > 0) {
    test <- wilcox.test(baseline_values, current_values, paired = TRUE, alternative = "two.sided")
    return(test$p.value)
  } else {
    return(NA)
  }
}

wilcoxon_results <- data.frame()
baseline_feat <- "M6_total_ahrq_trct_DF1_nm"
feature_sets <- county_only_feats[-1]

metrics_to_test <- c("auc", "f1", "eo_ratio")
results_dfs <- list(
  auc = primary_model_results %>% filter(subgroup == "blackorwhite"),
  f1 = primary_model_results %>% filter(subgroup == "blackorwhite"),
  eo_ratio = fair_xgb_hf_results 
)

for (metric_name in metrics_to_test) {
  full_results <- results_dfs[[metric_name]]
  baseline_data <- full_results %>% filter(feat == baseline_feat)
  
  for (current_feat in feature_sets) {
    current_data <- full_results %>% filter(feat == current_feat)
    p_value <- perform_wilcoxon_test(baseline_data, current_data, metric_name)
    
    wilcoxon_results <- bind_rows(wilcoxon_results, data.frame(
      metric = metric_name, feature_set = current_feat, p_value = p_value
    ))
  }
}

wilcoxon_table <- wilcoxon_results %>%
  pivot_wider(names_from = metric, values_from = p_value, names_glue = "{metric}_p") %>%
  mutate(
    across(where(is.numeric), ~ sprintf("%.3f", .)),
    feature_set = name_mapping[feature_set]
  )

kable(wilcoxon_table, 
      col.names = c("Feature Set", "AUC p-value", "F1 p-value", "EO Ratio p-value"),
      caption = paste("Wilcoxon Signed-Rank Test p-values vs Baseline:", name_mapping[baseline_feat]))
```

## 6. Plotting: Performance vs. Fairness (XGBoost)

```{r plot-prep}
# Filter for XGBoost
plot_pred_df <- primary_model_results %>% filter(subgroup == "blackorwhite", feat %in% total_feats_used)
plot_fair_df <- fair_xgb_hf_results %>% filter(feat %in% total_feats_used)

# Calculate averages
auc_avg <- plot_pred_df %>% group_by(feat) %>% summarize(mean_auc = round(mean(auc), 3))
f1_avg <- plot_pred_df %>% group_by(feat) %>% summarize(mean_f1 = round(mean(f1), 3))
eo_ratio_avg <- plot_fair_df %>% group_by(feat) %>% summarize(mean_eo_ratio = round(mean(eo_ratio), 3))

# Map to long names
auc_avg$feat <- name_mapping[auc_avg$feat]
f1_avg$feat <- name_mapping[f1_avg$feat]
eo_ratio_avg$feat <- name_mapping[eo_ratio_avg$feat]

# Balance Score (AUC * EO Ratio)
balance_scores <- auc_avg$mean_auc * eo_ratio_avg$mean_eo_ratio
best_index <- which.max(balance_scores)

best_feat_set <- auc_avg$feat[best_index]
best_auroc <- auc_avg$mean_auc[best_index]
best_eo <- eo_ratio_avg$mean_eo_ratio[best_index]

cat("Best Balance Feature Set:", best_feat_set, "| AUROC:", best_auroc, "| EO Ratio:", best_eo, "\n")
```

```{r plot-fairness-vs-performance, fig.width=10, fig.height=6}
# Define aesthetics mapping
long_county_feat_sets <- name_mapping[county_feat_sets]
long_county_only_feats <- name_mapping[county_only_feats]
long_tract_only_feats <- name_mapping[tract_only_feats]

colors_county_feat_sets <- brewer.pal(length(long_county_feat_sets), "Set1")
colors_county_only_feats <- brewer.pal(length(long_county_only_feats), "Set2")
colors_tract_only_feats <- brewer.pal(length(long_tract_only_feats), "Set3")

color_mapping <- c(
  setNames(colors_county_feat_sets, long_county_feat_sets),
  setNames(colors_county_only_feats, long_county_only_feats),
  setNames(colors_tract_only_feats, long_tract_only_feats)
)

shape_mapping <- c(
  setNames(rep(19, length(long_county_feat_sets)), long_county_feat_sets),
  setNames(rep(17, length(long_county_only_feats)), long_county_only_feats),
  setNames(rep(15, length(long_tract_only_feats)), long_tract_only_feats)
)

point_colors <- color_mapping[auc_avg$feat]
point_shapes <- shape_mapping[auc_avg$feat]

# Plot configurations
par(mar = c(5, 4, 4, 15) + 0.1)

plot(eo_ratio_avg$mean_eo_ratio, auc_avg$mean_auc, 
     type = "p", col = point_colors, pch = point_shapes, cex = 1.2,
     xlab = "Equal Opportunity (EO) Ratio", ylab = "AUROC", 
     xlim = c(0, 1), ylim = c(0, 1),
     main = "AUROC vs. Equal Opportunity Ratio (XGBoost)")

# Draw attention to maximums
max_auroc <- max(auc_avg$mean_auc)
max_eo <- max(eo_ratio_avg$mean_eo_ratio)

abline(h = max_auroc, col = "black", lty = 2)
abline(v = max_eo, col = "black", lty = 2)

# Highlight Best Balance Point
points(best_eo, best_auroc, col = "red", pch = 19, cex = 1.8)

# Legend Configuration
legend_labels <- c(long_county_feat_sets, long_county_only_feats, long_tract_only_feats)
legend_colors <- color_mapping[legend_labels]
legend_shapes <- shape_mapping[legend_labels]

legend("topright", inset = c(-0.55, 0), legend = legend_labels, 
       col = legend_colors, pch = legend_shapes, xpd = TRUE, 
       title = "Feature Sets", cex = 0.8)
```