---
title: "Patient Inclusion & Geocoding Pipeline"
author: "Chase Fensore"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Setup and Dependencies

```{r libraries}
library(tibble)
library(dplyr)
library(tidyr)
library(caret)
library(censusxy)
library(tidygeocoder)
library(stringr)
library(httr)
library(jsonlite)
library(purrr)

# Load custom data cleaning functions
source('data_cleaners.R')
```

## 1. Load Data and Apply Inclusion Criteria

```{r load-and-filter}
# 1. Read in raw patient data file
orig_df <- read.csv("./<your-raw-patient-data-file.csv>")
cat("Initial cohort size from raw data: ", nrow(orig_df), "\n")

# 2. Restrict to endpoints (Clean up the end-point related columns)
endpoints <- c("readmit30bin", "death30bin", "composite30")
filtered_df <- orig_df

for (colx in endpoints) {
  # Note: Assuming remove_na is a custom function from 'data_cleaners.R'
  filtered_df <- remove_na(filtered_df, colx) 
}
cat("After removing missing endpoints: ", nrow(filtered_df), "\n")

# 3. Exclude invalid street addresses
# Drop if: no value/missing/empty string for *ALL* of the useful columns. 
invalid_values <- c(NA, "", ".", "..", "...")

filtered_df <- filtered_df %>%
  filter(
    !(Patient_Home_Street_Address %in% invalid_values &
      Patient_Address %in% invalid_values &
      fips %in% invalid_values &
      censustract %in% invalid_values)
  )
cat("After invalid address exclusion: ", nrow(filtered_df), "\n")

# 4. Demographic exclusions
filtered_df <- filtered_df %>%
  filter((black == 1 | white == 1), age_current >= 18)

cat("After demographic inclusion (Black/White, Age >= 18): ", nrow(filtered_df), "\n")

# Write out valid encounters
write.csv(filtered_df$Encounter, "encouterIDs_validaddress.csv", row.names = FALSE)
```

## 2. Initial Geocoding Status Check

Identify how many patients already have valid geocodes (Census Tract or FIPS) versus those that need to be processed.

```{r check-geocode-status}
# Load previously geocoded patient cohort
patient_cohort <- read.csv("./final_allstates_modelinput.csv") 

# Standardize data types for safe joining
filtered_df$Encounter <- as.numeric(as.character(filtered_df$Encounter))
patient_cohort$Encounter <- as.numeric(as.character(patient_cohort$Encounter))

# Filter patients missing geocodes 
missing_geocode_df <- filtered_df %>%
  filter(fips %in% invalid_values | censustract %in% invalid_values)

# Patients successfully geocoded previously
non_missing_geocode_df <- anti_join(filtered_df, missing_geocode_df, by = "Encounter")

cat("Patients with successful geocodes: ", nrow(non_missing_geocode_df), "\n")
cat("Patients missing geocodes (require API lookup): ", nrow(missing_geocode_df), "\n")

# Save patients with unsuccessful geocodes for manual/batch review
drop_vars <- c("Patient", "patient_name", "Lon__Lat", "cxy_address", 
               "cxy_matched_address", "cxy_tiger_line_id", "cxy_tract_id", "cxy_block_id")

missing_geocode_export <- missing_geocode_df %>% select(-any_of(drop_vars))
fname <- paste0("./missing_geocodedf_n-", nrow(missing_geocode_export), ".csv")
write.csv(missing_geocode_export, fname, row.names = FALSE)
```

## 3. Secondary Geocoding via Census API

Attempt to geocode the remaining missing patients using their raw string addresses.
*(Note: API integration specifics are abstracted for publication. Contact corresponding author for full pipeline specifics).*

```{r secondary-geocoding}
# Create dataframe for the secondary geocoding attempt
fips_w_geocoded_data <- missing_geocode_df
fips_w_geocoded_data$tract_fips <- NA
fips_w_geocoded_data$county_fips <- NA

# Loop through missing addresses and assign FIPS/Tract via helper function
for (i in 1:nrow(fips_w_geocoded_data)) {
  # get_fips_from_address() is assumed to be a custom wrapper for the Census API
  tractcounty_both <- get_fips_from_address(fips_w_geocoded_data$Patient_Address[i])
  
  fips_w_geocoded_data$tract_fips[i]  <- as.character(tractcounty_both$tract_fips)
  fips_w_geocoded_data$county_fips[i] <- as.character(tractcounty_both$county_fips)
}

# Write out the new round of geocoded patients
write.csv(fips_w_geocoded_data, "ViaAddresses_FIPS_DONE_allstates_geocoded_patient_data.csv", row.names = FALSE)
```

## 4. Re-integrate Newly Geocoded Data

Isolate the successes from the secondary geocoding attempt and merge them back into the primary cohort.

```{r merge-geocoded-data}
# Filter for successful secondary geocodes
nonmissing_fips_w_geocoded_data <- fips_w_geocoded_data %>%
  filter(!(tract_fips %in% invalid_values | county_fips %in% invalid_values))

cat("Extra patients correctly geocoded from address API: ", nrow(nonmissing_fips_w_geocoded_data), "\n")

# Standardize column names to match the primary cohort
nonmissing_fips_w_geocoded_data <- nonmissing_fips_w_geocoded_data %>%
  rename(censustract = tract_fips, fips = county_fips)

# Identify common columns and bind the two distinct datasets
common_columns <- intersect(names(non_missing_geocode_df), names(nonmissing_fips_w_geocoded_data))

final_merged_df <- bind_rows(
  non_missing_geocode_df %>% select(all_of(common_columns)),
  nonmissing_fips_w_geocoded_data %>% select(all_of(common_columns))
)

cat("Total final patients geocoded across both methods: ", nrow(final_merged_df), "\n")

# Track patients who still failed geocoding for Table 1 analysis
failed_second_geocoding <- fips_w_geocoded_data %>%
  filter(!Encounter %in% nonmissing_fips_w_geocoded_data$Encounter)

cat("Patients who failed secondary geocoding: ", nrow(failed_second_geocoding), "\n")

filename_failed <- paste0("failed_geocode_table1_n-", nrow(failed_second_geocoding), ".csv")
write.csv(failed_second_geocoding, filename_failed, row.names = FALSE)
```

## 5. Final Formatting and PHI Removal

Strip Protected Health Information (PHI) and format geographic boundary identifiers.

```{r finalize-data}
# Define PHI columns to be removed (Update this vector as necessary per IRB guidelines)
remove_cols <- c(
  "Patient_Home_Street_Address", 
  "Patient_Address", 
  "Patient_Home_Zip_Code"
  # Add other PHI variables here...
)

# Remove PHI columns safely
final_merged_df <- final_merged_df %>%
  select(-any_of(remove_cols))

# Pad tract and FIPS values to ensure correct digit length and extract state num
final_merged_df <- final_merged_df %>%
  mutate(
    censustract = str_pad(as.character(censustract), width = 11, pad = "0"),
    fips = str_pad(as.character(fips), width = 5, pad = "0"),
    statenum = substr(censustract, 1, 2)
  )

# Write final non-missing cohort
write.csv(final_merged_df, "Nonmissing_Cohort.csv", row.names = FALSE)
cat("Final cohort saved to 'Nonmissing_Cohort.csv'. Pipeline complete.\n")
```