---
title: "Table 1 Generation & Cohort Comparisons"
author: "Chase Fensore"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Setup and Dependencies

```{r libraries}
library(caret)
library(data.table)
library(tidyverse) # Includes dplyr, tidyr, stringr, readr, ggplot2, tibble
library(tableone)
library(kableExtra)

# Suppress scientific notation for cleaner tables
options(scipen = 999)
```

## 1. Helper Functions for Data Formatting

Instead of repeating renaming operations for different dataframes, we define a standard processing function to unify column names and demographics across all cohorts.

```{r formatting-function}
# Function to safely convert specified columns to numeric, forcing errors to NA
convert_nonnumeric_to_na <- function(df, columns) {
  for(col in columns) {
    if(col %in% names(df)) {
      df[[col]] <- suppressWarnings(as.numeric(as.character(df[[col]])))
    }
  }
  return(df)
}

# Function to rename columns and engineer demographic factors for Table 1
format_patient_data <- function(df) {
  df <- df %>%
    # Filter out 'other' race if it exists, keeping only Black and White per inclusion criteria
    filter(is.na(other) | other != 1) %>%
    mutate(
      Race = case_when(white == 1 ~ "White", black == 1 ~ "Black", TRUE ~ NA_character_),
      `Insurance type` = case_when(
        privinsurance == 1 ~ "Private", 
        Medicare == 1 ~ "Medicare", 
        Medicaid == 1 ~ "Medicaid", 
        TRUE ~ "Not recorded"
      ),
      `HF classification` = case_when(
        hfsystolic == 1 ~ "Systolic (HFrEF)", 
        hfdiastolic == 1 ~ "Diastolic (HFpEF)", 
        hfother == 1 ~ "Other", 
        TRUE ~ "Not recorded"
      )
    ) %>%
    rename(
      `Age, years` = age_current,
      Female = female,
      Male = male,
      Black = black,
      White = white,
      `30-day Mortality` = death30bin,
      Hypertension = htn,
      `Coronary Artery Disease` = cad,
      `Chronic Kidney Disease` = ckd,
      `Charlson Comorbidity Index` = cci,
      `Systolic blood pressure, mm Hg` = Systolic_BP,
      `Heart Rate, beats per minute` = Heart_Rate,
      `Respiratory rate, breaths per minute` = Resp,
      `Diastolic Blood Pressure (20-200), mmHg` = Diastolic_BP,
      `B-type natriuretic peptide (BNP), pg/mL` = BNP,
      `Blood urea nitrogen (BUN) test, mg/dL` = BUN,
      Albumin = ALB,
      `Hemoglobin, g/dL` = Hb,
      `Hemoglobin A1C (0-20), %` = HbA1C,
      `Potassium, mEq/L` = K,
      `Serum Creatinine, mg/dL` = Creat,
      `Sodium, mEq/L` = Na,
      `Social deprivation index (percentile)` = sdi,
      `Area deprivation index, national rank (percentile)` = ADI_national_rank,
      `Area deprivation index, state rank (percentile)` = ADI_state_rank,
      `Social vulnerability index score (composite score)` = RPL_THEMES,
      `CVD Specialist` = specialcvd,
      `Int. Med. Specialist` = specialint,
      `EME Specialist` = specialeme,
      `Other Specialist` = specialoth,
      `UH Floor` = loceuhfloor,
      `CLH ICU` = loceclhicu,
      `CLH Floor` = loceclhfloor,
      `UH ICU` = loceuhicu
    )
  
  return(df)
}

# Variable lists used across all tables
contVars <- c("Age, years", "Systolic blood pressure, mm Hg", "Heart Rate, beats per minute", 
              "Respiratory rate, breaths per minute", "Diastolic Blood Pressure (20-200), mmHg",
              "B-type natriuretic peptide (BNP), pg/mL", "Blood urea nitrogen (BUN) test, mg/dL",
              "Albumin", "Hemoglobin, g/dL", "Hemoglobin A1C (0-20), %", "Potassium, mEq/L",
              "Serum Creatinine, mg/dL", "Sodium, mEq/L", "LOScheck",
              "Social deprivation index (percentile)", 
              "Area deprivation index, national rank (percentile)", 
              "Area deprivation index, state rank (percentile)", 
              "Social vulnerability index score (composite score)")

catVars <- c("Male", "Female", "Black", "White", "30-day Mortality", "Insurance type", 
             "HF classification", "Hypertension", "Coronary Artery Disease", "Chronic Kidney Disease",
             "CVD Specialist", "Int. Med. Specialist", "EME Specialist", "Other Specialist",
             "UH Floor", "CLH ICU", "CLH Floor", "UH ICU")
```

## 2. Load and Prepare Cohort Data

```{r load-data}
# 2.1 Load Main (Geocoded) Cohort
input_ahrq_patient_feats <- read.csv("../data/final_allstates_modelinput.csv") 
trans_input_feats <- format_patient_data(input_ahrq_patient_feats)

# 2.2 Load Unsuccessful/Missing Geocode Cohort
nonmissing_raw <- read.csv("../data/nonmissing_geocodedf_n-3102.csv")

# Clean deprivation columns in the missing cohort
deprivation_cols <- c("percnt_ltfpl100", "percnt_singlparntfly", "percnt_dropout", 
                      "percnt_hhnocar", "percnt_rentoccup", "percnt_crowding", 
                      "percnt_nonemp", "percnt_unemp", "percnt_highneeds", 
                      "percnt_hispanic", "percnt_frgnborn", "percnt_black", 
                      "sdi", "ADI_state_rank", "ADI_national_rank")

nonmissing_raw <- convert_nonnumeric_to_na(nonmissing_raw, deprivation_cols)
nonmissing_df <- format_patient_data(nonmissing_raw)
```

## 3. Table 1: Main Cohort (Stratified by 30-Day Readmission)

```{r table1-main}
# Ensure categorical variables are treated as factors
trans_input_feats <- trans_input_feats %>%
  mutate(across(all_of(catVars), as.factor))

# Filter to variables that actually exist in the dataframe to avoid errors
current_contVars <- intersect(contVars, names(trans_input_feats))
current_catVars <- intersect(catVars, names(trans_input_feats))

# Generate TableOne
tab1 <- CreateTableOne(
  vars = c(current_contVars, current_catVars), 
  strata = "readmit30bin", 
  data = trans_input_feats, 
  addOverall = TRUE,
  test = TRUE,
  smd = TRUE
)

# Print and format table
p_main <- print(tab1, printToggle = FALSE, noSpaces = TRUE, showAllLevels = TRUE)
kable(p_main, format = "markdown", booktabs = TRUE, caption = "Table 1: Baseline Characteristics by 30-Day Readmission Status")
```

## 4. Table 1 Comparison: Geocoded vs. Non-Geocoded Patients

```{r table1-comparison}
# Identify common columns to bind datasets safely
common_columns <- intersect(names(trans_input_feats), names(nonmissing_df))

# Subset and label cohorts
trans_subset <- trans_input_feats %>% 
  select(all_of(common_columns)) %>% 
  mutate(cohort = "Geocoded")

nonmissing_subset <- nonmissing_df %>% 
  select(all_of(common_columns)) %>% 
  mutate(cohort = "Non-Geocoded")

# Combine datasets
combined_df <- bind_rows(trans_subset, nonmissing_subset)

# Ensure categorical variables are factors
combined_catVars <- intersect(catVars, names(combined_df))
combined_contVars <- intersect(contVars, names(combined_df))

combined_df <- combined_df %>%
  mutate(across(all_of(combined_catVars), as.factor))

# Generate Comparison TableOne
comprehensive_table <- CreateTableOne(
    vars = c(combined_contVars, combined_catVars), 
    strata = "cohort", 
    data = combined_df, 
    addOverall = TRUE,
    test = TRUE, 
    smd = TRUE 
)

# Print and format table
print_comprehensive <- print(comprehensive_table, printToggle = FALSE, noSpaces = TRUE, showAllLevels = TRUE)

kable(print_comprehensive, format = "markdown", booktabs = TRUE, caption = "Comparison: Successfully Geocoded vs. Unsuccessfully Geocoded Patients")

# Export to CSV for manuscript inclusion
write.csv(print_comprehensive, "cohort_comparison_table1.csv", row.names = TRUE)
cat("Comparison table exported to 'cohort_comparison_table1.csv'\n")
```